---
title: "Data preperation"
author: "Julius Fenn"
format:
  html:
    toc: true
    toc-depth: 3
    html-math-method: katex
    number-sections: true
---

# Notes

This script prepares the pretest vignette data and outputs a single clean table
that only includes fully completed participants (JATOS FINISHED + all required
responses + 5 word associations).

# global variables

```{r}
#| label: global-variables

createRawFiles <- TRUE

data_root <- "data"
output_root <- "outputs"
raw_output_dir <- file.path(output_root, "raw")
questionnaire_output_dir <- file.path(output_root, "questionnaire")
associations_output_dir <- file.path(output_root, "associations")
```

# functions

```{r}
########################################
# from JATOS data set to classical data frame (2D)
########################################
### args:
# dataset = dat
# listvars = vec_ques
# notNumeric = vec_notNumeric
# verbose = TRUE

questionnairetype <- function(dataset,
                              listvars,
                              notNumeric,
                              verbose = FALSE) {
  datasetques <- data.frame(ID = unique(dataset$ID))

  for (c in seq_along(listvars)) {
    if (verbose) {
      print(c)
    }

    if (any(colnames(dataset) == listvars[c])) {
      if (verbose) {
        print(listvars[c])
      }

      tmpid <- dataset$ID[!is.na(dataset[, listvars[c]])]
      tmpvalue <- dataset[, listvars[c]][!is.na(dataset[, listvars[c]])]
      datasetques[listvars[c]] <- NA

      if (listvars[c] %in% notNumeric) {
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- tmpvalue
      } else if (is.list(tmpvalue)) {
        tmpvalue_tmp <- unique(tmpvalue)
        tmpvalue <- c()
        counter <- 1
        for (i in seq_along(tmpvalue_tmp)) {
          if (!is.null(tmpvalue_tmp[[i]])) {
            tmpvalue[counter] <- paste0(tmpvalue_tmp[[i]], collapse = " - ")
            counter <- counter + 1
          }
        }
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- tmpvalue
      } else {
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- as.numeric(tmpvalue)
      }
    }
  }
  return(datasetques)
}

########################################
# helpers
########################################

get_last_non_null <- function(x) {
  if (length(x) == 0) return(NULL)
  idx <- which(!purrr::map_lgl(x, is.null))
  if (length(idx) == 0) return(NULL)
  x[[idx[length(idx)]]]
}

parse_duration_seconds <- function(x) {
  if (is.na(x) || x == "") return(NA_real_)
  clean <- stringr::str_remove(x, "\\s*\\(.*\\)")
  if (is.na(clean) || clean == "") return(NA_real_)
  as.numeric(hms::as_hms(clean))
}

association_to_wide <- function(assoc_list) {
  if (is.null(assoc_list) || length(assoc_list) == 0) {
    return(tibble::tibble(
      assoc_cue = NA_character_,
      assoc1_response = NA_character_, assoc1_timestamp = NA_character_, assoc1_valence = NA_character_,
      assoc2_response = NA_character_, assoc2_timestamp = NA_character_, assoc2_valence = NA_character_,
      assoc3_response = NA_character_, assoc3_timestamp = NA_character_, assoc3_valence = NA_character_,
      assoc4_response = NA_character_, assoc4_timestamp = NA_character_, assoc4_valence = NA_character_,
      assoc5_response = NA_character_, assoc5_timestamp = NA_character_, assoc5_valence = NA_character_
    ))
  }

  assoc_df <- dplyr::as_tibble(assoc_list)
  assoc_df <- assoc_df[seq_len(min(5, nrow(assoc_df))), ]

  cue_value <- if ("cue" %in% names(assoc_df)) assoc_df$cue[1] else NA_character_

  out <- tibble::tibble(
    assoc_cue = cue_value,
    assoc1_response = assoc_df$response[1], assoc1_timestamp = assoc_df$timestamp[1], assoc1_valence = assoc_df$valence[1],
    assoc2_response = assoc_df$response[2], assoc2_timestamp = assoc_df$timestamp[2], assoc2_valence = assoc_df$valence[2],
    assoc3_response = assoc_df$response[3], assoc3_timestamp = assoc_df$timestamp[3], assoc3_valence = assoc_df$valence[3],
    assoc4_response = assoc_df$response[4], assoc4_timestamp = assoc_df$timestamp[4], assoc4_valence = assoc_df$valence[4],
    assoc5_response = assoc_df$response[5], assoc5_timestamp = assoc_df$timestamp[5], assoc5_valence = assoc_df$valence[5]
  )

  return(out)
}
```

# load packages

```{r}
#| label: load-packages
#| warning: false

require(pacman)
p_load(
  "tidyverse",
  "jsonlite",
  "hms",
  "writexl",
  "readxl",
  "haven"
)
```

# create raw data files

```{r}
#| label: process-studies
#| warning: false

if (createRawFiles) {
  dir.create(raw_output_dir, recursive = TRUE, showWarnings = FALSE)

  data_files <- list.files(
    data_root,
    pattern = "data.txt$",
    recursive = TRUE,
    full.names = TRUE
  )

  results <- list()
  for (file in data_files) {
    result_folder <- basename(dirname(dirname(file)))
    result_id <- stringr::str_remove(result_folder, "^study_result_")

    tmp <- jsonlite::fromJSON(file)
    tmp[, colnames(tmp) == ""] <- NULL
    tmp <- dplyr::as_tibble(tmp)
    tmp$result_id <- as.character(result_id)

    results[[length(results) + 1]] <- tmp
  }

  saveRDS(results, file = file.path(raw_output_dir, "study.rds"))
  json_lines <- purrr::map_chr(results, ~ jsonlite::toJSON(.x, auto_unbox = TRUE))
  writeLines(json_lines, con = file.path(raw_output_dir, "study.jsonl"))

  print("Raw files have been created in this run!")
} else {
  print("Raw files have been created in previous run!")
}
```

# load and prepare raw data

```{r}
#| label: load-raw-data

study <- readRDS(file.path(raw_output_dir, "study.rds"))
dat <- suppressMessages(bind_rows(study))
rm(study)



dat$ID <- dat$result_id

dat <- dat %>%
  dplyr::mutate(
    meta_language = purrr::map_chr(meta, ~ if (is.null(.x$language)) NA_character_ else .x$language),
    meta_screen_width = purrr::map_dbl(meta, ~ as.numeric(if (is.null(.x$screen_width)) NA else .x$screen_width)),
    meta_screen_height = purrr::map_dbl(meta, ~ as.numeric(if (is.null(.x$screen_height)) NA else .x$screen_height)),
    meta_userAgent = purrr::map_chr(meta, ~ if (is.null(.x$userAgent)) NA_character_ else .x$userAgent),
    meta_platform = purrr::map_chr(meta, ~ if (is.null(.x$platform)) NA_character_ else .x$platform),
    meta_timeZone = purrr::map_chr(meta, ~ if (is.null(.x$timeZone)) NA_character_ else .x$timeZone),
    meta_timezoneOffset = purrr::map_dbl(meta, ~ as.numeric(if (is.null(.x$timezoneOffset)) NA else .x$timezoneOffset)),
    meta_window_innerWidth = purrr::map_dbl(meta, ~ as.numeric(if (is.null(.x$window_innerWidth)) NA else .x$window_innerWidth)),
    meta_window_innerHeight = purrr::map_dbl(meta, ~ as.numeric(if (is.null(.x$window_innerHeight)) NA else .x$window_innerHeight)),
    meta_devicePixelRatio = purrr::map_dbl(meta, ~ as.numeric(if (is.null(.x$devicePixelRatio)) NA else .x$devicePixelRatio))
  )
```

# load and join JATOS metadata

```{r}
#| label: jatos-metadata

metadata_files <- list.files(
  data_root,
  pattern = "jatos_results_metadata_.*\\.csv$",
  full.names = TRUE
)

metadata_file <- metadata_files[which.max(file.info(metadata_files)$mtime)]
metadata <- readr::read_csv(metadata_file, show_col_types = FALSE)
names(metadata) <- make.names(names(metadata))

metadata <- metadata %>%
  mutate(
    result_id = as.character(Result.ID),
    duration_seconds = purrr::map_dbl(Duration, parse_duration_seconds)
  )

dat <- dat %>%
  left_join(metadata, by = "result_id")
```

# prepare questionnaire data

```{r}
#| label: prepare-questionnaire

dat_cols <- colnames(dat)

vec_visualtrap <- stringr::str_subset(dat_cols, "^visualTrap_")
vec_visualtrap_response <- stringr::str_subset(dat_cols, "^visualTrap_.*_response$")
vec_attributes <- stringr::str_subset(dat_cols, "^attributesFutureSociety-\\d+$")
vec_attcheck <- stringr::str_subset(dat_cols, "^attCheck_")

vec_notNumeric <- c(
  "ID",
  "result_id",
  "PROLIFIC_PID",
  "futureSocietyCondition",
  "cue",
  "cue_coding",
  "meta_language",
  "meta_userAgent",
  "meta_platform",
  "meta_timeZone",
  "dummy_informedconsent",
  "hp_exclusionCriteria",
  "hp_exclusionCriteria_value",
  "commCheck",
  "clearUtopiatext",
  "biasUtopiatext",
  "undUtopiaGeneral",
  "undUtopiaGoal",
  "undUtopiaHow",
  "utopiaPrototypeAssignment",
  "protoConfidenceRadio",
  "protoReasonText",
  "State",
  "Start.time",
  "Last.seen",
  "Duration",
  vec_attcheck,
  vec_visualtrap
)

vec_ques <- c(
  "ID",
  "result_id",
  "PROLIFIC_PID",
  "futureSocietyCondition",
  "cue",
  "cue_coding",
  "dummy_informedconsent",
  "hp_exclusionCriteria",
  "hp_exclusionCriteria_value",
  "commCheck",
  "clearUtopia",
  "clearUtopiatext",
  "biasUtopia",
  "biasUtopiatext",
  "undUtopiaGeneral",
  "undUtopiaGoal",
  "undUtopiaHow",
  vec_attributes,
  "utopiaPrototypeAssignment",
  "protoConfidenceRadio",
  "protoReasonText",
  vec_visualtrap,
  "meta_language",
  "meta_screen_width",
  "meta_screen_height",
  "meta_userAgent",
  "meta_platform",
  "meta_timeZone",
  "meta_timezoneOffset",
  "meta_window_innerWidth",
  "meta_window_innerHeight",
  "meta_devicePixelRatio",
  "State",
  "Start.time",
  "Last.seen",
  "Duration",
  "duration_seconds",
  vec_attcheck
)

ques_study <- questionnairetype(
  dataset = dat,
  listvars = vec_ques,
  notNumeric = vec_notNumeric,
  verbose = FALSE
)

ques_study <- ques_study %>%
  mutate(across(where(is.character), ~ na_if(., "")))
```

# prepare associations (wide) and merge

```{r}
#| label: prepare-associations

assoc_by_id <- dat %>%
  group_by(ID) %>%
  summarise(
    sucsessfulAssociations = get_last_non_null(sucsessfulAssociations),
    unsucsessfulAssociations = get_last_non_null(unsucsessfulAssociations),
    .groups = "drop"
  )

assoc_wide <- assoc_by_id %>%
  mutate(assoc_row = purrr::map(sucsessfulAssociations, association_to_wide)) %>%
  select(ID, assoc_row) %>%
  tidyr::unnest_wider(assoc_row)

dat_clean <- ques_study %>%
  left_join(assoc_wide, by = "ID")
```

# filter to complete datasets

```{r}
#| label: complete-filter

required_cols <- c(
  "PROLIFIC_PID",
  "futureSocietyCondition",
  "undUtopiaGeneral",
  "undUtopiaGoal",
  "undUtopiaHow",
  "clearUtopia",
  "biasUtopia",
  "utopiaPrototypeAssignment",
  "protoConfidenceRadio",
  vec_attributes,
  vec_visualtrap_response,
  "assoc1_response",
  "assoc2_response",
  "assoc3_response",
  "assoc4_response",
  "assoc5_response"
)

is_missing <- function(x) {
  if (is.character(x) || is.factor(x)) {
    is.na(x) | x == ""
  } else {
    is.na(x)
  }
}

clean_complete <- dat_clean %>%
  filter(State == "FINISHED") %>%
  filter(!if_any(all_of(required_cols), is_missing))

dim(clean_complete)
```

# save data sets

```{r}
#| label: save-outputs

dir.create(questionnaire_output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(associations_output_dir, recursive = TRUE, showWarnings = FALSE)

saveRDS(clean_complete, file = file.path(questionnaire_output_dir, "clean_complete.rds"))
writexl::write_xlsx(x = clean_complete, path = file.path(questionnaire_output_dir, "clean_complete.xlsx"))
write.csv(clean_complete, file = file.path(questionnaire_output_dir, "clean_complete.csv"), row.names = FALSE)
```
