---
title: "Data preperation"
author: "Julius Fenn"
format:
  html:
    toc: true
    toc-depth: 3
    html-math-method: katex
    number-sections: true
---

# Notes

This script creates a simplified participant-level dataset using the raw
`study.rds` file and produces:

- A clean wide table with core variables + associations (wide) + keyStats
- Paradata events in long format (clipboard, mouse, scroll, keyIntervals)
- Defocus events in long format

# global variables

```{r}
#| label: global-variables

createRawFiles <- TRUE

data_root <- "data"
output_root <- "outputs"
raw_output_dir <- file.path(output_root, "raw")
questionnaire_output_dir <- file.path(output_root, "questionnaire")
paradata_output_dir <- file.path(output_root, "paradata")
```

# functions

```{r}
last_non_empty <- function(x) {
  if (length(x) == 0) return(NA)
  for (i in seq_along(x)[length(x):1]) {
    val <- x[[i]]
    if (is.null(val)) next
    if (is.list(val) && length(val) == 0) next
    if (!is.list(val) && length(val) == 1 && is.na(val)) next
    if (is.character(val) && length(val) == 1 && isTRUE(val == "")) next
    return(val)
  }
  return(NA)
}

`%||%` <- function(x, y) {
  if (is.null(x)) y else x
}

parse_duration_seconds <- function(x) {
  if (is.na(x) || x == "") return(NA_real_)
  clean <- stringr::str_remove(x, "\\s*\\(.*\\)")
  if (is.na(clean) || clean == "") return(NA_real_)
  as.numeric(hms::as_hms(clean))
}

null_to_empty <- function(x) {
  if (is.null(x)) list() else x
}

normalize_cell <- function(val) {
  if (is.null(val)) return(NA)
  if (is.data.frame(val)) return(list(val))
  if (is.list(val)) {
    if (length(val) == 0) return(NA)
    return(list(val))
  }
  if (length(val) != 1) return(list(val))
  val
}

collapse_list_to_string <- function(x, sep = " // ") {
  if (is.null(x)) return(NA_character_)
  if (is.list(x)) {
    if (length(x) == 0) return(NA_character_)
    flat <- unlist(x, recursive = TRUE, use.names = FALSE)
    if (length(flat) == 0) return(NA_character_)
    return(paste0(flat, collapse = sep))
  }
  if (length(x) > 1) return(paste0(x, collapse = sep))
  as.character(x)
}

sanitize_paradata_name <- function(name) {
  cleaned <- gsub("[^A-Za-z0-9]+", "_", name)
  cleaned <- gsub("^_+|_+$", "", cleaned)
  paste0("para_", cleaned)
}

init_paradata_result <- function() {
  list(
    wide_vals = list(),
    wide_cols = character(),
    wide_cols_output = character(),
    long_rows = list(),
    long_cols = character()
  )
}

append_paradata_result <- function(target, source) {
  target$wide_vals <- c(target$wide_vals, source$wide_vals)
  target$wide_cols <- c(target$wide_cols, source$wide_cols)
  target$wide_cols_output <- c(target$wide_cols_output, source$wide_cols_output)
  target$long_rows <- c(target$long_rows, source$long_rows)
  target$long_cols <- c(target$long_cols, source$long_cols)
  target
}

add_paradata_wide <- function(res, field_label, value) {
  out_col <- sanitize_paradata_name(field_label)
  res$wide_vals[[out_col]] <- value
  res$wide_cols <- c(res$wide_cols, field_label)
  res$wide_cols_output <- c(res$wide_cols_output, out_col)
  res
}

add_paradata_long_row <- function(res, pid, result_id, field_label, value, row_index, timestamp) {
  res$long_rows[[length(res$long_rows) + 1]] <- tibble::tibble(
    PROLIFIC_PID = pid,
    result_id = result_id,
    origin = "paradata_general",
    event_type = "paradata_column",
    event_index = row_index,
    timestamp = timestamp %||% NA_character_,
    component = NA_character_,
    field = field_label,
    value = value,
    mouse_x = NA_real_,
    mouse_y = NA_real_,
    scroll_x = NA_real_,
    scroll_y = NA_real_,
    key_interval_ms = NA_real_,
    clipboard_type = NA_character_,
    clipboard_text = NA_character_,
    clipboard_length = NA_real_,
    target_tag = NA_character_,
    target_type = NA_character_,
    target_id = NA_character_,
    target_name = NA_character_,
    target_className = NA_character_
  )
  res$long_cols <- c(res$long_cols, field_label)
  res
}

get_event_value <- function(ev, name) {
  if (is.null(ev)) return(NA)
  if (is.data.frame(ev)) return(ev[[name]] %||% NA)
  if (is.list(ev)) return(ev[[name]] %||% NA)
  NA
}

get_target_value <- function(ev, field) {
  if (is.null(ev)) return(NA)
  if (is.data.frame(ev)) {
    col <- paste0("target.", field)
    return(ev[[col]] %||% NA)
  }
  if (is.list(ev) && !is.null(ev$target)) return(ev$target[[field]] %||% NA)
  NA
}

is_non_empty_element <- function(x) {
  if (is.null(x)) return(FALSE)
  if (is.data.frame(x)) return(nrow(x) > 0)
  if (is.list(x)) {
    if (length(x) == 0) return(FALSE)
    return(!all(purrr::map_lgl(x, ~ is.null(.x) || (is.list(.x) && length(.x) == 0))))
  }
  length(x) > 0
}

find_last_non_empty_element <- function(col_vals) {
  if (is.null(col_vals)) return(list(value = NULL, index = NA_integer_))
  if (!is.list(col_vals)) {
    if (length(col_vals) == 0) return(list(value = NULL, index = NA_integer_))
    return(list(value = col_vals, index = length(col_vals)))
  }
  for (i in seq_along(col_vals)[length(col_vals):1]) {
    v <- col_vals[[i]]
    if (is_non_empty_element(v)) return(list(value = v, index = i))
  }
  list(value = NULL, index = NA_integer_)
}

process_group_df <- function(df, prefix_label, timestamp_vec, pid, result_id) {
  res <- init_paradata_result()
  if (is.null(df) || !is.data.frame(df)) return(res)

  for (col in names(df)) {
    col_vals <- df[[col]]
    field_label <- paste0(prefix_label, ".", col)

    if (is.data.frame(col_vals)) {
      nested_res <- process_group_df(col_vals, field_label, timestamp_vec, pid, result_id)
      res <- append_paradata_result(res, nested_res)
      next
    }

    if (is.list(col_vals)) {
      vals_json <- vapply(col_vals, function(v) {
        if (is.null(v) || (is.list(v) && length(v) == 0)) return(NA_character_)
        jsonlite::toJSON(v, auto_unbox = TRUE)
      }, character(1))

      non_na <- vals_json[!is.na(vals_json)]
      if (length(non_na) == 0) next
      uniq <- unique(non_na)

      if (length(uniq) == 1) {
        res <- add_paradata_wide(res, field_label, uniq[[1]])
      } else {
        for (r in seq_along(vals_json)) {
          v <- vals_json[[r]]
          if (is.na(v)) next
          ts <- if (!is.null(timestamp_vec) && length(timestamp_vec) >= r) timestamp_vec[[r]] else NA_character_
          res <- add_paradata_long_row(res, pid, result_id, field_label, v, r, ts)
        }
      }
      next
    }

    non_na <- col_vals[!is.na(col_vals)]
    if (length(non_na) == 0) next
    uniq <- unique(non_na)

    if (length(uniq) == 1) {
      res <- add_paradata_wide(res, field_label, uniq[[1]])
    } else {
      for (r in seq_along(col_vals)) {
        v <- col_vals[[r]]
        if (is.na(v)) next
        ts <- if (!is.null(timestamp_vec) && length(timestamp_vec) >= r) timestamp_vec[[r]] else NA_character_
        res <- add_paradata_long_row(res, pid, result_id, field_label, as.character(v), r, ts)
      }
    }
  }

  res
}

flatten_with_prefix <- function(x, prefix) {
  if (is.null(x)) return(list())
  flat <- unlist(x, recursive = TRUE, use.names = TRUE)
  if (length(flat) == 0) return(list())
  names(flat) <- paste0(prefix, names(flat))
  as.list(flat)
}

assoc_to_wide <- function(assoc_list) {
  if (is.null(assoc_list) || length(assoc_list) == 0) {
    return(tibble::tibble(
      assoc1_response = NA_character_, assoc1_valence = NA_character_,
      assoc2_response = NA_character_, assoc2_valence = NA_character_,
      assoc3_response = NA_character_, assoc3_valence = NA_character_,
      assoc4_response = NA_character_, assoc4_valence = NA_character_,
      assoc5_response = NA_character_, assoc5_valence = NA_character_
    ))
  }

  assoc_df <- dplyr::as_tibble(assoc_list)
  assoc_df <- assoc_df[seq_len(min(5, nrow(assoc_df))), ]

  get_val <- function(df, col, idx) {
    if (!col %in% names(df)) return(NA_character_)
    if (nrow(df) < idx) return(NA_character_)
    df[[col]][idx]
  }

  tibble::tibble(
    assoc1_response = get_val(assoc_df, "response", 1),
    assoc1_valence = get_val(assoc_df, "valence", 1),
    assoc2_response = get_val(assoc_df, "response", 2),
    assoc2_valence = get_val(assoc_df, "valence", 2),
    assoc3_response = get_val(assoc_df, "response", 3),
    assoc3_valence = get_val(assoc_df, "valence", 3),
    assoc4_response = get_val(assoc_df, "response", 4),
    assoc4_valence = get_val(assoc_df, "valence", 4),
    assoc5_response = get_val(assoc_df, "response", 5),
    assoc5_valence = get_val(assoc_df, "valence", 5)
  )
}

build_paradata_events <- function(pid, result_id, paradata) {
  if (is.null(paradata) || is.null(paradata$behavior)) return(tibble::tibble())

  events <- list()
  behavior <- paradata$behavior

  clipboard_events <- null_to_empty(behavior$clipboardEvents)
  if (length(clipboard_events)) {
    events[[length(events) + 1]] <- purrr::map_dfr(seq_along(clipboard_events), function(i) {
      ev <- clipboard_events[[i]]
      target <- ev$target
      tibble::tibble(
        PROLIFIC_PID = pid,
        result_id = result_id,
        origin = "paradata_general",
        event_type = "clipboard",
        event_index = i,
        timestamp = ev$timestamp %||% NA_character_,
      component = ev$component %||% NA_character_,
      field = NA_character_,
      value = NA_character_,
      mouse_x = NA_real_,
      mouse_y = NA_real_,
        scroll_x = NA_real_,
        scroll_y = NA_real_,
        key_interval_ms = NA_real_,
        clipboard_type = ev$type %||% NA_character_,
        clipboard_text = ev$text %||% NA_character_,
        clipboard_length = ev$length %||% NA_real_,
        target_tag = target$tag %||% NA_character_,
        target_type = target$type %||% NA_character_,
        target_id = target$id %||% NA_character_,
        target_name = target$name %||% NA_character_,
        target_className = target$className %||% NA_character_
      )
    })
  }

  mouse_events <- null_to_empty(behavior$mouseEvents)
  if (length(mouse_events)) {
    events[[length(events) + 1]] <- purrr::map_dfr(seq_along(mouse_events), function(i) {
      ev <- mouse_events[[i]]
      tibble::tibble(
        PROLIFIC_PID = pid,
        result_id = result_id,
        origin = "paradata_general",
        event_type = "mouse",
        event_index = i,
        timestamp = ev$timestamp %||% NA_character_,
      component = ev$component %||% NA_character_,
      field = NA_character_,
      value = NA_character_,
      mouse_x = ev$x %||% NA_real_,
      mouse_y = ev$y %||% NA_real_,
        scroll_x = NA_real_,
        scroll_y = NA_real_,
        key_interval_ms = NA_real_,
        clipboard_type = NA_character_,
        clipboard_text = NA_character_,
        clipboard_length = NA_real_,
        target_tag = NA_character_,
        target_type = NA_character_,
        target_id = NA_character_,
        target_name = NA_character_,
        target_className = NA_character_
      )
    })
  }

  scroll_events <- null_to_empty(behavior$scrollEvents)
  if (length(scroll_events)) {
    events[[length(events) + 1]] <- purrr::map_dfr(seq_along(scroll_events), function(i) {
      ev <- scroll_events[[i]]
      tibble::tibble(
        PROLIFIC_PID = pid,
        result_id = result_id,
        origin = "paradata_general",
        event_type = "scroll",
        event_index = i,
        timestamp = ev$timestamp %||% NA_character_,
      component = ev$component %||% NA_character_,
      field = NA_character_,
      value = NA_character_,
      mouse_x = NA_real_,
      mouse_y = NA_real_,
        scroll_x = ev$scrollX %||% NA_real_,
        scroll_y = ev$scrollY %||% NA_real_,
        key_interval_ms = NA_real_,
        clipboard_type = NA_character_,
        clipboard_text = NA_character_,
        clipboard_length = NA_real_,
        target_tag = NA_character_,
        target_type = NA_character_,
        target_id = NA_character_,
        target_name = NA_character_,
        target_className = NA_character_
      )
    })
  }


  if (length(events) == 0) return(tibble::tibble())
  dplyr::bind_rows(events)
}

process_paradata_general <- function(pid, result_id, paradata) {
  empty <- list(
    wide = tibble::tibble(),
    long = tibble::tibble(),
    wide_cols = character(),
    wide_cols_output = character(),
    long_cols = character(),
    key_stats = tibble::tibble(
      keyStats_count = NA_real_,
      keyStats_mean = NA_real_,
      keyStats_median = NA_real_,
      keyStats_min = NA_real_,
      keyStats_max = NA_real_
    )
  )

  if (is.null(paradata)) return(empty)

  if (is.data.frame(paradata) && all(c("static", "ip", "behavior", "meta") %in% names(paradata))) {
    static_df <- paradata$static
    ip_df <- paradata$ip
    behavior_df <- paradata$behavior
    meta_df <- paradata$meta
    ip_error <- paradata$ipError

    timestamps <- NULL
    if (is.data.frame(static_df) && "timestamp" %in% names(static_df)) {
      timestamps <- static_df$timestamp
    } else if (is.data.frame(meta_df) && "updatedAt" %in% names(meta_df)) {
      timestamps <- meta_df$updatedAt
    }

    res <- init_paradata_result()
    res <- append_paradata_result(res, process_group_df(static_df, "static", timestamps, pid, result_id))
    res <- append_paradata_result(res, process_group_df(ip_df, "ip", timestamps, pid, result_id))
    res <- append_paradata_result(res, process_group_df(meta_df, "meta", timestamps, pid, result_id))

    if (!is.null(ip_error)) {
      non_na <- ip_error[!is.na(ip_error)]
      if (length(non_na) == 1) {
        res <- add_paradata_wide(res, "ipError", non_na[[1]])
      } else if (length(non_na) > 1) {
        for (r in seq_along(ip_error)) {
          v <- ip_error[[r]]
          if (is.na(v)) next
          ts <- if (!is.null(timestamps) && length(timestamps) >= r) timestamps[[r]] else NA_character_
          res <- add_paradata_long_row(res, pid, result_id, "ipError", as.character(v), r, ts)
        }
      }
    }

    key_stats <- empty$key_stats
    if (is.data.frame(behavior_df) && "keyStats" %in% names(behavior_df)) {
      key_df <- behavior_df$keyStats
      if (is.data.frame(key_df)) {
        key_stats <- tibble::tibble(
          keyStats_count = if ("count" %in% names(key_df)) last_non_empty(key_df$count) else NA_real_,
          keyStats_mean = if ("mean" %in% names(key_df)) last_non_empty(key_df$mean) else NA_real_,
          keyStats_median = if ("median" %in% names(key_df)) last_non_empty(key_df$median) else NA_real_,
          keyStats_min = if ("min" %in% names(key_df)) last_non_empty(key_df$min) else NA_real_,
          keyStats_max = if ("max" %in% names(key_df)) last_non_empty(key_df$max) else NA_real_
        )
      }
    }

    long_rows <- list()
    if (is.data.frame(behavior_df)) {
      if ("clipboardEvents" %in% names(behavior_df)) {
        last_clip <- find_last_non_empty_element(behavior_df$clipboardEvents)
        ev_list <- last_clip$value
        if (!is.null(ev_list)) {
          if (is.data.frame(ev_list)) {
            for (j in seq_len(nrow(ev_list))) {
              ev <- ev_list[j, ]
              long_rows[[length(long_rows) + 1]] <- tibble::tibble(
                PROLIFIC_PID = pid,
                result_id = result_id,
                origin = "paradata_general",
                event_type = "clipboard",
                event_index = j,
                timestamp = get_event_value(ev, "timestamp") %||% (if (!is.null(timestamps) && !is.na(last_clip$index) && length(timestamps) >= last_clip$index) timestamps[[last_clip$index]] else NA_character_),
                component = get_event_value(ev, "component"),
                field = NA_character_,
                value = NA_character_,
                mouse_x = NA_real_,
                mouse_y = NA_real_,
                scroll_x = NA_real_,
                scroll_y = NA_real_,
                key_interval_ms = NA_real_,
                clipboard_type = get_event_value(ev, "type"),
                clipboard_text = get_event_value(ev, "text"),
                clipboard_length = get_event_value(ev, "length"),
                target_tag = get_target_value(ev, "tag"),
                target_type = get_target_value(ev, "type"),
                target_id = get_target_value(ev, "id"),
                target_name = get_target_value(ev, "name"),
                target_className = get_target_value(ev, "className")
              )
            }
            res$long_cols <- c(res$long_cols, "behavior.clipboardEvents")
          } else if (is.list(ev_list)) {
            ev_list <- null_to_empty(ev_list)
            for (j in seq_along(ev_list)) {
              ev <- ev_list[[j]]
              long_rows[[length(long_rows) + 1]] <- tibble::tibble(
                PROLIFIC_PID = pid,
                result_id = result_id,
                origin = "paradata_general",
                event_type = "clipboard",
                event_index = j,
                timestamp = get_event_value(ev, "timestamp") %||% (if (!is.null(timestamps) && !is.na(last_clip$index) && length(timestamps) >= last_clip$index) timestamps[[last_clip$index]] else NA_character_),
                component = get_event_value(ev, "component"),
                field = NA_character_,
                value = NA_character_,
                mouse_x = NA_real_,
                mouse_y = NA_real_,
                scroll_x = NA_real_,
                scroll_y = NA_real_,
                key_interval_ms = NA_real_,
                clipboard_type = get_event_value(ev, "type"),
                clipboard_text = get_event_value(ev, "text"),
                clipboard_length = get_event_value(ev, "length"),
                target_tag = get_target_value(ev, "tag"),
                target_type = get_target_value(ev, "type"),
                target_id = get_target_value(ev, "id"),
                target_name = get_target_value(ev, "name"),
                target_className = get_target_value(ev, "className")
              )
            }
            res$long_cols <- c(res$long_cols, "behavior.clipboardEvents")
          }
        }
      }

      if ("mouseEvents" %in% names(behavior_df)) {
        last_mouse <- find_last_non_empty_element(behavior_df$mouseEvents)
        ev_df <- last_mouse$value
        if (!is.null(ev_df)) {
          if (is.data.frame(ev_df)) {
            for (j in seq_len(nrow(ev_df))) {
              ev <- ev_df[j, ]
              long_rows[[length(long_rows) + 1]] <- tibble::tibble(
                PROLIFIC_PID = pid,
                result_id = result_id,
                origin = "paradata_general",
                event_type = "mouse",
                event_index = j,
                timestamp = ev$timestamp %||% (if (!is.null(timestamps) && !is.na(last_mouse$index) && length(timestamps) >= last_mouse$index) timestamps[[last_mouse$index]] else NA_character_),
                component = ev$component %||% NA_character_,
                field = NA_character_,
                value = NA_character_,
                mouse_x = ev$x %||% NA_real_,
                mouse_y = ev$y %||% NA_real_,
                scroll_x = NA_real_,
                scroll_y = NA_real_,
                key_interval_ms = NA_real_,
                clipboard_type = NA_character_,
                clipboard_text = NA_character_,
                clipboard_length = NA_real_,
                target_tag = NA_character_,
                target_type = NA_character_,
                target_id = NA_character_,
                target_name = NA_character_,
                target_className = NA_character_
              )
            }
            res$long_cols <- c(res$long_cols, "behavior.mouseEvents")
          } else if (is.list(ev_df)) {
            ev_df <- null_to_empty(ev_df)
            for (j in seq_along(ev_df)) {
              ev <- ev_df[[j]]
              long_rows[[length(long_rows) + 1]] <- tibble::tibble(
                PROLIFIC_PID = pid,
                result_id = result_id,
                origin = "paradata_general",
                event_type = "mouse",
                event_index = j,
                timestamp = get_event_value(ev, "timestamp") %||% (if (!is.null(timestamps) && !is.na(last_mouse$index) && length(timestamps) >= last_mouse$index) timestamps[[last_mouse$index]] else NA_character_),
                component = get_event_value(ev, "component"),
                field = NA_character_,
                value = NA_character_,
                mouse_x = get_event_value(ev, "x"),
                mouse_y = get_event_value(ev, "y"),
                scroll_x = NA_real_,
                scroll_y = NA_real_,
                key_interval_ms = NA_real_,
                clipboard_type = NA_character_,
                clipboard_text = NA_character_,
                clipboard_length = NA_real_,
                target_tag = NA_character_,
                target_type = NA_character_,
                target_id = NA_character_,
                target_name = NA_character_,
                target_className = NA_character_
              )
            }
            res$long_cols <- c(res$long_cols, "behavior.mouseEvents")
          }
        }
      }

      if ("scrollEvents" %in% names(behavior_df)) {
        last_scroll <- find_last_non_empty_element(behavior_df$scrollEvents)
        ev_df <- last_scroll$value
        if (!is.null(ev_df)) {
          if (is.data.frame(ev_df)) {
            for (j in seq_len(nrow(ev_df))) {
              ev <- ev_df[j, ]
              long_rows[[length(long_rows) + 1]] <- tibble::tibble(
                PROLIFIC_PID = pid,
                result_id = result_id,
                origin = "paradata_general",
                event_type = "scroll",
                event_index = j,
                timestamp = ev$timestamp %||% (if (!is.null(timestamps) && !is.na(last_scroll$index) && length(timestamps) >= last_scroll$index) timestamps[[last_scroll$index]] else NA_character_),
                component = ev$component %||% NA_character_,
                field = NA_character_,
                value = NA_character_,
                mouse_x = NA_real_,
                mouse_y = NA_real_,
                scroll_x = ev$scrollX %||% NA_real_,
                scroll_y = ev$scrollY %||% NA_real_,
                key_interval_ms = NA_real_,
                clipboard_type = NA_character_,
                clipboard_text = NA_character_,
                clipboard_length = NA_real_,
                target_tag = NA_character_,
                target_type = NA_character_,
                target_id = NA_character_,
                target_name = NA_character_,
                target_className = NA_character_
              )
            }
            res$long_cols <- c(res$long_cols, "behavior.scrollEvents")
          } else if (is.list(ev_df)) {
            ev_df <- null_to_empty(ev_df)
            for (j in seq_along(ev_df)) {
              ev <- ev_df[[j]]
              long_rows[[length(long_rows) + 1]] <- tibble::tibble(
                PROLIFIC_PID = pid,
                result_id = result_id,
                origin = "paradata_general",
                event_type = "scroll",
                event_index = j,
                timestamp = get_event_value(ev, "timestamp") %||% (if (!is.null(timestamps) && !is.na(last_scroll$index) && length(timestamps) >= last_scroll$index) timestamps[[last_scroll$index]] else NA_character_),
                component = get_event_value(ev, "component"),
                field = NA_character_,
                value = NA_character_,
                mouse_x = NA_real_,
                mouse_y = NA_real_,
                scroll_x = get_event_value(ev, "scrollX"),
                scroll_y = get_event_value(ev, "scrollY"),
                key_interval_ms = NA_real_,
                clipboard_type = NA_character_,
                clipboard_text = NA_character_,
                clipboard_length = NA_real_,
                target_tag = NA_character_,
                target_type = NA_character_,
                target_id = NA_character_,
                target_name = NA_character_,
                target_className = NA_character_
              )
            }
            res$long_cols <- c(res$long_cols, "behavior.scrollEvents")
          }
        }
      }

    }

    if (length(long_rows)) {
      res$long_rows <- c(res$long_rows, long_rows)
    }

    return(list(
      wide = if (length(res$wide_vals)) tibble::as_tibble(res$wide_vals) else tibble::tibble(),
      long = if (length(res$long_rows)) dplyr::bind_rows(res$long_rows) else tibble::tibble(),
      wide_cols = res$wide_cols,
      wide_cols_output = res$wide_cols_output,
      long_cols = res$long_cols,
      key_stats = key_stats
    ))
  }

  if (is.list(paradata)) {
    wide_vals <- c(
      flatten_with_prefix(paradata$static, "para_static_"),
      flatten_with_prefix(paradata$ip, "para_ip_"),
      flatten_with_prefix(paradata$meta, "para_meta_")
    )

    key_stats <- tibble::tibble(
      keyStats_count = paradata$behavior$keyStats$count %||% NA_real_,
      keyStats_mean = paradata$behavior$keyStats$mean %||% NA_real_,
      keyStats_median = paradata$behavior$keyStats$median %||% NA_real_,
      keyStats_min = paradata$behavior$keyStats$min %||% NA_real_,
      keyStats_max = paradata$behavior$keyStats$max %||% NA_real_
    )

    return(list(
      wide = if (length(wide_vals)) tibble::as_tibble(wide_vals) else tibble::tibble(),
      long = build_paradata_events(pid, result_id, paradata),
      wide_cols = names(wide_vals),
      wide_cols_output = names(wide_vals),
      long_cols = character(),
      key_stats = key_stats
    ))
  }

  empty
}

build_defocus_events <- function(pid, result_id, defocus_list) {
  if (is.null(defocus_list)) return(tibble::tibble())

  if (is.data.frame(defocus_list)) {
    if (nrow(defocus_list) == 0) return(tibble::tibble())
    return(dplyr::mutate(
      tibble::as_tibble(defocus_list),
      PROLIFIC_PID = pid,
      result_id = result_id,
      origin = "paradata_focus",
      event_index = dplyr::row_number()
    ) %>%
      dplyr::select(PROLIFIC_PID, result_id, origin, event_index, durationblur, senderblur))
  }

  if (is.list(defocus_list)) {
    if (length(defocus_list) == 0) return(tibble::tibble())
    return(purrr::map_dfr(seq_along(defocus_list), function(i) {
      ev <- defocus_list[[i]]
      tibble::tibble(
        PROLIFIC_PID = pid,
        result_id = result_id,
        origin = "paradata_focus",
        event_index = i,
        durationblur = ev$durationblur %||% NA_real_,
        senderblur = ev$senderblur %||% NA_character_
      )
    }))
  }

  tibble::tibble()
}

is_missing_value <- function(x) {
  if (is.null(x)) return(TRUE)
  if (is.list(x)) {
    return(length(x) == 0 || all(is.na(unlist(x))))
  }
  if (length(x) > 1) {
    if (is.character(x)) return(all(is.na(x) | x == ""))
    return(all(is.na(x)))
  }
  if (is.character(x)) return(is.na(x) | x == "")
  is.na(x)
}
```

# load packages

```{r}
#| label: load-packages
#| warning: false

require(pacman)
p_load(
  "tidyverse",
  "jsonlite",
  "hms",
  "psych",
  "afex",
  "writexl",
  "readxl",
  "haven"
)
```

# create raw data files

```{r}
#| label: process-studies
#| warning: false

if (createRawFiles) {
  dir.create(raw_output_dir, recursive = TRUE, showWarnings = FALSE)

  data_files <- list.files(
    data_root,
    pattern = "data.txt$",
    recursive = TRUE,
    full.names = TRUE
  )

  results <- list()
  for (file in data_files) {
    result_folder <- basename(dirname(dirname(file)))
    result_id <- stringr::str_remove(result_folder, "^study_result_")

    tmp <- jsonlite::fromJSON(file)
    tmp[, colnames(tmp) == ""] <- NULL
    tmp <- dplyr::as_tibble(tmp)
    tmp$result_id <- as.character(result_id)

    results[[length(results) + 1]] <- tmp
  }

  saveRDS(results, file = file.path(raw_output_dir, "study.rds"))
  json_lines <- purrr::map_chr(results, ~ jsonlite::toJSON(.x, auto_unbox = TRUE))
  writeLines(json_lines, con = file.path(raw_output_dir, "study.jsonl"))

  print("Raw files have been created in this run!")
} else {
  print("Raw files have been created in previous run!")
}
```

# load JATOS metadata

```{r}
#| label: jatos-metadata

metadata_files <- list.files(
  data_root,
  pattern = "jatos_results_metadata_.*\\.csv$",
  full.names = TRUE
)

metadata_file <- metadata_files[which.max(file.info(metadata_files)$mtime)]
metadata <- readr::read_csv(metadata_file, show_col_types = FALSE)
names(metadata) <- make.names(names(metadata))

metadata <- metadata %>%
  mutate(
    result_id = as.character(Result.ID),
    duration_seconds = purrr::map_dbl(Duration, parse_duration_seconds)
  )

metadata_small <- metadata %>%
  select(result_id, State, Duration, duration_seconds)
```

# build participant-level data

```{r}
#| label: build-participant-data

study <- readRDS(file.path(raw_output_dir, "study.rds"))

core_vars <- c(
  "dummy_informedconsent",
  "hp_exclusionCriteria",
  "commCheck",
  "counterVisualTraps",
  "visualTrapResponse",
  "visualTrap_Planets_response",
  "visualTrap_Planets_responseLabel",
  "visualTrap_Planets_correct",
  "visualTrap_Planets_rtMs",
  "visualTrap_Collision_response",
  "visualTrap_Collision_responseLabel",
  "visualTrap_Collision_correct",
  "visualTrap_Collision_rtMs",
  "visualTrap_CafeWall_response",
  "visualTrap_CafeWall_responseLabel",
  "visualTrap_CafeWall_correct",
  "visualTrap_CafeWall_rtMs",
  "visualTrap_Circles_response",
  "visualTrap_Circles_responseLabel",
  "visualTrap_Circles_correct",
  "visualTrap_Circles_rtMs",
  "visualTrap_Lines_response",
  "visualTrap_Lines_responseLabel",
  "visualTrap_Lines_correct",
  "visualTrap_Lines_rtMs",
  "visualTrap_Robot_response",
  "visualTrap_Robot_responseLabel",
  "visualTrap_Robot_correct",
  "visualTrap_Robot_rtMs",
  "visualTrap_order",
  "condition_FutureSociety",
  "cue",
  "cue_coding",
  "R1",
  "R2",
  "R3",
  "R4",
  "R5",
  "para_countclicks",
  "AR-2",
  "AR-5",
  "AR-3",
  "AR-1",
  "AR-4",
  "undUtopiaGeneral",
  "undUtopiaGoal",
  "undUtopiaHow",
  "clearUtopia",
  "clearUtopiatext",
  "biasUtopia",
  "biasUtopiatext",
  "attributesFutureSociety-5",
  "attributesFutureSociety-6",
  "attributesFutureSociety-2",
  "attributesFutureSociety-3",
  "attributesFutureSociety-4",
  "attributesFutureSociety-7",
  "attributesFutureSociety-1",
  "attributesFutureSociety-8",
  "utopiaPrototypeAssignment",
  "protoConfidenceRadio",
  "protoReasonText",
  "feedback_critic"
)

core_rows <- list()
paradata_events_rows <- list()
paradata_defocus_rows <- list()
paradata_wide_cols_rows <- list()
paradata_long_cols_rows <- list()

for (i in seq_along(study)) {
  print(i)
  df <- study[[i]]

  pid <- if ("PROLIFIC_PID" %in% names(df)) last_non_empty(df$PROLIFIC_PID) else NA
  result_id <- if ("result_id" %in% names(df)) last_non_empty(df$result_id) else NA

  core_values <- lapply(core_vars, function(v) {
    val <- if (v %in% names(df)) {
      last_non_empty(df[[v]])
    } else if (v == "attributesFutureSociety-8" && "attributesFutureSociety-undefined" %in% names(df)) {
      last_non_empty(df[["attributesFutureSociety-undefined"]])
    } else {
      NA
    }
    normalize_cell(val)
  })
  names(core_values) <- core_vars
  core_row <- tibble::as_tibble(core_values)

  if ("visualTrap_order" %in% names(core_row)) {
    core_row$visualTrap_order <- collapse_list_to_string(core_row$visualTrap_order)
  }

  assoc_list <- if ("sucsessfulAssociations" %in% names(df)) last_non_empty(df$sucsessfulAssociations) else NULL
  assoc_wide <- assoc_to_wide(assoc_list)

  paradata <- if ("paradata_general" %in% names(df)) df$paradata_general else NULL
  paradata_summary <- process_paradata_general(pid, result_id, paradata)
  paradata_wide <- paradata_summary$wide
  key_stats_row <- paradata_summary$key_stats

  core_row <- dplyr::bind_cols(
    tibble::tibble(result_id = result_id, PROLIFIC_PID = pid),
    core_row,
    assoc_wide,
    paradata_wide,
    key_stats_row
  )

  core_rows[[length(core_rows) + 1]] <- core_row

  paradata_events <- paradata_summary$long
  if (nrow(paradata_events) > 0) {
    paradata_events_rows[[length(paradata_events_rows) + 1]] <- paradata_events
  }

  if (length(paradata_summary$wide_cols) > 0) {
    paradata_wide_cols_rows[[length(paradata_wide_cols_rows) + 1]] <- tibble::tibble(
      PROLIFIC_PID = pid,
      result_id = result_id,
      field = paradata_summary$wide_cols,
      output_field = paradata_summary$wide_cols_output,
      destination = "wide"
    )
  }

  if (length(paradata_summary$long_cols) > 0) {
    paradata_long_cols_rows[[length(paradata_long_cols_rows) + 1]] <- tibble::tibble(
      PROLIFIC_PID = pid,
      result_id = result_id,
      field = paradata_summary$long_cols,
      output_field = NA_character_,
      destination = "long"
    )
  }

  defocus_list <- if ("para_defocuscount" %in% names(df)) last_non_empty(df$para_defocuscount) else NULL
  defocus_events <- build_defocus_events(pid, result_id, defocus_list)
  if (nrow(defocus_events) > 0) {
    paradata_defocus_rows[[length(paradata_defocus_rows) + 1]] <- defocus_events
  }
}

core_df <- dplyr::bind_rows(core_rows)
paradata_events_long <- dplyr::bind_rows(paradata_events_rows)
paradata_defocus_long <- dplyr::bind_rows(paradata_defocus_rows)
paradata_column_map <- dplyr::bind_rows(paradata_wide_cols_rows, paradata_long_cols_rows)

if (nrow(paradata_events_long) == 0) {
  paradata_events_long <- tibble::tibble(
    PROLIFIC_PID = character(),
    result_id = character(),
    origin = character(),
    event_type = character(),
    event_index = integer(),
    timestamp = character(),
    component = character(),
    field = character(),
    value = character(),
    mouse_x = double(),
    mouse_y = double(),
    scroll_x = double(),
    scroll_y = double(),
    key_interval_ms = double(),
    clipboard_type = character(),
    clipboard_text = character(),
    clipboard_length = double(),
    target_tag = character(),
    target_type = character(),
    target_id = character(),
    target_name = character(),
    target_className = character()
  )
}

if (nrow(paradata_defocus_long) == 0) {
  paradata_defocus_long <- tibble::tibble(
    PROLIFIC_PID = character(),
    result_id = character(),
    origin = character(),
    event_index = integer(),
    durationblur = double(),
    senderblur = character()
  )
}

if (nrow(paradata_column_map) == 0) {
  paradata_column_map <- tibble::tibble(
    PROLIFIC_PID = character(),
    result_id = character(),
    field = character(),
    output_field = character(),
    destination = character()
  )
}

core_df <- core_df %>%
  left_join(metadata_small, by = "result_id")
```

# filter to complete datasets

```{r}
#| label: complete-filter

optional_text <- c("clearUtopiatext", "biasUtopiatext", "protoReasonText", "feedback_critic")

required_cols <- setdiff(
  c("PROLIFIC_PID",
    core_vars,
    "assoc1_response", "assoc2_response", "assoc3_response", "assoc4_response", "assoc5_response",
    "assoc1_valence", "assoc2_valence", "assoc3_valence", "assoc4_valence", "assoc5_valence"
  ),
  optional_text
)

clean_complete <- core_df %>%
  filter(State == "FINISHED")

dim(clean_complete)
```


```{r}
#| label: type-coercion

to_numeric <- function(x) {
  if (is.numeric(x)) return(x)
  suppressWarnings(as.numeric(as.character(x)))
}

to_logical <- function(x) {
  if (is.logical(x)) return(x)
  if (is.factor(x)) x <- as.character(x)
  as.logical(x)
}

num_cols <- c(
  "counterVisualTraps",
  "para_countclicks",
  "visualTrap_Planets_rtMs",
  "visualTrap_Collision_rtMs",
  "visualTrap_CafeWall_rtMs",
  "visualTrap_Circles_rtMs",
  "visualTrap_Lines_rtMs",
  "visualTrap_Robot_rtMs",
  "AR-1",
  "AR-2",
  "AR-3",
  "AR-4",
  "AR-5",
  "clearUtopia",
  "biasUtopia",
  "attributesFutureSociety-1",
  "attributesFutureSociety-2",
  "attributesFutureSociety-3",
  "attributesFutureSociety-4",
  "attributesFutureSociety-5",
  "attributesFutureSociety-6",
  "attributesFutureSociety-7",
  "attributesFutureSociety-8",
  "protoConfidenceRadio",
  "assoc1_valence",
  "assoc2_valence",
  "assoc3_valence",
  "assoc4_valence",
  "assoc5_valence",
  "keyStats_count",
  "keyStats_mean",
  "keyStats_median",
  "keyStats_min",
  "keyStats_max",
  "duration_seconds"
)

bool_cols <- c(
  "hp_exclusionCriteria",
  "visualTrap_Planets_correct",
  "visualTrap_Collision_correct",
  "visualTrap_CafeWall_correct",
  "visualTrap_Circles_correct",
  "visualTrap_Lines_correct",
  "visualTrap_Robot_correct"
)

factor_cols <- c(
  "dummy_informedconsent",
  "commCheck",
  "condition_FutureSociety",
  "cue_coding",
  "utopiaPrototypeAssignment",
  "visualTrapResponse",
  "visualTrap_Planets_response",
  "visualTrap_Collision_response",
  "visualTrap_CafeWall_response",
  "visualTrap_Circles_response",
  "visualTrap_Lines_response",
  "visualTrap_Robot_response"
)

clean_complete <- clean_complete %>%
  mutate(
    across(any_of(bool_cols), ~ to_logical(.)),
    across(any_of(num_cols), ~ to_numeric(.)),
    across(any_of(factor_cols), ~ as.factor(.))
  )
```

# save outputs

```{r}
#| label: save-outputs

dir.create(questionnaire_output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(paradata_output_dir, recursive = TRUE, showWarnings = FALSE)

saveRDS(clean_complete, file = file.path(questionnaire_output_dir, "clean_complete_wide.rds"))
writexl::write_xlsx(x = clean_complete, path = file.path(questionnaire_output_dir, "clean_complete_wide.xlsx"))
write.csv(clean_complete, file = file.path(questionnaire_output_dir, "clean_complete_wide.csv"), row.names = FALSE)

saveRDS(paradata_events_long, file = file.path(paradata_output_dir, "paradata_events_long.rds"))
writexl::write_xlsx(x = paradata_events_long, path = file.path(paradata_output_dir, "paradata_events_long.xlsx"))
write.csv(paradata_events_long, file = file.path(paradata_output_dir, "paradata_events_long.csv"), row.names = FALSE)

saveRDS(paradata_defocus_long, file = file.path(paradata_output_dir, "paradata_defocus_long.rds"))
writexl::write_xlsx(x = paradata_defocus_long, path = file.path(paradata_output_dir, "paradata_defocus_long.xlsx"))
write.csv(paradata_defocus_long, file = file.path(paradata_output_dir, "paradata_defocus_long.csv"), row.names = FALSE)
```

# Summary tables

```{r}
#| label: summary-tables

paradata_column_summary <- paradata_column_map %>%
  group_by(destination, field, output_field) %>%
  summarise(n_participants = dplyr::n_distinct(PROLIFIC_PID), .groups = "drop") %>%
  arrange(destination, field)

paradata_participant_summary <- core_df %>%
  select(PROLIFIC_PID, result_id, State, duration_seconds) %>%
  distinct() %>%
  left_join(paradata_events_long %>% count(PROLIFIC_PID, name = "paradata_event_rows"), by = "PROLIFIC_PID") %>%
  left_join(paradata_defocus_long %>% count(PROLIFIC_PID, name = "defocus_event_rows"), by = "PROLIFIC_PID") %>%
  left_join(paradata_column_map %>% filter(destination == "wide") %>% count(PROLIFIC_PID, name = "paradata_wide_fields"), by = "PROLIFIC_PID") %>%
  left_join(paradata_column_map %>% filter(destination == "long") %>% count(PROLIFIC_PID, name = "paradata_long_fields"), by = "PROLIFIC_PID") %>%
  mutate(
    paradata_event_rows = tidyr::replace_na(paradata_event_rows, 0),
    defocus_event_rows = tidyr::replace_na(defocus_event_rows, 0),
    paradata_wide_fields = tidyr::replace_na(paradata_wide_fields, 0),
    paradata_long_fields = tidyr::replace_na(paradata_long_fields, 0)
  )

paradata_column_summary
paradata_participant_summary
```

# Summary statistics

```{r}
#| label: summary-statistics

summary_stats_df <- clean_complete %>%
  select(
    PROLIFIC_PID,
    duration_seconds,
    keyStats_count,
    keyStats_mean,
    keyStats_median,
    keyStats_min,
    keyStats_max
  ) %>%
  left_join(
    paradata_participant_summary %>%
      select(PROLIFIC_PID, paradata_event_rows, defocus_event_rows, paradata_wide_fields, paradata_long_fields),
    by = "PROLIFIC_PID"
  ) %>%
  mutate(
    paradata_event_rows = tidyr::replace_na(paradata_event_rows, 0),
    defocus_event_rows = tidyr::replace_na(defocus_event_rows, 0),
    paradata_wide_fields = tidyr::replace_na(paradata_wide_fields, 0),
    paradata_long_fields = tidyr::replace_na(paradata_long_fields, 0)
  )

psych::describe(summary_stats_df %>% select(where(is.numeric)))
```

```{r}
#| label: summary-plots-anova

assoc_cols <- c("assoc1_valence", "assoc2_valence", "assoc3_valence", "assoc4_valence", "assoc5_valence")
attr_cols <- paste0("attributesFutureSociety-", 1:8)

clean_complete <- clean_complete %>%
  mutate(
    assoc_valence_mean = rowMeans(select(., all_of(assoc_cols)), na.rm = FALSE),
    attributes_mean = rowMeans(select(., all_of(attr_cols)), na.rm = FALSE)
  )

analysis_df <- clean_complete %>%
  select(PROLIFIC_PID, condition_FutureSociety, assoc_valence_mean, attributes_mean, all_of(attr_cols)) %>%
  drop_na(assoc_valence_mean, attributes_mean, condition_FutureSociety)

mean_summary <- analysis_df %>%
  pivot_longer(
    cols = c(assoc_valence_mean, attributes_mean),
    names_to = "measure",
    values_to = "value"
  ) %>%
  group_by(condition_FutureSociety, measure) %>%
  summarise(
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = dplyr::n(),
    se = sd / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(
    measure = recode(
      measure,
      assoc_valence_mean = "Association Valence Mean",
      attributes_mean = "Attributes Mean"
    )
  )

ggplot(mean_summary, aes(x = condition_FutureSociety, y = mean, fill = measure)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(
    aes(ymin = mean - se, ymax = mean + se),
    position = position_dodge(width = 0.8),
    width = 0.2
  ) +
  labs(x = "Condition", y = "Mean", fill = "Measure") +
  theme_minimal()

attr_labels <- c(
  "1" = "Utopian",
  "2" = "Desirable",
  "3" = "Ideal",
  "4" = "Beneficial for the greater good",
  "5" = "Imaginative",
  "6" = "Innovative",
  "7" = "Creative",
  "8" = "Possible"
)

attr_long <- analysis_df %>%
  pivot_longer(
    cols = all_of(attr_cols),
    names_to = "item",
    values_to = "value"
  ) %>%
  mutate(
    item_num = stringr::str_extract(item, "[0-9]+"),
    item_label = recode(item_num, !!!attr_labels)
  )

attr_effects <- attr_long %>%
  filter(!is.na(value)) %>%
  group_by(item_label) %>%
  summarise(
    eta_sq = {
      fit <- aov(value ~ condition_FutureSociety)
      ss <- summary(fit)[[1]]
      ss_between <- ss["condition_FutureSociety", "Sum Sq"]
      ss_total <- sum(ss[, "Sum Sq"])
      if (is.na(ss_between) || is.na(ss_total) || ss_total == 0) NA_real_ else ss_between / ss_total
    },
    .groups = "drop"
  )

attr_effects_table <- attr_effects %>%
  arrange(desc(eta_sq))

attr_effects_table

item_order <- attr_effects %>%
  mutate(eta_order = ifelse(is.na(eta_sq), -Inf, eta_sq)) %>%
  arrange(desc(eta_order)) %>%
  pull(item_label)

attr_summary <- attr_long %>%
  group_by(condition_FutureSociety, item_label) %>%
  summarise(
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = dplyr::n(),
    se = sd / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(item_label = factor(item_label, levels = item_order))

condition_eta <- analysis_df %>%
  group_by(condition_FutureSociety) %>%
  summarise(
    mean = mean(attributes_mean, na.rm = TRUE),
    n = dplyr::n(),
    .groups = "drop"
  ) %>%
  mutate(
    grand_mean = mean(analysis_df$attributes_mean, na.rm = TRUE),
    ss_level = n * (mean - grand_mean)^2
  ) %>%
  mutate(
    ss_total = sum((analysis_df$attributes_mean - grand_mean)^2, na.rm = TRUE),
    eta_sq = ifelse(ss_total == 0, NA_real_, ss_level / ss_total)
  ) %>%
  arrange(desc(eta_sq))

condition_eta

condition_order <- condition_eta %>%
  pull(condition_FutureSociety)

attr_summary <- attr_summary %>%
  mutate(condition_FutureSociety = factor(condition_FutureSociety, levels = condition_order))

ggplot(attr_summary, aes(x = item_label, y = mean, fill = item_label)) +
  geom_col(width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  geom_text(aes(label = round(mean, 2)),
            position = position_stack(vjust = 0.5),
            size = 3) +
  facet_wrap(~ condition_FutureSociety, ncol = 2) +
  labs(x = "Attribute", y = "Mean", fill = "Attribute") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 30, hjust = 1))

anova_assoc <- afex::aov_ez(
  id = "PROLIFIC_PID",
  dv = "assoc_valence_mean",
  between = "condition_FutureSociety",
  data = analysis_df
)
anova_attr <- afex::aov_ez(
  id = "PROLIFIC_PID",
  dv = "attributes_mean",
  between = "condition_FutureSociety",
  data = analysis_df
)

anova_assoc
anova_attr
```
