---
title: "Data preperation"
author: "Julius Fenn"
format:
  html:
    toc: true
    toc-depth: 3
    html-math-method: katex
    number-sections: true
---

# Notes

This script creates a simplified participant-level dataset using the raw
`study.rds` file and produces:

- A clean wide table with core variables + associations (wide) + keyStats
- Paradata events in long format (clipboard, mouse, scroll, keyIntervals)
- Defocus events in long format

# global variables

```{r}
#| label: global-variables

data_root <- "../01_dataPreperation/outputs"
output_root <- "outputs"
```

# functions

```{r}
# not needed
```

# load packages

```{r}
#| label: load-packages
#| warning: false

require(pacman)
p_load(
  "tidyverse",
  "psych",
  "afex",
  "mlr3",
  "mlr3cluster",
  "cluster",
  "writexl",
  "readxl",
  "DT"
)
```

# load processed data

```{r}
#| label: load-data
#| warning: false

setwd(data_root)
setwd("questionnaire")

dat_ques <- readRDS(file = "clean_complete_wide.rds")

setwd("../paradata")
dat_para_defocus <- readRDS(file = "paradata_defocus_long.rds")
dat_para_events <- readRDS(file = "paradata_events_long.rds")
```


# Descriptive Statistics

## sample


```{r}
#| label: descriptive-stats-sample-1

socidem_vars <- str_subset(string = colnames(dat_ques), pattern = "^socidem")
socidem_vars

socidem_df <- dat_ques %>%
  select(all_of(socidem_vars))

socidem_overview <- tibble(
  variable = socidem_vars,
  class = purrr::map_chr(socidem_df, ~ paste(class(.x), collapse = "/")),
  n = nrow(socidem_df),
  n_missing = purrr::map_int(socidem_df, ~ sum(is.na(.x))),
  missing_pct = round(100 * n_missing / n, 2),
  n_unique = purrr::map_int(socidem_df, ~ dplyr::n_distinct(.x, na.rm = TRUE))
)

socidem_descriptives <- purrr::map_dfr(socidem_vars, function(v) {
  x <- socidem_df[[v]]

  if (is.numeric(x)) {
    n_non_missing <- sum(!is.na(x))

    tibble(
      variable = v,
      type = "numeric",
      mean = if (n_non_missing == 0) NA_real_ else mean(x, na.rm = TRUE),
      sd = if (n_non_missing <= 1) NA_real_ else sd(x, na.rm = TRUE),
      median = if (n_non_missing == 0) NA_real_ else median(x, na.rm = TRUE),
      min = if (n_non_missing == 0) NA_real_ else min(x, na.rm = TRUE),
      max = if (n_non_missing == 0) NA_real_ else max(x, na.rm = TRUE),
      top_levels = NA_character_
    )
  } else {
    x_chr <- as.character(x)
    x_chr[x_chr == ""] <- NA_character_
    tab <- sort(table(x_chr, useNA = "no"), decreasing = TRUE)
    k <- min(5, length(tab))

    tibble(
      variable = v,
      type = "categorical_or_text",
      mean = NA_real_,
      sd = NA_real_,
      median = NA_real_,
      min = NA_real_,
      max = NA_real_,
      top_levels = if (k == 0) {
        NA_character_
      } else {
        paste0(names(tab)[seq_len(k)], " (", as.integer(tab[seq_len(k)]), ")", collapse = "; ")
      }
    )
  }
})

socidem_table <- socidem_overview %>%
  left_join(socidem_descriptives, by = "variable")


DT::datatable(data = socidem_table)
```


```{r}
#| label: descriptive-stats-sample-2

## factor vars:
factor_vars <- as.character(socidem_table$variable[socidem_table$class == "factor"])

for (var_name in factor_vars) {
  cat("\nTable for:", var_name, "\n")
  
  tab <- table(socidem_df[[var_name]], useNA = "ifany")
  pct <- prop.table(tab) * 100
  
  print(cbind(Frequency = tab, Percentage = round(pct, 1)))
}

## numeric vars:
numeric_vars <- as.character(socidem_table$variable[socidem_table$class == "numeric"])

for (var_name in numeric_vars) {
  if (var_name %in% names(socidem_df)) {
    x <- socidem_df[[var_name]]
    x <- x[!is.na(x)]
    
    stats <- boxplot.stats(x)$stats
    names(stats) <- c("Min", "Q1", "Median", "Q3", "Max")
    
    mean_x <- mean(x)
    sd_x <- sd(x)
    
    boxplot(
      x,
      main = paste("Boxplot of", var_name),
      ylab = var_name
    )
    
    legend(
      "topright",
      legend = c(
        paste(names(stats), round(stats, 2), sep = ": "),
        paste("Mean:", round(mean_x, 2)),
        paste("SD:", round(sd_x, 2))
      ),
      bty = "n",
      cex = 0.8
    )
    
    cat("\nBoxplot for:", var_name, "\n")
    print(stats)
    cat("Mean:", round(mean_x, 2), "\n")
    cat("SD:", round(sd_x, 2), "\n")
  }
}


rm(socidem_vars); rm(socidem_df)
```


## paradata


```{r}
#| label: descriptive-stats-paradata-1

para_visualTraps_vars <- str_subset(string = colnames(dat_ques), pattern = "^visualTrap")
para_visualTraps_vars

para_visualTraps_vars_df <- dat_ques %>%
  select(all_of(para_visualTraps_vars))

para_visualTraps_overview <- tibble(
  variable = names(para_visualTraps_vars_df),
  class = purrr::map_chr(para_visualTraps_vars_df, ~ paste(class(.x), collapse = "/")),
  n = nrow(para_visualTraps_vars_df),
  n_missing = purrr::map_int(para_visualTraps_vars_df, ~ sum(is.na(.x))),
  n_unique = purrr::map_int(para_visualTraps_vars_df, ~ dplyr::n_distinct(.x, na.rm = TRUE))
) %>%
  mutate(
    missing_pct = round(100 * n_missing / n, 2)
  )


para_visualTrap_descriptives <- purrr::map_dfr(para_visualTraps_vars, function(v) {
  x <- para_visualTraps_vars_df[[v]]

  if (is.numeric(x)) {
    n_non_missing <- sum(!is.na(x))

    tibble(
      variable = v,
      type = "numeric",
      mean = if (n_non_missing == 0) NA_real_ else mean(x, na.rm = TRUE),
      sd = if (n_non_missing <= 1) NA_real_ else sd(x, na.rm = TRUE),
      median = if (n_non_missing == 0) NA_real_ else median(x, na.rm = TRUE),
      min = if (n_non_missing == 0) NA_real_ else min(x, na.rm = TRUE),
      max = if (n_non_missing == 0) NA_real_ else max(x, na.rm = TRUE),
      top_levels = NA_character_
    )
  } else {
    x_chr <- as.character(x)
    x_chr[x_chr == ""] <- NA_character_
    tab <- sort(table(x_chr, useNA = "no"), decreasing = TRUE)
    k <- min(5, length(tab))

    tibble(
      variable = v,
      type = "categorical_or_text",
      mean = NA_real_,
      sd = NA_real_,
      median = NA_real_,
      min = NA_real_,
      max = NA_real_,
      top_levels = if (k == 0) {
        NA_character_
      } else {
        paste0(names(tab)[seq_len(k)], " (", as.integer(tab[seq_len(k)]), ")", collapse = "; ")
      }
    )
  }
})


para_visualTraps_table <- para_visualTraps_overview %>%
  left_join(para_visualTrap_descriptives, by = "variable")


DT::datatable(data = para_visualTraps_table)
```



```{r}
#| label: descriptive-stats-paradata-2

## factor vars:
factor_vars <- as.character(para_visualTraps_table$variable[para_visualTraps_table$class == "factor"])

for (var_name in factor_vars) {
  cat("\nTable for:", var_name, "\n")
  
  tab <- table(para_visualTraps_vars_df[[var_name]], useNA = "ifany")
  pct <- prop.table(tab) * 100
  
  print(cbind(Frequency = tab, Percentage = round(pct, 1)))
}

## numeric vars:
numeric_vars <- as.character(para_visualTraps_table$variable[para_visualTraps_table$class %in% c("numeric", "integer")])

for (var_name in numeric_vars) {
  if (var_name %in% names(para_visualTraps_vars_df)) {
    x <- para_visualTraps_vars_df[[var_name]]
    x <- x[!is.na(x)]
    
    stats <- boxplot.stats(x)$stats
    names(stats) <- c("Min", "Q1", "Median", "Q3", "Max")
    
    mean_x <- mean(x)
    sd_x <- sd(x)
    
    boxplot(
      x,
      main = paste("Boxplot of", var_name),
      ylab = var_name
    )
    
    legend(
      "topright",
      legend = c(
        paste(names(stats), round(stats, 2), sep = ": "),
        paste("Mean:", round(mean_x, 2)),
        paste("SD:", round(sd_x, 2))
      ),
      bty = "n",
      cex = 0.8
    )
    
    cat("\nBoxplot for:", var_name, "\n")
    print(stats)
    cat("Mean:", round(mean_x, 2), "\n")
    cat("SD:", round(sd_x, 2), "\n")
  }
}


rm(para_visualTraps_vars); rm(para_visualTraps_vars_df)
```

for static paradata:

```{r}
#| label: descriptive-stats-paradata-3

para_statics_vars <- str_subset(string = colnames(dat_ques), pattern = "^para")
para_statics_vars

para_statics_vars_df <- dat_ques %>%
  select(all_of(para_statics_vars))

para_static_overview <- tibble(
  variable = names(para_statics_vars_df),
  class = purrr::map_chr(para_statics_vars_df, ~ paste(class(.x), collapse = "/")),
  n = nrow(para_statics_vars_df),
  n_missing = purrr::map_int(para_statics_vars_df, ~ sum(is.na(.x))),
  n_unique = purrr::map_int(para_statics_vars_df, ~ dplyr::n_distinct(.x, na.rm = TRUE))
) %>%
  mutate(
    missing_pct = round(100 * n_missing / n, 2)
  )


para_static_descriptives <- purrr::map_dfr(para_statics_vars, function(v) {
  x <- para_statics_vars_df[[v]]

  if (is.numeric(x)) {
    n_non_missing <- sum(!is.na(x))

    tibble(
      variable = v,
      type = "numeric",
      mean = if (n_non_missing == 0) NA_real_ else mean(x, na.rm = TRUE),
      sd = if (n_non_missing <= 1) NA_real_ else sd(x, na.rm = TRUE),
      median = if (n_non_missing == 0) NA_real_ else median(x, na.rm = TRUE),
      min = if (n_non_missing == 0) NA_real_ else min(x, na.rm = TRUE),
      max = if (n_non_missing == 0) NA_real_ else max(x, na.rm = TRUE),
      top_levels = NA_character_
    )
  } else {
    x_chr <- as.character(x)
    x_chr[x_chr == ""] <- NA_character_
    tab <- sort(table(x_chr, useNA = "no"), decreasing = TRUE)
    k <- min(5, length(tab))

    tibble(
      variable = v,
      type = "categorical_or_text",
      mean = NA_real_,
      sd = NA_real_,
      median = NA_real_,
      min = NA_real_,
      max = NA_real_,
      top_levels = if (k == 0) {
        NA_character_
      } else {
        paste0(names(tab)[seq_len(k)], " (", as.integer(tab[seq_len(k)]), ")", collapse = "; ")
      }
    )
  }
})


para_static_table <- para_static_overview %>%
  left_join(para_static_descriptives, by = "variable")


DT::datatable(data = para_static_table)


rm(para_statics_vars); rm(para_statics_vars_df)
```

# Search for Bots

```{r}
#| label: visualtrap-rt-right-vs-wrong

visualtrap_rt_vars <- str_subset(
  string = colnames(dat_ques),
  pattern = "^visualTrap_.*_rtMs$"
)

visualtrap_items <- stringr::str_remove(visualtrap_rt_vars, "_rtMs$")
visualtrap_correct_vars <- paste0(visualtrap_items, "_correct")

visualtrap_pairs <- tibble(
  item = visualtrap_items,
  rt_var = visualtrap_rt_vars,
  correct_var = visualtrap_correct_vars
) %>%
  filter(correct_var %in% colnames(dat_ques))

parse_correct_binary <- function(x) {
  x_chr <- tolower(trimws(as.character(x)))

  dplyr::case_when(
    x_chr %in% c("1", "true", "t", "yes", "y", "correct", "right") ~ 1,
    x_chr %in% c("0", "false", "f", "no", "n", "incorrect", "wrong") ~ 0,
    TRUE ~ NA_real_
  )
}

visualtrap_rt_ttest <- purrr::map_dfr(seq_len(nrow(visualtrap_pairs)), function(i) {
  item_name <- visualtrap_pairs$item[[i]]
  rt_name <- visualtrap_pairs$rt_var[[i]]
  correct_name <- visualtrap_pairs$correct_var[[i]]

  correct_bin_all <- parse_correct_binary(dat_ques[[correct_name]])

  n_correct_all <- sum(correct_bin_all == 1, na.rm = TRUE)
  n_wrong_all <- sum(correct_bin_all == 0, na.rm = TRUE)
  n_correct_non_missing <- n_correct_all + n_wrong_all

  rt_all <- suppressWarnings(as.numeric(dat_ques[[rt_name]]))

  item_df <- tibble(
    rt_ms = rt_all,
    correct_bin = correct_bin_all
  ) %>%
    filter(!is.na(rt_ms), !is.na(correct_bin)) %>%
    mutate(group = factor(ifelse(correct_bin == 1, "right", "wrong"), levels = c("wrong", "right")))

  n_right_rt <- sum(item_df$group == "right")
  n_wrong_rt <- sum(item_df$group == "wrong")

  mean_right <- if (n_right_rt > 0) mean(item_df$rt_ms[item_df$group == "right"], na.rm = TRUE) else NA_real_
  mean_wrong <- if (n_wrong_rt > 0) mean(item_df$rt_ms[item_df$group == "wrong"], na.rm = TRUE) else NA_real_
  sd_right <- if (n_right_rt > 1) sd(item_df$rt_ms[item_df$group == "right"], na.rm = TRUE) else NA_real_
  sd_wrong <- if (n_wrong_rt > 1) sd(item_df$rt_ms[item_df$group == "wrong"], na.rm = TRUE) else NA_real_

  t_out <- tryCatch(
    {
      if (n_right_rt >= 2 && n_wrong_rt >= 2) {
        t.test(rt_ms ~ group, data = item_df)
      } else {
        NULL
      }
    },
    error = function(e) NULL
  )

  pooled_sd <- if (!is.na(sd_right) && !is.na(sd_wrong) && (n_right_rt + n_wrong_rt - 2) > 0) {
    sqrt(((n_right_rt - 1) * sd_right^2 + (n_wrong_rt - 1) * sd_wrong^2) / (n_right_rt + n_wrong_rt - 2))
  } else {
    NA_real_
  }

  cohen_d <- if (!is.na(pooled_sd) && pooled_sd > 0) {
    (mean_right - mean_wrong) / pooled_sd
  } else {
    NA_real_
  }

  hedges_g <- if (!is.na(cohen_d) && (n_right_rt + n_wrong_rt) > 3) {
    cohen_d * (1 - (3 / (4 * (n_right_rt + n_wrong_rt) - 9)))
  } else {
    NA_real_
  }

  tibble(
    item = item_name,
    correct_var = correct_name,
    rt_var = rt_name,
    n_correct_non_missing = n_correct_non_missing,
    n_right = n_correct_all,
    n_wrong = n_wrong_all,
    prop_right = if (n_correct_non_missing > 0) n_correct_all / n_correct_non_missing else NA_real_,
    prop_wrong = if (n_correct_non_missing > 0) n_wrong_all / n_correct_non_missing else NA_real_,
    n_rt_ttest = nrow(item_df),
    n_right_rt = n_right_rt,
    n_wrong_rt = n_wrong_rt,
    mean_rt_right = mean_right,
    sd_rt_right = sd_right,
    mean_rt_wrong = mean_wrong,
    sd_rt_wrong = sd_wrong,
    mean_diff_right_minus_wrong = mean_right - mean_wrong,
    t_statistic = if (!is.null(t_out)) unname(t_out$statistic) else NA_real_,
    df = if (!is.null(t_out)) unname(t_out$parameter) else NA_real_,
    p_value = if (!is.null(t_out)) t_out$p.value else NA_real_,
    conf_low = if (!is.null(t_out)) t_out$conf.int[[1]] else NA_real_,
    conf_high = if (!is.null(t_out)) t_out$conf.int[[2]] else NA_real_,
    cohen_d = cohen_d,
    hedges_g = hedges_g
  )
}) %>%
  mutate(
    across(c(prop_right, prop_wrong), ~ round(.x, 3)),
    across(c(mean_rt_right, sd_rt_right, mean_rt_wrong, sd_rt_wrong, mean_diff_right_minus_wrong), ~ round(.x, 2)),
    across(c(t_statistic, df, cohen_d, hedges_g), ~ round(.x, 3)),
    p_value = round(p_value, 4),
    across(c(conf_low, conf_high), ~ round(.x, 2))
  )

visualtrap_rt_ttest

DT::datatable(
  visualtrap_rt_ttest,
  caption = "Visual-trap item analysis: right vs wrong reaction times (Welch t-test + effect sizes)",
  options = list(pageLength = 10, scrollX = TRUE)
)
```


```{r}
dat_para_events$clipboard_text[!is.na(dat_para_events$clipboard_text)]
dat_para_events$clipboard_length[!is.na(dat_para_events$clipboard_length)]
dat_para_events$PROLIFIC_PID[!is.na(dat_para_events$clipboard_length)]

```






```{r}
#| label: bot-indicators-and-clustering
#| warning: false
#| message: false

parse_bool <- function(x) {
  x_chr <- tolower(trimws(as.character(x)))
  dplyr::case_when(
    x_chr %in% c("1", "true", "t", "yes", "y") ~ 1,
    x_chr %in% c("0", "false", "f", "no", "n") ~ 0,
    TRUE ~ NA_real_
  )
}

safe_num <- function(x) suppressWarnings(as.numeric(as.character(x)))

vt_correct_vars <- str_subset(colnames(dat_ques), "^visualTrap_.*_correct$")
vt_rt_vars <- str_subset(colnames(dat_ques), "^visualTrap_.*_rtMs$")

vt_correct_mat <- dat_ques %>%
  select(all_of(vt_correct_vars)) %>%
  mutate(across(everything(), parse_bool))

vt_rt_mat <- dat_ques %>%
  select(all_of(vt_rt_vars)) %>%
  mutate(across(everything(), safe_num))

vt_wrong_count <- vt_correct_mat %>%
  mutate(across(everything(), ~ ifelse(.x == 0, 1, 0))) %>%
  transmute(vt_wrong_count = rowSums(across(everything()), na.rm = TRUE)) %>%
  pull(vt_wrong_count)

vt_answered_count <- vt_correct_mat %>%
  transmute(vt_answered_count = rowSums(!is.na(across(everything())))) %>%
  pull(vt_answered_count)

vt_correct_count <- vt_correct_mat %>%
  transmute(vt_correct_count = rowSums(across(everything()), na.rm = TRUE)) %>%
  pull(vt_correct_count)

vt_prop_correct <- ifelse(vt_answered_count > 0, vt_correct_count / vt_answered_count, NA_real_)

vt_mean_rt <- vt_rt_mat %>%
  transmute(vt_mean_rt = rowMeans(across(everything()), na.rm = TRUE)) %>%
  pull(vt_mean_rt)

vt_mean_rt_p05 <- quantile(vt_mean_rt, probs = 0.05, na.rm = TRUE)
vt_mean_rt_p95 <- quantile(vt_mean_rt, probs = 0.95, na.rm = TRUE)


vt_fast_count <- vt_rt_mat %>%
  mutate(across(everything(), ~ ifelse(!is.na(.x) & .x < vt_mean_rt_p05, 1, 0))) %>%
  transmute(vt_fast_count = rowSums(across(everything()), na.rm = TRUE)) %>%
  pull(vt_fast_count)

vt_slow_count <- vt_rt_mat %>%
  mutate(across(everything(), ~ ifelse(!is.na(.x) & .x > vt_mean_rt_p95, 1, 0))) %>%
  transmute(vt_slow_count = rowSums(across(everything()), na.rm = TRUE)) %>%
  pull(vt_slow_count)

webdriver_flag <- parse_bool(dat_ques$para_static_webdriver)
honeypot_flag <- parse_bool(dat_ques$hp_exclusionCriteria)

duration_seconds <- safe_num(dat_ques$duration_seconds)
duration_p05 <- quantile(duration_seconds, probs = 0.05, na.rm = TRUE)
duration_p95 <- quantile(duration_seconds, probs = 0.95, na.rm = TRUE)

very_fast_duration_flag <- ifelse(!is.na(duration_seconds) & duration_seconds < duration_p05, 1, 0)
very_slow_duration_flag <- ifelse(!is.na(duration_seconds) & duration_seconds > duration_p95, 1, 0)

events_by_pid <- dat_para_events %>%
  mutate(
    key_interval_ms = safe_num(key_interval_ms),
    clipboard_length = safe_num(clipboard_length)
  ) %>%
  group_by(PROLIFIC_PID) %>%
  summarise(
    n_events = n(),
    n_copy = sum(event_type == "copy", na.rm = TRUE),
    n_paste = sum(event_type == "paste", na.rm = TRUE),
    n_mouse = sum(event_type == "mousemove", na.rm = TRUE),
    n_scroll = sum(event_type == "scroll", na.rm = TRUE),
    n_key_intervals = sum(!is.na(key_interval_ms)),
    key_interval_mean = mean(key_interval_ms, na.rm = TRUE),
    key_interval_sd = sd(key_interval_ms, na.rm = TRUE),
    clipboard_total_chars = sum(clipboard_length, na.rm = TRUE),
    .groups = "drop"
  )

defocus_by_pid <- dat_para_defocus %>%
  mutate(durationblur = safe_num(durationblur)) %>%
  group_by(PROLIFIC_PID) %>%
  summarise(
    defocus_n = n(),
    defocus_total_ms = sum(durationblur, na.rm = TRUE),
    defocus_mean_ms = mean(durationblur, na.rm = TRUE),
    .groups = "drop"
  )

bot_df <- dat_ques %>%
  select(PROLIFIC_PID, result_id, duration_seconds = duration_seconds) %>%
  mutate(
    duration_seconds = safe_num(duration_seconds),
    vt_wrong_count = vt_wrong_count,
    vt_answered_count = vt_answered_count,
    vt_correct_count = vt_correct_count,
    vt_prop_correct = vt_prop_correct,
    vt_mean_rt = vt_mean_rt,
    vt_fast_count = vt_fast_count,
    vt_slow_count = vt_slow_count,
    webdriver_flag = webdriver_flag,
    honeypot_flag = honeypot_flag,
    very_fast_duration_flag = very_fast_duration_flag,
    very_slow_duration_flag = very_slow_duration_flag
  ) %>%
  left_join(events_by_pid, by = "PROLIFIC_PID") %>%
  left_join(defocus_by_pid, by = "PROLIFIC_PID") %>%
  mutate(
    across(c(n_events, n_copy, n_paste, n_mouse, n_scroll, n_key_intervals,
             clipboard_total_chars, defocus_n, defocus_total_ms, defocus_mean_ms),
           ~ replace_na(.x, 0)),
    key_interval_mean = ifelse(is.nan(key_interval_mean), NA_real_, key_interval_mean),
    key_interval_sd = ifelse(is.nan(key_interval_sd), NA_real_, key_interval_sd)
  )

vt_rt_p10 <- quantile(bot_df$vt_mean_rt, 0.10, na.rm = TRUE)
vt_rt_p90 <- quantile(bot_df$vt_mean_rt, 0.90, na.rm = TRUE)
paste_p95 <- quantile(bot_df$n_paste, 0.95, na.rm = TRUE)

bot_df <- bot_df %>%
  mutate(
    risk_webdriver = ifelse(webdriver_flag == 1, 1, 0),
    risk_honeypot = ifelse(honeypot_flag == 1, 1, 0),
    risk_many_wrong_visual = ifelse(vt_wrong_count >= 3, 1, 0),
    risk_low_visual_accuracy = ifelse(!is.na(vt_prop_correct) & vt_prop_correct < 0.50, 1, 0),
    risk_very_fast_visual_rt = ifelse(!is.na(vt_mean_rt) & vt_mean_rt < vt_rt_p10, 1, 0),
    risk_very_slow_visual_rt = ifelse(!is.na(vt_mean_rt) & vt_mean_rt > vt_rt_p90, 1, 0),
    risk_very_fast_duration = ifelse(very_fast_duration_flag == 1, 1, 0),
    risk_high_paste = ifelse(!is.na(n_paste) & n_paste > paste_p95, 1, 0),
    risk_low_mouse_events = ifelse(!is.na(n_mouse) & n_mouse <= 5, 1, 0),
    risk_high_defocus = ifelse(!is.na(defocus_n) & defocus_n >= 5, 1, 0),
    bot_risk_score = risk_webdriver + risk_honeypot + risk_many_wrong_visual +
      risk_low_visual_accuracy + risk_very_fast_visual_rt + risk_very_slow_visual_rt +
      risk_very_fast_duration + risk_high_paste + risk_low_mouse_events + risk_high_defocus
  )

risk_indicator_table <- bot_df %>%
  summarise(
    n = n(),
    across(starts_with("risk_"), ~ mean(.x, na.rm = TRUE), .names = "{.col}_prop")
  ) %>%
  pivot_longer(cols = starts_with("risk_"), names_to = "indicator", values_to = "proportion_flagged") %>%
  mutate(proportion_flagged = round(proportion_flagged, 3)) %>%
  arrange(desc(proportion_flagged))

risk_indicator_table

bot_risk_score_dist <- bot_df %>%
  count(bot_risk_score, name = "n") %>%
  mutate(prop = round(n / sum(n), 3)) %>%
  arrange(bot_risk_score)

bot_risk_score_dist

cluster_features <- bot_df %>%
  select(
    vt_wrong_count, vt_prop_correct, vt_mean_rt, vt_fast_count, vt_slow_count,
    duration_seconds, n_events, n_paste, n_mouse, n_scroll,
    n_key_intervals, key_interval_mean, key_interval_sd,
    defocus_n, defocus_total_ms, bot_risk_score
  ) %>%
  mutate(across(everything(), ~ ifelse(is.infinite(.x), NA_real_, .x)))

for (j in seq_len(ncol(cluster_features))) {
  med_j <- median(cluster_features[[j]], na.rm = TRUE)
  cluster_features[[j]][is.na(cluster_features[[j]])] <- med_j
}


# remove only variables that contain at least one missing value
vars_with_na <- names(cluster_features_scaled)[colSums(is.na(cluster_features_scaled)) > 0]
vars_with_na

cluster_features_scaled_noNAvars <- cluster_features_scaled %>%
  dplyr::select(where(is.numeric)) %>%
  dplyr::select(-dplyr::all_of(vars_with_na))

# clustering task
task <- mlr3cluster::TaskClust$new(
  id = "bot_profile",
  backend = cluster_features_scaled_noNAvars
)

# evaluate candidate k values
k_grid <- 2:6
dmat <- dist(cluster_features_scaled_noNAvars)

k_eval <- purrr::map_dfr(k_grid, function(k) {
  learner_k <- mlr3::lrn("clust.kmeans", centers = k, nstart = 30)
  learner_k$train(task)
  pred_k <- learner_k$predict(task)
  cl <- as.integer(pred_k$partition)

  sil <- tryCatch({
    mean(cluster::silhouette(cl, dmat)[, "sil_width"])
  }, error = function(e) NA_real_)

  tibble::tibble(
    k = k,
    avg_silhouette = sil
  )
})

k_eval

# choose best k
best_k <- k_eval %>%
  dplyr::filter(!is.na(avg_silhouette)) %>%
  dplyr::arrange(dplyr::desc(avg_silhouette)) %>%
  dplyr::slice(1) %>%
  dplyr::pull(k)

if (length(best_k) == 0 || is.na(best_k)) best_k <- 3

# final model
learner_final <- mlr3::lrn("clust.kmeans", centers = best_k, nstart = 50)
learner_final$train(task)
pred_final <- learner_final$predict(task)

# attach cluster labels back to original data
bot_df <- bot_df %>%
  dplyr::mutate(bot_cluster = factor(pred_final$partition))

# cluster summaries
cluster_profile <- bot_df %>%
  dplyr::group_by(bot_cluster) %>%
  dplyr::summarise(
    n = dplyr::n(),
    prop_sample = round(n / nrow(bot_df), 3),
    mean_bot_risk_score = round(mean(bot_risk_score, na.rm = TRUE), 3),
    mean_vt_prop_correct = round(mean(vt_prop_correct, na.rm = TRUE), 3),
    mean_vt_mean_rt = round(mean(vt_mean_rt, na.rm = TRUE), 1),
    mean_duration_seconds = round(mean(duration_seconds, na.rm = TRUE), 1),
    mean_n_paste = round(mean(n_paste, na.rm = TRUE), 2),
    mean_n_mouse = round(mean(n_mouse, na.rm = TRUE), 2),
    mean_defocus_n = round(mean(defocus_n, na.rm = TRUE), 2),
    prop_webdriver = round(mean(webdriver_flag == 1, na.rm = TRUE), 3),
    prop_honeypot = round(mean(honeypot_flag == 1, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  dplyr::arrange(dplyr::desc(mean_bot_risk_score))

cluster_profile

# identify highest-risk cluster
high_risk_cluster <- cluster_profile %>%
  dplyr::slice(1) %>%
  dplyr::pull(bot_cluster)

suspected_bot_table <- bot_df %>%
  dplyr::mutate(suspected_bot_cluster = ifelse(bot_cluster == high_risk_cluster, 1, 0)) %>%
  dplyr::select(
    PROLIFIC_PID, result_id, bot_cluster, suspected_bot_cluster, bot_risk_score,
    vt_wrong_count, vt_prop_correct, vt_mean_rt, duration_seconds,
    webdriver_flag, honeypot_flag, n_paste, n_mouse, defocus_n
  ) %>%
  dplyr::arrange(dplyr::desc(suspected_bot_cluster), dplyr::desc(bot_risk_score), vt_prop_correct)

DT::datatable(
  suspected_bot_table,
  caption = paste0("Bot-risk profiling table (best k = ", best_k, ")"),
  options = list(pageLength = 12, scrollX = TRUE)
)
```


