---
title: "Data preperation"
author: "Julius Fenn"
format:
  html:
    toc: true
    toc-depth: 3
    html-math-method: katex
    number-sections: true
---

# Notes

This script creates a simplified participant-level dataset using the raw
`study.rds` file and produces:

- A clean wide table with core variables + associations (wide) + keyStats
- Paradata events in long format (clipboard, mouse, scroll, keyIntervals)
- Defocus events in long format

# global variables

```{r}
#| label: global-variables

data_root <- "../01_dataPreperation/outputs"
output_root <- "outputs"
```

# functions

```{r}
# not needed
```

# load packages

```{r}
#| label: load-packages
#| warning: false

require(pacman)
p_load(
  "tidyverse",
  "psych",
  "afex",
  "writexl",
  "readxl",
  "DT"
)
```

# load processed data

```{r}
#| label: load-data
#| warning: false

setwd(data_root)
setwd("questionnaire")

dat_ques <- readRDS(file = "clean_complete_wide.rds")

setwd("../paradata")

dat_para_defocus <- readRDS(file = "paradata_defocus_long.rds")
dat_para_events <- readRDS(file = "dat_para_events.rds")
```


# Descriptive Statistics

## sample


```{r}
#| label: descriptive-stats-sample-1

socidem_vars <- str_subset(string = colnames(dat_ques), pattern = "^socidem")
socidem_vars

socidem_df <- dat_ques %>%
  select(all_of(socidem_vars))

socidem_overview <- tibble(
  variable = socidem_vars,
  class = purrr::map_chr(socidem_df, ~ paste(class(.x), collapse = "/")),
  n = nrow(socidem_df),
  n_missing = purrr::map_int(socidem_df, ~ sum(is.na(.x))),
  missing_pct = round(100 * n_missing / n, 2),
  n_unique = purrr::map_int(socidem_df, ~ dplyr::n_distinct(.x, na.rm = TRUE))
)

socidem_descriptives <- purrr::map_dfr(socidem_vars, function(v) {
  x <- socidem_df[[v]]

  if (is.numeric(x)) {
    n_non_missing <- sum(!is.na(x))

    tibble(
      variable = v,
      type = "numeric",
      mean = if (n_non_missing == 0) NA_real_ else mean(x, na.rm = TRUE),
      sd = if (n_non_missing <= 1) NA_real_ else sd(x, na.rm = TRUE),
      median = if (n_non_missing == 0) NA_real_ else median(x, na.rm = TRUE),
      min = if (n_non_missing == 0) NA_real_ else min(x, na.rm = TRUE),
      max = if (n_non_missing == 0) NA_real_ else max(x, na.rm = TRUE),
      top_levels = NA_character_
    )
  } else {
    x_chr <- as.character(x)
    x_chr[x_chr == ""] <- NA_character_
    tab <- sort(table(x_chr, useNA = "no"), decreasing = TRUE)
    k <- min(5, length(tab))

    tibble(
      variable = v,
      type = "categorical_or_text",
      mean = NA_real_,
      sd = NA_real_,
      median = NA_real_,
      min = NA_real_,
      max = NA_real_,
      top_levels = if (k == 0) {
        NA_character_
      } else {
        paste0(names(tab)[seq_len(k)], " (", as.integer(tab[seq_len(k)]), ")", collapse = "; ")
      }
    )
  }
})

socidem_table <- socidem_overview %>%
  left_join(socidem_descriptives, by = "variable")


DT::datatable(data = socidem_table)
```


```{r}
#| label: descriptive-stats-sample-2

## factor vars:
factor_vars <- as.character(socidem_table$variable[socidem_table$class == "factor"])

for (var_name in factor_vars) {
  cat("\nTable for:", var_name, "\n")
  
  tab <- table(socidem_df[[var_name]], useNA = "ifany")
  pct <- prop.table(tab) * 100
  
  print(cbind(Frequency = tab, Percentage = round(pct, 1)))
}

## numeric vars:
numeric_vars <- as.character(socidem_table$variable[socidem_table$class == "numeric"])

for (var_name in numeric_vars) {
  if (var_name %in% names(socidem_df)) {
    x <- socidem_df[[var_name]]
    x <- x[!is.na(x)]
    
    stats <- boxplot.stats(x)$stats
    names(stats) <- c("Min", "Q1", "Median", "Q3", "Max")
    
    mean_x <- mean(x)
    sd_x <- sd(x)
    
    boxplot(
      x,
      main = paste("Boxplot of", var_name),
      ylab = var_name
    )
    
    legend(
      "topright",
      legend = c(
        paste(names(stats), round(stats, 2), sep = ": "),
        paste("Mean:", round(mean_x, 2)),
        paste("SD:", round(sd_x, 2))
      ),
      bty = "n",
      cex = 0.8
    )
    
    cat("\nBoxplot for:", var_name, "\n")
    print(stats)
    cat("Mean:", round(mean_x, 2), "\n")
    cat("SD:", round(sd_x, 2), "\n")
  }
}


rm(socidem_vars); rm(socidem_df)
```


## paradata


```{r}
#| label: descriptive-stats-paradata-1

para_visualTraps_vars <- str_subset(string = colnames(dat_ques), pattern = "^visualTrap")
para_visualTraps_vars

para_visualTraps_vars_df <- dat_ques %>%
  select(all_of(para_visualTraps_vars))

para_visualTraps_overview <- tibble(
  variable = names(para_visualTraps_vars_df),
  class = purrr::map_chr(para_visualTraps_vars_df, ~ paste(class(.x), collapse = "/")),
  n = nrow(para_visualTraps_vars_df),
  n_missing = purrr::map_int(para_visualTraps_vars_df, ~ sum(is.na(.x))),
  n_unique = purrr::map_int(para_visualTraps_vars_df, ~ dplyr::n_distinct(.x, na.rm = TRUE))
) %>%
  mutate(
    missing_pct = round(100 * n_missing / n, 2)
  )


para_visualTrap_descriptives <- purrr::map_dfr(para_visualTraps_vars, function(v) {
  x <- para_visualTraps_vars_df[[v]]

  if (is.numeric(x)) {
    n_non_missing <- sum(!is.na(x))

    tibble(
      variable = v,
      type = "numeric",
      mean = if (n_non_missing == 0) NA_real_ else mean(x, na.rm = TRUE),
      sd = if (n_non_missing <= 1) NA_real_ else sd(x, na.rm = TRUE),
      median = if (n_non_missing == 0) NA_real_ else median(x, na.rm = TRUE),
      min = if (n_non_missing == 0) NA_real_ else min(x, na.rm = TRUE),
      max = if (n_non_missing == 0) NA_real_ else max(x, na.rm = TRUE),
      top_levels = NA_character_
    )
  } else {
    x_chr <- as.character(x)
    x_chr[x_chr == ""] <- NA_character_
    tab <- sort(table(x_chr, useNA = "no"), decreasing = TRUE)
    k <- min(5, length(tab))

    tibble(
      variable = v,
      type = "categorical_or_text",
      mean = NA_real_,
      sd = NA_real_,
      median = NA_real_,
      min = NA_real_,
      max = NA_real_,
      top_levels = if (k == 0) {
        NA_character_
      } else {
        paste0(names(tab)[seq_len(k)], " (", as.integer(tab[seq_len(k)]), ")", collapse = "; ")
      }
    )
  }
})


para_visualTraps_table <- para_visualTraps_overview %>%
  left_join(para_visualTrap_descriptives, by = "variable")


DT::datatable(data = para_visualTraps_table)
```



```{r}
#| label: descriptive-stats-paradata-2

## factor vars:
factor_vars <- as.character(para_visualTraps_table$variable[para_visualTraps_table$class == "factor"])

for (var_name in factor_vars) {
  cat("\nTable for:", var_name, "\n")
  
  tab <- table(para_visualTraps_vars_df[[var_name]], useNA = "ifany")
  pct <- prop.table(tab) * 100
  
  print(cbind(Frequency = tab, Percentage = round(pct, 1)))
}

## numeric vars:
numeric_vars <- as.character(para_visualTraps_table$variable[para_visualTraps_table$class %in% c("numeric", "integer")])

for (var_name in numeric_vars) {
  if (var_name %in% names(para_visualTraps_vars_df)) {
    x <- para_visualTraps_vars_df[[var_name]]
    x <- x[!is.na(x)]
    
    stats <- boxplot.stats(x)$stats
    names(stats) <- c("Min", "Q1", "Median", "Q3", "Max")
    
    mean_x <- mean(x)
    sd_x <- sd(x)
    
    boxplot(
      x,
      main = paste("Boxplot of", var_name),
      ylab = var_name
    )
    
    legend(
      "topright",
      legend = c(
        paste(names(stats), round(stats, 2), sep = ": "),
        paste("Mean:", round(mean_x, 2)),
        paste("SD:", round(sd_x, 2))
      ),
      bty = "n",
      cex = 0.8
    )
    
    cat("\nBoxplot for:", var_name, "\n")
    print(stats)
    cat("Mean:", round(mean_x, 2), "\n")
    cat("SD:", round(sd_x, 2), "\n")
  }
}


rm(para_visualTraps_vars); rm(para_visualTraps_vars_df)
```

for static paradata:

```{r}
#| label: descriptive-stats-paradata-3

para_statics_vars <- str_subset(string = colnames(dat_ques), pattern = "^para")
para_statics_vars

para_statics_vars_df <- dat_ques %>%
  select(all_of(para_statics_vars))

para_static_overview <- tibble(
  variable = names(para_statics_vars_df),
  class = purrr::map_chr(para_statics_vars_df, ~ paste(class(.x), collapse = "/")),
  n = nrow(para_statics_vars_df),
  n_missing = purrr::map_int(para_statics_vars_df, ~ sum(is.na(.x))),
  n_unique = purrr::map_int(para_statics_vars_df, ~ dplyr::n_distinct(.x, na.rm = TRUE))
) %>%
  mutate(
    missing_pct = round(100 * n_missing / n, 2)
  )


para_static_descriptives <- purrr::map_dfr(para_statics_vars, function(v) {
  x <- para_statics_vars_df[[v]]

  if (is.numeric(x)) {
    n_non_missing <- sum(!is.na(x))

    tibble(
      variable = v,
      type = "numeric",
      mean = if (n_non_missing == 0) NA_real_ else mean(x, na.rm = TRUE),
      sd = if (n_non_missing <= 1) NA_real_ else sd(x, na.rm = TRUE),
      median = if (n_non_missing == 0) NA_real_ else median(x, na.rm = TRUE),
      min = if (n_non_missing == 0) NA_real_ else min(x, na.rm = TRUE),
      max = if (n_non_missing == 0) NA_real_ else max(x, na.rm = TRUE),
      top_levels = NA_character_
    )
  } else {
    x_chr <- as.character(x)
    x_chr[x_chr == ""] <- NA_character_
    tab <- sort(table(x_chr, useNA = "no"), decreasing = TRUE)
    k <- min(5, length(tab))

    tibble(
      variable = v,
      type = "categorical_or_text",
      mean = NA_real_,
      sd = NA_real_,
      median = NA_real_,
      min = NA_real_,
      max = NA_real_,
      top_levels = if (k == 0) {
        NA_character_
      } else {
        paste0(names(tab)[seq_len(k)], " (", as.integer(tab[seq_len(k)]), ")", collapse = "; ")
      }
    )
  }
})


para_static_table <- para_static_overview %>%
  left_join(para_static_descriptives, by = "variable")


DT::datatable(data = para_static_table)


rm(para_statics_vars); rm(para_statics_vars_df)
```

# Search for Bots

```{r}
#| label: visualtrap-rt-right-vs-wrong

visualtrap_rt_vars <- str_subset(
  string = colnames(dat_ques),
  pattern = "^visualTrap_.*_rtMs$"
)

visualtrap_items <- stringr::str_remove(visualtrap_rt_vars, "_rtMs$")
visualtrap_correct_vars <- paste0(visualtrap_items, "_correct")

visualtrap_pairs <- tibble(
  item = visualtrap_items,
  rt_var = visualtrap_rt_vars,
  correct_var = visualtrap_correct_vars
) %>%
  filter(correct_var %in% colnames(dat_ques))

parse_correct_binary <- function(x) {
  x_chr <- tolower(trimws(as.character(x)))

  dplyr::case_when(
    x_chr %in% c("1", "true", "t", "yes", "y", "correct", "right") ~ 1,
    x_chr %in% c("0", "false", "f", "no", "n", "incorrect", "wrong") ~ 0,
    TRUE ~ NA_real_
  )
}

visualtrap_rt_ttest <- purrr::map_dfr(seq_len(nrow(visualtrap_pairs)), function(i) {
  item_name <- visualtrap_pairs$item[[i]]
  rt_name <- visualtrap_pairs$rt_var[[i]]
  correct_name <- visualtrap_pairs$correct_var[[i]]

  correct_bin_all <- parse_correct_binary(dat_ques[[correct_name]])

  n_correct_all <- sum(correct_bin_all == 1, na.rm = TRUE)
  n_wrong_all <- sum(correct_bin_all == 0, na.rm = TRUE)
  n_correct_non_missing <- n_correct_all + n_wrong_all

  rt_all <- suppressWarnings(as.numeric(dat_ques[[rt_name]]))

  item_df <- tibble(
    rt_ms = rt_all,
    correct_bin = correct_bin_all
  ) %>%
    filter(!is.na(rt_ms), !is.na(correct_bin)) %>%
    mutate(group = factor(ifelse(correct_bin == 1, "right", "wrong"), levels = c("wrong", "right")))

  n_right_rt <- sum(item_df$group == "right")
  n_wrong_rt <- sum(item_df$group == "wrong")

  mean_right <- if (n_right_rt > 0) mean(item_df$rt_ms[item_df$group == "right"], na.rm = TRUE) else NA_real_
  mean_wrong <- if (n_wrong_rt > 0) mean(item_df$rt_ms[item_df$group == "wrong"], na.rm = TRUE) else NA_real_
  sd_right <- if (n_right_rt > 1) sd(item_df$rt_ms[item_df$group == "right"], na.rm = TRUE) else NA_real_
  sd_wrong <- if (n_wrong_rt > 1) sd(item_df$rt_ms[item_df$group == "wrong"], na.rm = TRUE) else NA_real_

  t_out <- tryCatch(
    {
      if (n_right_rt >= 2 && n_wrong_rt >= 2) {
        t.test(rt_ms ~ group, data = item_df)
      } else {
        NULL
      }
    },
    error = function(e) NULL
  )

  pooled_sd <- if (!is.na(sd_right) && !is.na(sd_wrong) && (n_right_rt + n_wrong_rt - 2) > 0) {
    sqrt(((n_right_rt - 1) * sd_right^2 + (n_wrong_rt - 1) * sd_wrong^2) / (n_right_rt + n_wrong_rt - 2))
  } else {
    NA_real_
  }

  cohen_d <- if (!is.na(pooled_sd) && pooled_sd > 0) {
    (mean_right - mean_wrong) / pooled_sd
  } else {
    NA_real_
  }

  hedges_g <- if (!is.na(cohen_d) && (n_right_rt + n_wrong_rt) > 3) {
    cohen_d * (1 - (3 / (4 * (n_right_rt + n_wrong_rt) - 9)))
  } else {
    NA_real_
  }

  tibble(
    item = item_name,
    correct_var = correct_name,
    rt_var = rt_name,
    n_correct_non_missing = n_correct_non_missing,
    n_right = n_correct_all,
    n_wrong = n_wrong_all,
    prop_right = if (n_correct_non_missing > 0) n_correct_all / n_correct_non_missing else NA_real_,
    prop_wrong = if (n_correct_non_missing > 0) n_wrong_all / n_correct_non_missing else NA_real_,
    n_rt_ttest = nrow(item_df),
    n_right_rt = n_right_rt,
    n_wrong_rt = n_wrong_rt,
    mean_rt_right = mean_right,
    sd_rt_right = sd_right,
    mean_rt_wrong = mean_wrong,
    sd_rt_wrong = sd_wrong,
    mean_diff_right_minus_wrong = mean_right - mean_wrong,
    t_statistic = if (!is.null(t_out)) unname(t_out$statistic) else NA_real_,
    df = if (!is.null(t_out)) unname(t_out$parameter) else NA_real_,
    p_value = if (!is.null(t_out)) t_out$p.value else NA_real_,
    conf_low = if (!is.null(t_out)) t_out$conf.int[[1]] else NA_real_,
    conf_high = if (!is.null(t_out)) t_out$conf.int[[2]] else NA_real_,
    cohen_d = cohen_d,
    hedges_g = hedges_g
  )
}) %>%
  mutate(
    across(c(prop_right, prop_wrong), ~ round(.x, 3)),
    across(c(mean_rt_right, sd_rt_right, mean_rt_wrong, sd_rt_wrong, mean_diff_right_minus_wrong), ~ round(.x, 2)),
    across(c(t_statistic, df, cohen_d, hedges_g), ~ round(.x, 3)),
    p_value = round(p_value, 4),
    across(c(conf_low, conf_high), ~ round(.x, 2))
  )

visualtrap_rt_ttest

DT::datatable(
  visualtrap_rt_ttest,
  caption = "Visual-trap item analysis: right vs wrong reaction times (Welch t-test + effect sizes)",
  options = list(pageLength = 10, scrollX = TRUE)
)
```





```{r}
XXX
```



