---
title: "Data analysis all studies"
author: "Richard Bergs, Julius Fenn"
format:
  html:
    toc: true
    toc-depth: 3
    html-math-method: katex
    number-sections: true
---


# Notes


# global variables

Define your global variables (e.g., to reduce run time):

```{r}
#| label: global variables


```


# functions

```{r}
########################################
# xx
########################################
### argss:
# xx


```


# load packages, functions, and data

## packages

```{r}
#| label: load packages
#| warning: false

### load packages (also packages loaded, which are not needed for data preperation)
require(pacman)
p_load('tidyverse', 'jsonlite',
       'stargazer',  'DT', 'psych',
       'writexl', 'associatoR',
       'ggstatsplot', 
       'afex', 'emmeans', 'effectsize',
       'moments', 'lavaan', 'semPlot')
# devtools::install_github("samuelae/associatoR")
```

## load external functions

```{r}
#| echo: true
#| warning: false


### load functions
# print(getwd())
setwd("functions")
for(i in 1:length(dir())){
  # print(dir()[i])
  source(dir()[i], encoding = "utf-8")
}
```


## load prepared data

```{r}
setwd("outputs/questionnaire")

ques_study <- readRDS(file = paste0("ques_study", ".rds"))
```

```{r}
setwd("outputs/associations_translated")

ass_study <- readxl::read_excel(path = paste0("ass_study_translated_sentiment", ".xlsx"), sheet = 1)
str(ass_study)
dim(ass_study)
```





# check and compute mean values of valence, scales

## mean values of valence

```{r}
library(dplyr)

compute_mean_sd_valence <- function(data,
                                 id_col = "participant_id",
                                 condition_col = "study_condition",
                                 valence_col = "valence") {
  
  data %>%
    mutate(
      valence = as.numeric(.data[[valence_col]])
    ) %>%
    filter(!is.na(valence)) %>%
    group_by(
      participant_id = .data[[id_col]],
      study_condition = .data[[condition_col]]
    ) %>%
    summarise(
      mean_valence = mean(valence, na.rm = TRUE),
      sd_valence   = sd(valence, na.rm = TRUE),
      n_ratings    = n(),
      .groups = "drop"
    )
}

mean_sd_valence_per_person <- compute_mean_sd_valence(ass_study)

str(mean_sd_valence_per_person)
```

append variable to ques_study:

```{r}
ques_study <- ques_study %>%
  left_join(
    mean_sd_valence_per_person %>%
      select(participant_id, mean_valence, sd_valence),
    by = c("PROLIFIC_PID" = "participant_id")
  )
```


## mean values of scales

### EFA overall


Remark: factors scores were computed using "tenBerge", which preserves correlation structure, but not as widely used in applied psych research (often "regression" used).


```{r}
#### Overall EFA
regExOverall <- "^AIAS|^PTTA"

psych::cor.plot(r = cor(ques_study[, str_detect(string = colnames(ques_study),
                                                   pattern = regExOverall)]
                        , use = "pairwise.complete.obs"),
                upper = FALSE, xlas = 2, main = "Overall")

#> parallel analysis
tmp_parallelAnalysis <- dimensionalityTest(label = "Overall",
                             regEx = regExOverall, dataset = ques_study)

#> EFA
tmp_EFA_overall <- explorativeFactorAnalysis(label = "Overall",
                                 regEx = regExOverall,
                                 dataset = ques_study, nfac = tmp_parallelAnalysis$nfact) # manually set
tmp_EFA_overall[[1]]
tmp_EFA_overall[[1]]$loadings
```

### Descriptives, correlation plot, EFA, CFA for AIAS

```{r}
regEx <- "^AIAS"
nameVariable <- "AIAS"
  
tmp_numVars <- sum(str_detect(string = colnames(ques_study), pattern = regEx))
cat("\n>>>",  tmp_numVars, "items )<<<\n\n")
tmp_dat <- na.omit(ques_study[,str_detect(string = colnames(ques_study), pattern = regEx)])
tmp <- CFAstats(dataset = tmp_dat, regularExp = regEx, labelLatent = str_remove(string = nameVariable, pattern = "mean_"), 
                  showPlots = TRUE, 
                  computeEFA = TRUE, 
                  computeCFA = TRUE, 
                  computeCFAMplus = FALSE)
```




### Descriptives, correlation plot, EFA, CFA for PTTA

```{r}
regEx <- "^PTTA"
nameVariable <- "PTTA"
  
tmp_numVars <- sum(str_detect(string = colnames(ques_study), pattern = regEx))
cat("\n>>>",  tmp_numVars, "items )<<<\n\n")
tmp_dat <- na.omit(ques_study[,str_detect(string = colnames(ques_study), pattern = regEx)])
tmp <- CFAstats(dataset = tmp_dat, regularExp = regEx, labelLatent = str_remove(string = nameVariable, pattern = "mean_"), 
                  showPlots = TRUE, 
                  computeEFA = TRUE, 
                  computeCFA = TRUE, 
                  computeCFAMplus = FALSE)
```

### compute mean scores

```{r}
ques_study$mean_AIAS <- rowMeans(x = ques_study[, str_subset(string = colnames(ques_study), pattern = "^AIAS")], na.rm = FALSE) # no missing
ques_study$mean_PTTA <- rowMeans(x = ques_study[, str_subset(string = colnames(ques_study), pattern = "^PTTA")], na.rm = FALSE) # no missing
```



# Descriptive Statistics

**sample**

```{r}
table(ques_study$sociodemo_gender)
table(ques_study$sociodemo_gender) / nrow(ques_study)

summary(ques_study$sociodemo_age)
hist(ques_study$sociodemo_age)
```

```{r}
# Ich finde es wichtig, dass ein Mensch in dem beschriebenen Kontext Aufsicht über KI hat.
barplot(table(ques_study$`PUR-1`))
# Ich halte den Einsatz von KI in diesem Kontext für riskant.
barplot(table(ques_study$`PUR-2`))
```

**scales**

```{r}
psych::corPlot(r = cor(ques_study[, c("mean_AIAS", "mean_PTTA", "PUR-1", "PUR-2", "mean_valence")], use = "pairwise"))
```


# Analyze data - scales

```{r}
ggstatsplot::ggbetweenstats(data = ques_study, x = study_condition, y = mean_valence, type = "parametric")
# ggstatsplot::ggbetweenstats(data = mean_sd_valence_per_person, x = study_condition, y = mean_valence, type = "parametric")

# Run the Between-Subjects ANOVA
model_anova <- aov_ez(
  id = "PROLIFIC_PID",      # Replace with your subject ID column name
  dv = "mean_valence",           # Your Dependent Variable
  data = ques_study, 
  between = "study_condition" # Your Independent Variable
)

# Print the ANOVA table
print(model_anova)

# Get the marginal means
ls_means <- emmeans(model_anova, specs = "study_condition")

# Perform pairwise comparisons
post_hocs <- pairs(ls_means, adjust = "tukey")

# View results
print(post_hocs)
```

```{r}
ass_study_clean <- ass_study %>%
  filter(!is.na(sentiment_label),
         !is.na(study_condition)) %>%
  mutate(
    sentiment_label = factor(sentiment_label),
    study_condition = factor(study_condition)
  )

tab <- table(ass_study_clean$sentiment_label,
             ass_study_clean$study_condition)

tab

chisq_res <- chisq.test(tab)

chisq_res

chisq_res$expected

cramers_v(tab)


ggplot(ass_study_clean,
       aes(x = study_condition,
           fill = sentiment_label)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Study Condition",
    y = "Proportion",
    fill = "Sentiment",
    title = "Sentiment Distribution by Study Condition"
  ) +
  theme_minimal(base_size = 14)
```




```{r}
ggstatsplot::ggbetweenstats(data = ques_study, x = study_condition, y = `PUR-2`, type = "parametric")

# Run the Between-Subjects ANOVA
model_anova <- aov_ez(
  id = "PROLIFIC_PID",      # Replace with your subject ID column name
  dv = "PUR-2",           # Your Dependent Variable
  data = ques_study, 
  between = "study_condition" # Your Independent Variable
)

# Print the ANOVA table
print(model_anova)

# Get the marginal means
ls_means <- emmeans(model_anova, specs = "study_condition")

# Perform pairwise comparisons
post_hocs <- pairs(ls_means, adjust = "tukey")

# View results
print(post_hocs)
```

```{r}
ggstatsplot::ggbetweenstats(data = ques_study, x = study_condition, y = mean_AIAS, type = "parametric")

# Run the Between-Subjects ANOVA
model_anova <- aov_ez(
  id = "PROLIFIC_PID",      # Replace with your subject ID column name
  dv = "mean_AIAS",           # Your Dependent Variable
  data = ques_study, 
  between = "study_condition" # Your Independent Variable
)

# Print the ANOVA table
print(model_anova)


# Get the marginal means
ls_means <- emmeans(model_anova, specs = "study_condition")

# Perform pairwise comparisons
post_hocs <- pairs(ls_means, adjust = "tukey")

# View results
print(post_hocs)
```



```{r}
ggstatsplot::ggbetweenstats(data = ques_study, x = study_condition, y = mean_PTTA, type = "parametric")

# Run the Between-Subjects ANOVA
model_anova <- aov_ez(
  id = "PROLIFIC_PID",      # Replace with your subject ID column name
  dv = "mean_PTTA",           # Your Dependent Variable
  data = ques_study, 
  between = "study_condition" # Your Independent Variable
)

# Print the ANOVA table
print(model_anova)


# Get the marginal means
ls_means <- emmeans(model_anova, specs = "study_condition")

# Perform pairwise comparisons
post_hocs <- pairs(ls_means, adjust = "tukey")

# View results
print(post_hocs)
```

# Analyze data - word associations

```{r}
normalize_text <- function(text,
                           lowercase = TRUE,
                           remove_punctuation = TRUE,
                           remove_numbers = TRUE,
                           remove_extra_spaces = TRUE,
                           normalize_umlauts = TRUE) {
  
  # ensure character
  text <- as.character(text)
  
  # convert encoding safely (important for German text)
  text <- iconv(text, to = "UTF-8")
  
  # lowercase
  if (lowercase) {
    text <- tolower(text)
  }
  
  # normalize German umlauts (optional but common)
  if (normalize_umlauts) {
    text <- gsub("ä", "ae", text, fixed = TRUE)
    text <- gsub("ö", "oe", text, fixed = TRUE)
    text <- gsub("ü", "ue", text, fixed = TRUE)
    text <- gsub("ß", "ss", text, fixed = TRUE)
  }
  
  # remove punctuation
  if (remove_punctuation) {
    text <- gsub("[[:punct:]]+", " ", text)
  }
  
  # remove numbers
  if (remove_numbers) {
    text <- gsub("[0-9]+", " ", text)
  }
  
  # remove extra whitespace
  if (remove_extra_spaces) {
    text <- gsub("\\s+", " ", text)
  }
  
  # trim leading/trailing spaces
  text <- trimws(text)
  
  return(text)
}

length(unique(ass_study$response))
length(unique(ass_study$response_translated))

ass_study$response_translated_clean <- normalize_text(text = ass_study$response_translated)

length(unique(ass_study$response_translated_clean))
```

