---
title: "Data preperation all studies"
author: "Richard Bergs, Julius Fenn"
format:
  html:
    toc: true
    toc-depth: 3
    html-math-method: katex
    number-sections: true
---


# Notes


# global variables

Define your global variables (e.g., to reduce run time):

```{r}
#| label: global variables

createRawFiles <- TRUE
```

# functions

```{r}
########################################
# from JATOS data set to classical data frame (2D)
########################################
### argss:
# dataset = dat_secondPostCAM
# listvars = vec_ques
# notNumeric = vec_notNumeric
# verbose = TRUE


questionnairetype <- function(dataset,
                              listvars = ques_mixed,
                              notNumeric = vec_notNumeric,
                              verbose=FALSE){

  datasetques <- data.frame(ID = unique(dataset$ID))

  for(c in 1:length(listvars)){
    if(verbose){
      print(c)
    }

    if(any(colnames(dataset) == listvars[c])){
      if(verbose){
        print(listvars[c])
      }

      ## tmp IDs
      tmpid <- dataset$ID[!is.na(dataset[, listvars[c]])]
      ## tmp value variable
      tmpvalue <- dataset[, listvars[c]][!is.na(dataset[, listvars[c]])]
      datasetques[listvars[c]]  <- NA




      if(listvars[c] %in% notNumeric){
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- tmpvalue
      }else if(is.list(tmpvalue)){
        tmpvalue_tmp <- unique(tmpvalue)
        tmpvalue <- c()
        counter = 1
        for(i in 1:length(tmpvalue_tmp)){
          if(!is.null(tmpvalue_tmp[[i]])){
            tmpvalue[counter] <- paste0(tmpvalue_tmp[[i]], collapse = " - ")
            counter = counter + 1
          }
        }
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- tmpvalue
      }else{
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- as.numeric(tmpvalue)
      }
    }
  }
  return(datasetques)
}
```



```{r}
########################################
# from list of associations to classical data frame (2D)
########################################
### argss:
# association_list = sucsessfulAssociations
# metadata_df = ques_study


# association_list = sucsessfulAssociations
# metadata_df = ques_study
process_association_data <- function(association_list, metadata_df) {
  # Initialize output data frame
  output_df <- data.frame(
    participant_id = character(),
    study_condition = character(),
    gender = character(),
    age = integer(),
    cue = character(),
    response = character(),
    response_position = integer(),
    response_level = integer(),
    timestamp = character(),
    time_diff_sec = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Loop through all participants
  for (i in seq_along(association_list)) {
    
    # Get associations
    assoc <- association_list[[i]]
    assoc_L1 <- assoc[1:5, ]
    assoc_L2 <- assoc[6:nrow(assoc), ]
    
    # Get participant metadata
    meta <- metadata_df[i, ]
    pid <- meta$PROLIFIC_PID
    study_condition <- meta$study_condition
    gender <- meta$sociodemo_gender
    age <- meta$sociodemo_age 

    if(nrow(assoc_L1) != 5){
      cat("assoc_L1 not 5 for:", pid, "\n")
    }
    if(nrow(assoc_L2) != 25){
      cat("assoc_L2 not 25 but", nrow(assoc_L2), "for:", pid, "\n")
    }
    
    if(is.null(assoc_L1$valence)){
      assoc_L1$valence <- NA
    }
    if(is.null(assoc_L2$valence)){
      assoc_L2$valence <- NA
    }
    
    # --- Level 1: responses ---
    df_level1 <- data.frame(
      participant_id = pid,
      study_condition = study_condition,
      gender = gender,
      age = age,
      cue = assoc_L1$cue,
      valence = assoc_L1$valence,
      response = assoc_L1$response,
      response_position = seq_along(assoc_L1$cue),
      response_level = 1,
      timestamp = assoc_L1$timestamp,
      stringsAsFactors = FALSE
    )
    
    # --- Level 2: responses ---
    df_level2 <- data.frame(
 participant_id = pid,
      study_condition = study_condition,
      gender = gender,
      age = age,
      cue = assoc_L2$cue,
      valence = assoc_L2$valence,
      response = assoc_L2$response,
      response_position = as.numeric(ave(assoc_L2$cue, assoc_L2$cue, FUN = seq_along)), #rep(1:5, times = 5)
      response_level = 2,
      timestamp = assoc_L2$timestamp,
      stringsAsFactors = FALSE
    )
    
    # Combine both levels
    combined <- rbind(df_level1, df_level2)
    
    # Convert timestamp and calculate time difference from first response
    combined$timestamp <- as.POSIXct(combined$timestamp, format = "%Y-%m-%dT%H:%M:%OSZ", tz = "UTC")
    combined$time_diff_sec <- as.numeric(difftime(combined$timestamp, combined$timestamp[1], units = "secs"))
    
    # Append to final output
    output_df <- rbind(output_df, combined)
  }
  
  return(output_df)
}
```


# load packages

```{r}
#| label: load packages
#| warning: false

### load packages (also packages loaded, which are not needed for data preperation)
require(pacman)
p_load('tidyverse', 'jsonlite',
       'stargazer',  'DT', 'psych',
       'writexl', 'associatoR')
# devtools::install_github("samuelae/associatoR")
```

# create raw data files

```{r}
#| label: process-studies
#| warning: false

if(createRawFiles){
  # Define root paths
  source_root <- "data"
  output_dir <- "outputs/raw"
  
  
  # Define studies with paths
  studies <- list(
    study = file.path(source_root)
  )
  
  # Loop over each study
  for (study_name in names(studies)) {
    study_path <- studies[[study_name]]
    
    # Get folders matching pattern "study_result*"
    tmp_folders <- list.files(study_path, pattern = "^study_result.*", full.names = TRUE)
    
    results <- list()
    
    for (folder in tmp_folders) {
      # Look for a single subdirectory inside the folder
      inner_folders <- list.files(folder, full.names = TRUE)
      
      if (length(inner_folders) == 1) {
        data_file <- file.path(inner_folders[1], "data.txt")
        
        if (file.exists(data_file)) {
          # Load and store the JSON data
          tmp <- fromJSON(data_file)
          results[[length(results) + 1]] <- tmp
        }
      }
    }
    
    # Save as RDS
    saveRDS(results, file = file.path(output_dir, paste0(study_name, ".rds")))
    
    # Optionally also save as JSON Lines (.jsonl)
    json_lines <- map_chr(results, ~ toJSON(.x, auto_unbox = TRUE))
    writeLines(json_lines, con = file.path(output_dir, paste0(study_name, ".jsonl")))
  }
  print("Raw files have been created in this run!")
}else{
  print("Raw files have been created in previous run!")
}
```


# prepare data questionnaires

## study 1

load data:

```{r}
output_dir <- "outputs/raw"

# Load the RDS files into a named list
study <- readRDS(file.path(output_dir, "study.rds"))

# Combine into single data.frame
suppressMessages({
  dat <- bind_rows(study)
})

rm(study)
```

counter variable:

```{r}
dat$ID <- cumsum(dat$sender == "Greetings" & !is.na(dat$sender))

table(dat$ID)

tmp <- dat %>%
  group_by(ID) %>%
  summarise(N = n())
table(tmp$N)
```


filter:

```{r}
# IDs, die zu nicht-NULL gehören
tmp_ids <- dat$ID[!sapply(dat$sucsessfulAssociations, is.null)]

# die nicht-NULL Dataframes
tmp_cleaned <- dat$sucsessfulAssociations[!sapply(dat$sucsessfulAssociations, is.null)]

# alles zu einem Dataframe zusammenkleben + IDs
all_cols <- Reduce(union, lapply(tmp_cleaned, names))

tmp_cleaned2 <- lapply(tmp_cleaned, function(d) {
    missing <- setdiff(all_cols, names(d))
    for (m in missing) d[[m]] <- NA
    d[all_cols]  # reorder consistently
})

tmp_df <- do.call(rbind, Map(function(d, id) cbind(ID = id, d),
                             tmp_cleaned2, tmp_ids))
nrow(tmp_df)

tmp <- tmp_df %>%
  group_by(ID) %>%
  summarise(N = n())
table(tmp$N)

tmp_df <- tmp_df[tmp_df$ID %in% names(table(tmp_df$ID))[table(tmp_df$ID) == 30],]
nrow(tmp_df)

nrow(dat)
dat <- dat[dat$ID %in% tmp_df$ID, ]
nrow(dat)

tmp <- dat %>%
  group_by(ID) %>%
  summarise(N = n())
table(tmp$N)
```

```{r}
# sum(table(dat$ID) != max(table(dat$ID)))
# sum(table(dat$ID) == max(table(dat$ID)))
# 
# 
# length(unique(dat$ID))
# dat <-
#   dat[dat$ID %in% names(table(dat$ID))[table(dat$ID) == max(table(dat$ID))], ]
# length(unique(dat$ID))
```

**questionnaire**:

some elements are lists:

```{r}
colnames(dat)[sapply(dat, is.list)]
```

add meta manually and remove it:

```{r}
meta_df <- dat$meta[, c("language", "screen_width", "screen_height", "userAgent")]
dat <- bind_cols(dat, meta_df)
rm(meta_df)
dat$meta <- NULL
```
create questionnaire:

```{r}
colnames(dat)
```


```{r}
vec_notNumeric <- c("PROLIFIC_PID",
                    
                    "study_condition",
                    
                     "sociodemo_gender", 
                    
                    "feedback_attentionCheck", "feedback_technicalprobs", "feedback_technicalprobsText",
                    
                    "language", "screen_width", "screen_height", "userAgent")

### get survey
vec_ques <- c("PROLIFIC_PID",
              "dummy_informedconsent",
              "sociodemo_age", "lrscale", "rlgdgr",
              "feedback_conscientiousCompletion",
              "commCheck", "attCheck", "feedback_conscientiousCompletion",
                    str_subset(string = colnames(dat), pattern = "^PUR"),
                    str_subset(string = colnames(dat), pattern = "^AIAS"),
                    str_subset(string = colnames(dat), pattern = "^PTTA"),

              vec_notNumeric)



ques_study <- questionnairetype(
  dataset = dat,
  listvars = vec_ques,
  notNumeric = vec_notNumeric,
  verbose = FALSE
)

ques_study$study_condition <- factor(x = ques_study$study_condition)
ques_study$sociodemo_gender <- factor(x = ques_study$sociodemo_gender)


dim(ques_study)
```

### reverse code items


```{r}
psych::cor.plot(r = cor(ques_study[,str_subset(string = colnames(ques_study), pattern = "^PUR")]))
psych::cor.plot(r = cor(ques_study[,str_subset(string = colnames(ques_study), pattern = "^AIAS")]))
# items which have a suffix "r" need to be reverse coded:
psych::cor.plot(r = cor(ques_study[,str_subset(string = colnames(ques_study), pattern = "^PTTA")]))
```

```{r}
ques_study$`PTTA-2r` <- 6 - ques_study$`PTTA-2r`
ques_study$`PTTA-4r` <- 6 - ques_study$`PTTA-4r`
ques_study$`PTTA-5r` <- 6 - ques_study$`PTTA-5r`
```


```{r}
psych::cor.plot(r = cor(ques_study[,str_subset(string = colnames(ques_study), pattern = "^PTTA")]))
```



# prepare data associations


```{r}
colnames(dat)[sapply(dat, is.list)]
```

```{r}
sucsessfulAssociations <- dat$sucsessfulAssociations[!sapply(dat$sucsessfulAssociations, is.null)]

ass_study <- process_association_data(
  association_list = sucsessfulAssociations,
  metadata_df = ques_study
)

ass_study$study_condition <- factor(x = ass_study$study_condition)
ass_study$gender <- factor(x = ass_study$gender)
# Convert the character strings to POSIXct date-time objects
ass_study$timestamp <- ymd_hms(ass_study$timestamp)



dim(ass_study)


hist(ass_study$timestamp, breaks = "days", 
     main = "Frequency of Data Collection - Associations", 
     xlab = "Time", col = "skyblue")


# DT::datatable(data = ass_study)
```

```{r}
table(ass_study$study_condition) /30

tmp <- ass_study$study_condition[ass_study$response_position == 1 & ass_study$response_level == 1]
table(tmp)
```


# delete test data

```{r}
# first 30 associations are test data
ass_study$response[1:30]

dim(ass_study)
ass_study <- ass_study[31:nrow(ass_study), ]
dim(ass_study)


dim(ques_study)
ques_study <- ques_study[ques_study$PROLIFIC_PID %in% ass_study$participant_id, ]
dim(ques_study)
```

## compare study conditions between data sets

```{r}
## questionnaire
table(ques_study$study_condition)
## association data
table(ass_study$study_condition) /30
```




# save data sets



## save questionnaire

```{r}
setwd("outputs/questionnaire")

saveRDS(ques_study, file = paste0("ques_study", ".rds"))
writexl::write_xlsx(x = ques_study, path = paste0("ques_study", ".xlsx"))
```

## save associations


```{r}
setwd("outputs/associations")

# Save RDS
saveRDS(ass_study, file = "ass_study.rds")

# Save CSV
write.csv(ass_study, file = "ass_study.csv", row.names = FALSE)

# Save XLSX using writexl
writexl::write_xlsx(ass_study, path = "ass_study.xlsx")
```

