---
title: "Data Preparation: Survey + Transcripts + Sentiment Analysis"
author: "Julius Fenn"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    html-math-method: katex
    number-sections: true
    embed-resources: true
---

# Overview

This document prepares the complete dataset by combining:
1. **Survey Data** from `data/results_survey/` (questionnaire responses)
2. **Audio Transcripts** from `outputs/audio_transcripts/` (text_missingUtopia, text_rankingUtopias)
3. **Sentiment Analysis** from `outputs/sentiment_analysis/` (overall + window-level)

Final output: One row per participant with all metadata integrated.

---

# Setup & Configuration

```{r}
#| label: setup
#| message: false
#| warning: false

# Clear environment
rm(list = ls())

# Set control variables
createRawFiles <- TRUE  # Set to FALSE to skip raw file creation
VERBOSE <- TRUE

# Load packages
require(pacman)
p_load('tidyverse', 'jsonlite', 'writexl', 'lubridate', 'DT', 'psych')

# Set base directory
BASE_DIR <- getwd()
print(paste("Working directory:", BASE_DIR))

# Define paths
DATA_SURVEY_DIR <- file.path(BASE_DIR, "data", "results_survey")
AUDIO_TRANSCRIPT_DIR <- file.path(BASE_DIR, "outputs", "audio_transcripts")
SENTIMENT_DIR <- file.path(BASE_DIR, "outputs", "sentiment_analysis")

OUTPUT_RAW_DIR <- file.path(BASE_DIR, "outputs", "raw")
OUTPUT_QUESTIONNAIRE_DIR <- file.path(BASE_DIR, "outputs", "questionnaire")

# Create output directories if they don't exist
dir.create(OUTPUT_RAW_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(OUTPUT_QUESTIONNAIRE_DIR, showWarnings = FALSE, recursive = TRUE)

print("✓ Setup complete")
```

---

# Helper Functions

## Questionnaire Type Function

```{r}
#| label: questionnaire-function

questionnairetype <- function(dataset,
                              listvars = ques_mixed,
                              notNumeric = vec_notNumeric,
                              verbose = FALSE) {

  # Convert JATOS data to classical 2D data frame
  # 
  # Args:
  #   dataset: Raw data frame from JATOS
  #   listvars: Vector of column names to extract
  #   notNumeric: Vector of column names to keep as character
  #   verbose: Print progress
  # 
  # Returns:
  #   Data frame with 1 row per unique ID

  
  datasetques <- data.frame(ID = unique(dataset$ID))
  
  for (c in 1:length(listvars)) {
    if (verbose) {
      print(c)
    }
    
    if (any(colnames(dataset) == listvars[c])) {
      if (verbose) {
        print(listvars[c])
      }
      
      ## tmp IDs
      tmpid <- dataset$ID[!is.na(dataset[, listvars[c]])]
      ## tmp value variable
      tmpvalue <- dataset[, listvars[c]][!is.na(dataset[, listvars[c]])]
      datasetques[listvars[c]] <- NA
      
      if (listvars[c] %in% notNumeric) {
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- tmpvalue
      } else if (is.list(tmpvalue)) {
        tmpvalue_tmp <- unique(tmpvalue)
        tmpvalue <- c()
        counter <- 1
        for (i in 1:length(tmpvalue_tmp)) {
          if (!is.null(tmpvalue_tmp[[i]])) {
            tmpvalue[counter] <- paste0(tmpvalue_tmp[[i]], collapse = " - ")
            counter <- counter + 1
          }
        }
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- tmpvalue
      } else {
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- as.numeric(tmpvalue)
      }
    }
  }
  return(datasetques)
}


last_non_empty <- function(x) {
  if (length(x) == 0) return(NA)
  for (i in seq_along(x)[length(x):1]) {
    val <- x[[i]]
    if (is.null(val)) next
    if (!is.list(val) && length(val) == 1 && is.na(val)) next
    if (is.character(val) && length(val) == 1 && isTRUE(val == "")) next
    return(val)
  }
  return(NA)
}



print("✓ Helper functions defined")
```

---

# Load & Process Raw Survey Data

## Load all data.txt files

```{r}
#| label: load-raw-survey-data
#| message: false

if (createRawFiles) {
  print("Loading raw survey data from data.txt files...")
  
  # Find all data.txt files
  data_files <- list.files(
    path = DATA_SURVEY_DIR,
    pattern = "^data\\.txt$",
    full.names = TRUE,
    recursive = TRUE
  )
  
  print(paste("Found", length(data_files), "survey data files"))
  
  if (VERBOSE) {
    print("Files found:")
    print(data_files)
  }
  
  # Load all files
  results <- list()
  
  for (file in data_files) {
    tryCatch({
      tmp <- fromJSON(file)
      tmp[, colnames(tmp) == ""] <- NULL # remove empty variable
      results[[length(results) + 1]] <- tmp
      
      if (VERBOSE) {
        print(paste("✓ Loaded:", basename(dirname(dirname(file)))))
      }
    }, error = function(e) {
      print(paste("✗ Error loading", file, ":", e$message))
    })
  }
  
  print(paste("Successfully loaded", length(results), "files"))
  
  # Save as RDS
  saveRDS(results, file = file.path(OUTPUT_RAW_DIR, "study.rds"))
  
  # Optionally save as JSONL
  json_lines <- map_chr(results, ~ toJSON(.x, auto_unbox = TRUE))
  writeLines(json_lines, con = file.path(OUTPUT_RAW_DIR, "study.jsonl"))
  
  print("✓ Raw files saved (study.rds and study.jsonl)")
  
} else {
  print("Loading pre-saved raw data...")
  results <- readRDS(file = file.path(OUTPUT_RAW_DIR, "study.rds"))
  print(paste("Loaded", length(results), "records from study.rds"))
}
```

---

# Create Questionnaire DataFrame with Filtering


## Combine and prepare raw data

```{r}
#| label: combine-raw-data

for(i in 1:length(results)) {
  tmp <- results[[i]]$meta
  results[[i]]$meta <- NULL
  
    if(any(colnames(results[[i]]) %in% "paradata_general")) {
      results[[i]]$paradata_general <- NULL
    }
  
  results[[i]] <- cbind(results[[i]], tmp)
}


# Combine all records into single dataframe
suppressMessages({
  dat <- bind_rows(results)
})


# Combine all records into single dataframe
suppressMessages({
  dat <- bind_rows(results)
})

print(paste("Combined data shape:", nrow(dat), "rows x", ncol(dat), "columns"))
rm(results)

# Add ID counter
dat$ID <- cumsum(dat$sender == "Greetings" & !is.na(dat$sender))

print("\n=== Initial ID Distribution ===")
print(table(dat$ID))

# Check record counts per ID
tmp <- dat %>%
  group_by(ID) %>%
  summarise(N = n(), .groups = 'drop')

print("\n=== Record Count Distribution ===")
print(table(tmp$N))

max_N <- max(tmp$N)
print(paste("Maximum records per ID:", max_N))
```


## Identify incomplete records and test data

```{r}
#| label: filter-incomplete-records

# Initialize filtering tracking
filtering_stats <- data.frame(
  metric = character(),
  value = numeric(),
  stringsAsFactors = FALSE
)

starting_N <- nrow(dat)
filtering_stats <- rbind(filtering_stats, 
                         data.frame(metric = "starting_total_records", value = starting_N))

# Find IDs with incomplete records
tmp_full <- dat %>%
  group_by(ID) %>%
  summarise(N = n(), .groups = 'drop')

incomplete_IDs <- tmp_full$ID[tmp_full$N != max_N]
n_incomplete_before <- length(incomplete_IDs)

print(paste("\n=== Filtering Step 1: Identify Incomplete Records ==="))
print(paste("IDs with fewer than", max_N, "records:", n_incomplete_before))
if (VERBOSE & n_incomplete_before > 0) {
  print("Incomplete IDs:")
  print(incomplete_IDs)
  
  filtering_stats <- rbind(filtering_stats, 
                         data.frame(metric = "incomplete_ids", value = paste0(incomplete_IDs, collapse = " // ")))
}

dat <- dat[!dat$ID %in% incomplete_IDs, ]

if (VERBOSE & n_incomplete_before > 0) {
  filtering_stats <- rbind(filtering_stats, 
                         data.frame(metric = "after_filtering_incomplete_records", value = nrow(dat)))
}



# Handle test ID "111" - Special logic
print("\n=== Filtering Step 2: Handle Test ID '111' ===")

test_ids <- dat$ID[dat$PROLIFIC_PID == "111"] # !!!
test_ids <- test_ids[!is.na(test_ids)]
n_test_runs <- length(unique(test_ids))

print(paste("Found", n_test_runs, "test run(s) with ID '111'"))

if (VERBOSE & n_test_runs > 0) {
  print("Test IDs:")
  print(test_ids)
  
  filtering_stats <- rbind(filtering_stats, 
                         data.frame(metric = "test_ids", value = paste0(test_ids, collapse = " // ")))
}

dat <- dat[!dat$ID %in% test_ids, ]


if (VERBOSE & n_test_runs > 0) {
  filtering_stats <- rbind(filtering_stats, 
                         data.frame(metric = "after_removing_testIDs_records", value = nrow(dat)))
}


# Update participant count
n_participants_final <- length(unique(dat$ID))
filtering_stats <- rbind(filtering_stats,
                         data.frame(metric = "final_unique_participants", value = n_participants_final))

cat("\n✓ Final unique participants:", n_participants_final)
```

## Extract questionnaire

```{r}
colnames(dat)[sapply(dat, is.list)]
```


```{r}
colnames(dat)
```

```{r}
tmp <- dat %>%
  select(-where(is.list)) %>%
  mutate(across(-ID, as.character)) %>%
  pivot_longer(-ID) %>%
  group_by(ID, name) %>%
  summarise(n_unique = n_distinct(value, na.rm = TRUE), .groups = 'drop') %>%
  filter(n_unique > 1)

tmp$name
```



```{r}
print("\n=== Creating Questionnaire DataFrame ===")


# ============================================================================
# PSYCHOLOGICAL SCALES - Organized by Type
# ============================================================================

# --- ACTIVISM & RADICALISM INTENTIONS (ARIS-adapted) ---
aris_activism <- c(
  "aris_scale-ais1",   # Join organization
  "aris_scale-ais2",   # Donate money
  "aris_scale-ais3",   # Volunteer time
  "aris_scale-ais4"    # Travel to rally
)

aris_radicalism <- c(
  "aris_scale-ris1",   # Support despite breaking law
  "aris_scale-ris2",   # Support despite violence
  "aris_scale-ris3",   # Protest despite violence
  "aris_scale-ris4"    # Attack security forces
)

# --- UTOPIAN THINKING (Fernando et al. 2018) ---
utopianism <- c(
  "utopianism_scale-utop1",   # Think about ideal society
  "utopianism_scale-utop2",   # Spend time thinking
  "utopianism_scale-utop3",   # What world I'd like
  "utopianism_scale-utop4"    # Important to think
)

anti_utopianism <- c(
  "anti_utopianism_scale-antiutop1",   # Could be dangerous
  "anti_utopianism_scale-antiutop2",   # Negative consequences
  "anti_utopianism_scale-antiutop3",   # Shouldn't envision
  "anti_utopianism_scale-antiutop4"    # Useless to dream
)

# --- NEED FOR CHAOS (NFC) - Arceneaux et al. 2021 ---
need_for_chaos <- c(
  "nfc_scale-nfc1",   # Kick from disasters
  "nfc_scale-nfc2",   # Fantasize wipeout
  "nfc_scale-nfc3",   # Burn society
  "nfc_scale-nfc4",   # Let them burn
  "nfc_scale-nfc5",   # Tear down & restart
  "nfc_scale-nfc6",   # Need chaos
  "nfc_scale-nfc7"    # Destroy beautiful things
)

# --- POLITICAL/INSTITUTIONAL ATTITUDES ---
system_justification <- c(
  "sjs_scale-sjs1",    # Society is fair
  "sjs_scale-sjs2",    # Political system works
  "sjs_scale-sjs3r",   # [REVERSE] Needs restructuring
  "sjs_scale-sjs4",    # Best country
  "sjs_scale-sjs5",    # Policies serve greater good
  "sjs_scale-sjs6",    # Fair shot at wealth
  "sjs_scale-sjs7r",   # [REVERSE] Getting worse
  "sjs_scale-sjs8"     # Get what they deserve
)

# --- WELL-BEING & RESILIENCE ---
satisfaction_with_life <- c(
  "swls_scale-swls1",  # Close to ideal
  "swls_scale-swls2",  # Excellent conditions
  "swls_scale-swls3",  # Satisfied with life
  "swls_scale-swls4",  # Got important things
  "swls_scale-swls5"   # Change almost nothing
)

brief_resilience <- c(
  "brs_scale-brs1",    # Bounce back quickly
  "brs_scale-brs2r",   # [REVERSE] Hard time through stress
  "brs_scale-brs3",    # Don't take long to recover
  "brs_scale-brs4r",   # [REVERSE] Hard to snap back
  "brs_scale-brs5",    # Come through with trouble
  "brs_scale-brs6r"    # [REVERSE] Long time to get over
)

# --- DEPRIVATION & CONSPIRACY ---
personal_relative_deprivation <- c(
  "prds_scale-prds1",  # Feel deprived
  "prds_scale-prds2r", # [REVERSE] Feel privileged
  "prds_scale-prds3",  # Feel resentful
  "prds_scale-prds4r", # [REVERSE] Quite well off
  "prds_scale-prds5"   # Dissatisfied
)

conspiracy_mentality <- c(
  "cmq_scale-cmq1",    # Important things hidden
  "cmq_scale-cmq2",    # Politicians don't tell truth
  "cmq_scale-cmq3",    # Government monitors
  "cmq_scale-cmq4",    # Secret activities
  "cmq_scale-cmq5"     # Secret organizations influence
)



# ============================================================================
# SOCIO-DEMOGRAPHIC VARIABLES, FEEDBACK
# ============================================================================
socio_demo <- c(
  "age",
  "gender",
  "education",
  "employment_status",
  "employment_is_employed",
  "lrscale",           # Left-right political scale
  "rlgdgr"             # Religious degree
)

# --- COMBINED SCALE VECTOR ---
vec_ques <- c(
  "PROLIFIC_PID",
  "dummy_informedconsent",
  "commCheck",
  "mic_test_passed",
  "feedback_critic",
  socio_demo,
  # scales:
  aris_activism,
  aris_radicalism,
  utopianism,
  anti_utopianism,
  need_for_chaos,
  system_justification,
  satisfaction_with_life,
  brief_resilience,
  personal_relative_deprivation,
  conspiracy_mentality
)


vec_notNumeric <- c(
  "PROLIFIC_PID",
  # Socio-demographic non-numeric
  "gender",
  "education",
  "employment_status",
  # Other non-numeric
  "mic_test_passed",
  "feedback_critic"
)


# ============================================================================
# Create questionnaire
# ============================================================================
ques_study <- questionnairetype(
  dataset = dat,
  listvars = vec_ques,
  notNumeric = vec_notNumeric,
  verbose = FALSE
)


ques_study$gender <- factor(x = ques_study$gender)
ques_study$education <- factor(x = ques_study$education)
ques_study$employment_status <- factor(x = ques_study$employment_status)


# Save summary
filtering_stats <- rbind(filtering_stats,
                         data.frame(metric = "final_questionnaire_rows", value = nrow(ques_study)))
```


## add rankings



```{r}
## PROLIFIC_PIDs first run BEFORE improved ranking task:
potentialInconsistentRankings <- c("6a08b810d807f5805aabe02e", "67103e4ed56cde03a0f2476b", "59a07638bfd73c00010ea394", "63d7b1a77e23b5460ab88069", "69c4160ec318cfcd7d4f37a4", "69fe9e67a20599144f688b50", "69e91fd1a613772e648e565a", "673e5fc42cbf961c153d09d3", "6a04a5a7e230bb73b88bd6e6", "5f68fdce1c675a0baf13e0ff")

ranking_data <- dat %>%
  group_by(ID) %>%
  reframe(
    PROLIFIC_PID = last_non_empty(PROLIFIC_PID),
    ranking_final_order = last_non_empty(ranking_final_order)
  ) 


ranking_data$flagPotentialIncon <- FALSE
ranking_data$flagPotentialIncon[ranking_data$PROLIFIC_PID %in% potentialInconsistentRankings] <- TRUE

ranking_data$rankingOrder <- NA
ranking_data$rankingOrder[ranking_data$flagPotentialIncon] <- rep(1:7, times = length(potentialInconsistentRankings))


## for non-inconsistent potential rankings the new order:
num_non_InconsistentRankings <- sum(!unique(ranking_data$PROLIFIC_PID) %in% potentialInconsistentRankings)
ranking_data$rankingOrder[!ranking_data$flagPotentialIncon] <- rep(7:1, times = length(num_non_InconsistentRankings))


# to wide
ranking_data$ID <- NULL
ranking_wide <- ranking_data %>%
  pivot_wider(
    names_from = ranking_final_order,
    values_from = rankingOrder,
    names_prefix = ""
  )


ques_study <- ques_study %>%
  left_join(ranking_wide, by = "PROLIFIC_PID")


# !!! check with prompt below
table(ques_study$flagPotentialIncon)
```




# Add Transcripts

## Load transcript files

```{r}
#| label: load-transcripts

print("\n=== Loading Transcripts ===")

# Find all transcript files
transcript_files <- list.files(
  path = AUDIO_TRANSCRIPT_DIR,
  pattern = "_transcript\\.json$",
  full.names = TRUE
)

print(paste("Found", length(transcript_files), "transcript files"))

# Parse all transcripts
transcripts <- list()

for (file in transcript_files) {
  tryCatch({
    transcript <- fromJSON(file)
    filename <- basename(file)
    
    # Extract participant ID from filename
    # Format: {id_person}_{type}_{timestamp}_transcript.json
    participant_id <- str_extract(filename, "^[^_]+")
    audio_type <- str_extract(filename, "(?<=_)(audio_[^_]+)(?=_)")
    
    transcripts[[length(transcripts) + 1]] <- list(
      filename = filename,
      participant_id = participant_id,
      audio_type = audio_type,
      transcript_text = transcript$transcript,
      timestamp = transcript$timestamp
    )
    
  }, error = function(e) {
    print(paste("✗ Error loading", file, ":", e$message))
  })
}

print(paste("Successfully parsed", length(transcripts), "transcripts"))

# Convert to dataframe for easier processing
transcripts_df <- bind_rows(lapply(transcripts, as.data.frame, stringsAsFactors = FALSE))

if (VERBOSE) {
  print("Sample transcripts:")
  print(head(transcripts_df))
  print("\nTranscript types found:")
  print(table(transcripts_df$audio_type))
}
```

## Match transcripts to questionnaire and add

```{r}
#| label: add-transcripts-to-questionnaire

print("\n=== Adding Transcripts to Questionnaire ===")

# Initialize columns
ques_study$text_missingUtopia <- NA_character_
ques_study$text_rankingUtopias <- NA_character_

# For each participant in questionnaire
for (i in 1:nrow(ques_study)) {
  participant_id <- ques_study$PROLIFIC_PID[i]
  
  # Find matching transcripts
  matching_transcripts <- transcripts_df %>%
    filter(participant_id == !! participant_id)
  
  if (nrow(matching_transcripts) > 0) {
    # Process missing utopia transcripts
    missing_utopia <- matching_transcripts %>%
      filter(audio_type == "audio_missing") %>%
      arrange(desc(timestamp))
    
    if (nrow(missing_utopia) > 0) {
      # Keep only most recent
      ques_study$text_missingUtopia[i] <- missing_utopia$transcript_text[1]
      
      if (VERBOSE & nrow(missing_utopia) > 1) {
        print(paste("  Participant", participant_id, "- Kept 1 of", nrow(missing_utopia), "missingUtopia transcripts"))
      }
    }
    
    # Process ranking explanation transcripts
    ranking <- matching_transcripts %>%
      filter(audio_type == "audio_ranking") %>%
      arrange(desc(timestamp))
    
    if (nrow(ranking) > 0) {
      # Keep only most recent
      ques_study$text_rankingUtopias[i] <- ranking$transcript_text[1]
      
      if (VERBOSE & nrow(ranking) > 1) {
        print(paste("  Participant", participant_id, "- Kept 1 of", nrow(ranking), "rankingUtopias transcripts"))
      }
    }
  }
}

# Check coverage
n_with_missing_utopia <- sum(!is.na(ques_study$text_missingUtopia))
n_with_ranking <- sum(!is.na(ques_study$text_rankingUtopias))

print(paste("Transcripts added:"))
print(paste("  - text_missingUtopia:", n_with_missing_utopia, "/", nrow(ques_study)))
print(paste("  - text_rankingUtopias:", n_with_ranking, "/", nrow(ques_study)))

print("✓ Transcripts merged successfully")
```


---

# Add Sentiment Analysis

## Load overall sentiment data

```{r}
#| label: load-overall-sentiment

print("\n=== Loading Overall Sentiment Analysis ===")

# Find sentiment file
sentiment_files <- list.files(
  path = SENTIMENT_DIR,
  pattern = "overall_sentiment_and_embedding.*\\.csv$",
  full.names = TRUE
)

if (length(sentiment_files) > 0) {
  # Use most recent if multiple files exist
  sentiment_file <- sentiment_files[which.max(file.info(sentiment_files)$mtime)]
  
  print(paste("Using:", basename(sentiment_file)))
  
  sentiment_overall <- read.csv(sentiment_file, stringsAsFactors = FALSE)
  
  print(paste("Loaded", nrow(sentiment_overall), "sentiment records"))
  print("\nSample data:")
  print(head(sentiment_overall))
  
} else {
  print("✗ No overall sentiment file found!")
  sentiment_overall <- NULL
}
```

## Add overall sentiment to questionnaire

```{r}
#| label: add-overall-sentiment

if (!is.null(sentiment_overall)) {
  print("\n=== Adding Overall Sentiment to Questionnaire ===")
  
  # Initialize columns
  ques_study$sentiment_label <- NA_character_
  ques_study$sentiment_score <- NA_real_
  ques_study$num_windows <- NA_integer_
  
  # For each participant
  for (i in 1:nrow(ques_study)) {
    participant_id <- ques_study$PROLIFIC_PID[i]
    
    # Find matching sentiment records
    # Match using json_file which contains participant_id
    matching_sentiment <- sentiment_overall %>%
      filter(grepl(participant_id, json_file, fixed = TRUE))
    
    if (nrow(matching_sentiment) > 0) {
      # If multiple matches, use most recent (first one after sort)
      # For now, combine or average if multiple
      
      if (nrow(matching_sentiment) > 1) {
        # Average sentiment scores, keep first label
        ques_study$sentiment_label[i] <- matching_sentiment$sentiment_label[1]
        ques_study$sentiment_score[i] <- mean(matching_sentiment$sentiment_score, na.rm = TRUE)
        ques_study$num_windows[i] <- sum(matching_sentiment$num_windows, na.rm = TRUE)
        
        if (VERBOSE) {
          print(paste("  Participant", participant_id, "- Averaged", nrow(matching_sentiment), "sentiment records"))
        }
      } else {
        ques_study$sentiment_label[i] <- matching_sentiment$sentiment_label[1]
        ques_study$sentiment_score[i] <- matching_sentiment$sentiment_score[1]
        ques_study$num_windows[i] <- matching_sentiment$num_windows[1]
      }
    }
  }
  
  n_with_sentiment <- sum(!is.na(ques_study$sentiment_label))
  print(paste("Overall sentiment added to", n_with_sentiment, "participants"))
  
} else {
  print("Skipping overall sentiment (file not found)")
}
```

## Load window-level sentiment data

```{r}
#| label: load-window-sentiment

print("\n=== Loading Window-Level Sentiment ===")

# Find window sentiment file
window_files <- list.files(
  path = SENTIMENT_DIR,
  pattern = "window_sentiments.*\\.csv$",
  full.names = TRUE
)

if (length(window_files) > 0) {
  # Use most recent if multiple files exist
  window_file <- window_files[which.max(file.info(window_files)$mtime)]
  
  print(paste("Using:", basename(window_file)))
  
  sentiment_windows <- read.csv(window_file, stringsAsFactors = FALSE)
  
  # Calculate max window index
  max_window_idx <- max(sentiment_windows$window_index, na.rm = TRUE)
  
  print(paste("Loaded", nrow(sentiment_windows), "window sentiment records"))
  print(paste("Maximum window index:", max_window_idx))
  
  # Show sample
  print("\nSample data:")
  print(head(sentiment_windows))
  
} else {
  print("✗ No window sentiment file found!")
  sentiment_windows <- NULL
  max_window_idx <- 0
}
```

## Add window sentiment columns

```{r}
#| label: add-window-sentiment

if (!is.null(sentiment_windows)) {
  print("\n=== Adding Window-Level Sentiment Columns ===")
  
  # Create column names dynamically
  window_cols <- paste0("window_", 0:max_window_idx, "_sentimentLabel")
  
  print(paste("Creating", length(window_cols), "window sentiment columns"))
  
  # Initialize columns with NA
  for (col in window_cols) {
    ques_study[[col]] <- NA_character_
  }
  
  # Fill window sentiment for each participant
  for (i in 1:nrow(ques_study)) {
    participant_id <- ques_study$PROLIFIC_PID[i]
    
    # Find windows for this participant
    participant_windows <- sentiment_windows %>%
      filter(grepl(participant_id, json_file, fixed = TRUE))
    
    if (nrow(participant_windows) > 0) {
      # Fill window columns
      for (j in 1:nrow(participant_windows)) {
        window_idx <- participant_windows$window_index[j]
        col_name <- paste0("window_", window_idx, "_sentimentLabel")
        
        # Only fill if column exists
        if (col_name %in% colnames(ques_study)) {
          ques_study[i, col_name] <- participant_windows$sentiment_label[j]
        }
      }
      
      if (VERBOSE & i %in% c(1, 5, 10)) {
        print(paste("  Participant", participant_id, "- Added", nrow(participant_windows), "window sentiments"))
      }
    }
  }
  
  print(paste("✓ Window sentiment columns added"))
  
} else {
  print("Skipping window sentiment (file not found)")
}
```
```{r}
# Before
cat("Original columns:", ncol(ques_study), "\n")
cat("Columns that are all NA:", sum(sapply(ques_study, function(x) all(is.na(x)))), "\n")

# Apply removal (Option 1)
ques_study <- ques_study[, colSums(is.na(ques_study)) < nrow(ques_study)]

# After
cat("Final columns:", ncol(ques_study), "\n")
```


---

# Get within part of study



## vignette part

```{r}
# --- Get PROLIFIC_PID mapping ---
pid_lookup <- dat %>%
  group_by(ID) %>%
  summarise(PROLIFIC_PID = last_non_empty(PROLIFIC_PID), .groups = 'drop')

# --- Initialize results list ---
vignette_list <- list()

# --- Get unique IDs ---
unique_ids <- unique(dat$ID)

# --- LOOP 1: For each participant ---
for (id in unique_ids) {
  
  # Filter data for this participant
  dat_id <- dat %>% filter(ID == id)
  
  # Get their PROLIFIC_PID
  prolific_pid <- pid_lookup %>% 
    filter(ID == id) %>% 
    pull(PROLIFIC_PID)
  
  # --- LOOP 2: For each vignette trial (counterScenarios) ---
  unique_trials <- unique(na.omit(dat_id$counterScenarios))
  
  for (trial in unique_trials) {
    
    # Filter data for this trial
    trial_data <- dat_id %>% filter(counterScenarios == trial)
    
    # Extract values (take last non-empty per variable)
    row <- tibble(
      ID = id,
      PROLIFIC_PID = prolific_pid,
      counterScenarios = trial,
      vignette_order_position = last_non_empty(trial_data$vignette_order_position),
      vignette_id = last_non_empty(trial_data$vignette_id),
      condition_FutureSociety = last_non_empty(trial_data$condition_FutureSociety),
      # Participative efficacy
      pe1 = last_non_empty(trial_data$pe1),
      pe2 = last_non_empty(trial_data$pe2),
      pe3 = last_non_empty(trial_data$pe3),
      pe4 = last_non_empty(trial_data$pe4),
      # Attributes
      att_ideal = last_non_empty(trial_data$att_ideal),
      att_beneficial = last_non_empty(trial_data$att_beneficial),
      att_utopian = last_non_empty(trial_data$att_utopian),
      att_desirable = last_non_empty(trial_data$att_desirable),
      att_innovative = last_non_empty(trial_data$att_innovative),
      att_live_in = last_non_empty(trial_data$att_live_in),
      # Paradata
      para_countclicks = last_non_empty(trial_data$para_countclicks),
      time_run = last_non_empty(trial_data$time_run),
      time_render = last_non_empty(trial_data$time_render),
      time_show = last_non_empty(trial_data$time_show)
    )
    
    # Add to results list
    vignette_list[[length(vignette_list) + 1]] <- row
  }
  
  # Progress indicator
  # cat("Processed ID:", id, "\n")
}

# --- Combine all results into one dataframe ---
ques_study_within_vignettes <- bind_rows(vignette_list)

# --- Remove trials with all NA in PE/ATT ---
ques_study_within_vignettes <- ques_study_within_vignettes %>%
  filter(!if_all(pe1:att_live_in, is.na)) %>%
  arrange(ID, counterScenarios)

# --- Inspect results ---
head(ques_study_within_vignettes)
cat("Total vignette trials:", nrow(ques_study_within_vignettes), "\n")
cat("Trials per participant:\n")
print(table(ques_study_within_vignettes$ID))
```

## visual trap part

```{r}
trap_types <- c("Planets", "Robot", "Circles", "Lines", "CafeWall", "Collision")

visual_trap_within <- map_df(trap_types, function(trap) {
  dat %>%
    select(
      ID,
      counterVisualTraps,
      all_of(paste0("visualTrap_", trap, "_response")),
      all_of(paste0("visualTrap_", trap, "_responseLabel")),
      all_of(paste0("visualTrap_", trap, "_correct")),
      all_of(paste0("visualTrap_", trap, "_rtMs"))
    ) %>%
    # Remove rows where this trap is all NA
    filter(!if_all(starts_with("visualTrap_"), is.na)) %>%
    distinct(ID, counterVisualTraps, .keep_all = TRUE) %>%
    mutate(
      trap_type = trap,
      response = .data[[paste0("visualTrap_", trap, "_response")]],
      response_label = .data[[paste0("visualTrap_", trap, "_responseLabel")]],
      correct = .data[[paste0("visualTrap_", trap, "_correct")]],
      rt_ms = .data[[paste0("visualTrap_", trap, "_rtMs")]]
    ) %>%
    select(ID, counterVisualTraps, trap_type, response, response_label, correct, rt_ms)
}) %>%
  # Add PROLIFIC_PID from lookup
  left_join(pid_lookup, by = "ID") %>%
  select(ID, PROLIFIC_PID, counterVisualTraps, trap_type, everything()) %>%
  arrange(ID, counterVisualTraps, trap_type)

# --- Inspect results ---
head(visual_trap_within)
cat("Total vignette trials:", nrow(visual_trap_within), "\n")
cat("Trials per participant:\n")
print(table(visual_trap_within$ID))
```

# Summary stats and Quality checks


```{r}
print("FINAL DATA SUMMARY & VALIDATION")
print(strrep("=", 80))
 
# --- Count unique IDs across datasets ---
n_ids_wide <- n_distinct(ques_study$ID)
n_ids_vignettes <- n_distinct(ques_study_within_vignettes$ID)
n_ids_traps <- n_distinct(visual_trap_within$ID)
 
print("\n📊 PARTICIPANT COUNTS:")
print(paste("  - Between-subjects (ques_study):", n_ids_wide, "unique IDs"))
print(paste("  - Within-vignettes dataset:", n_ids_vignettes, "unique IDs"))
print(paste("  - Within-traps dataset:", n_ids_traps, "unique IDs"))
```

## VALIDATION CHECK 1: Matching ID counts

```{r}
print("VALIDATION 1: ID Count Matching")
print(strrep("-", 80))
 
id_count_mismatch <- FALSE
 
if (n_ids_wide != n_ids_vignettes) {
  print("❌ ERROR: ID count mismatch between ques_study and vignettes dataset!")
  print(paste("   ques_study IDs:", n_ids_wide))
  print(paste("   vignettes IDs:", n_ids_vignettes))
  id_count_mismatch <- TRUE
}
 
if (n_ids_wide != n_ids_traps) {
  print("❌ ERROR: ID count mismatch between ques_study and traps dataset!")
  print(paste("   ques_study IDs:", n_ids_wide))
  print(paste("   traps IDs:", n_ids_traps))
  id_count_mismatch <- TRUE
}
 
if (n_ids_vignettes != n_ids_traps) {
  print("❌ ERROR: ID count mismatch between vignettes and traps datasets!")
  print(paste("   vignettes IDs:", n_ids_vignettes))
  print(paste("   traps IDs:", n_ids_traps))
  id_count_mismatch <- TRUE
}
 
if (!id_count_mismatch) {
  print("✓ All datasets have matching ID counts")
}
```

## VALIDATION CHECK 2: ID set matching

```{r}
print("VALIDATION 2: ID Set Matching")
print(strrep("-", 80))
 
ids_wide <- sort(unique(ques_study$ID))
ids_vignettes <- sort(unique(ques_study_within_vignettes$ID))
ids_traps <- sort(unique(visual_trap_within$ID))
 
# Check vignettes vs wide
missing_in_vignettes <- setdiff(ids_wide, ids_vignettes)
extra_in_vignettes <- setdiff(ids_vignettes, ids_wide)
 
if (length(missing_in_vignettes) > 0) {
  print("❌ ERROR: IDs in ques_study but missing from vignettes dataset:")
  print(paste("   Missing:", paste(missing_in_vignettes, collapse = ", ")))
  id_count_mismatch <- TRUE
} else {
  print("✓ All ques_study IDs present in vignettes dataset")
}
 
if (length(extra_in_vignettes) > 0) {
  print("❌ ERROR: IDs in vignettes but not in ques_study:")
  print(paste("   Extra:", paste(extra_in_vignettes, collapse = ", ")))
  id_count_mismatch <- TRUE
}
 
# Check traps vs wide
missing_in_traps <- setdiff(ids_wide, ids_traps)
extra_in_traps <- setdiff(ids_traps, ids_wide)
 
if (length(missing_in_traps) > 0) {
  print("❌ ERROR: IDs in ques_study but missing from traps dataset:")
  print(paste("   Missing:", paste(missing_in_traps, collapse = ", ")))
  id_count_mismatch <- TRUE
} else {
  print("✓ All ques_study IDs present in traps dataset")
}
 
if (length(extra_in_traps) > 0) {
  print("❌ ERROR: IDs in traps but not in ques_study:")
  print(paste("   Extra:", paste(extra_in_traps, collapse = ", ")))
  id_count_mismatch <- TRUE
}
 
# STOP IF VALIDATION FAILED
if (id_count_mismatch) {
  stop("\n🛑 DATA INTEGRITY VALIDATION FAILED!\n   Please review ID mismatches above before proceeding.")
}
 
print("✓ ALL VALIDATION CHECKS PASSED!")
print(strrep("✓", 80))
```


## MISSING DATA REPORT

```{r}
print("MISSING DATA REPORT")
print(strrep("=", 80))
 
# For between-subjects (top columns with most missing)
missing_wide <- data.frame(
  Column = colnames(ques_study),
  Missing_Count = colSums(is.na(ques_study)),
  Missing_Percent = round(colSums(is.na(ques_study)) / nrow(ques_study) * 100, 2),
  stringsAsFactors = FALSE
) %>%
  filter(Missing_Count > 0) %>%
  arrange(desc(Missing_Count)) %>%
  head(15)
 
if (nrow(missing_wide) > 0) {
  print("\nBetween-subjects (ques_study) - Top 15 columns with missing data:")
  print(missing_wide)
} else {
  print("\n✓ No missing values in ques_study")
}
```

# Save Datasets


```{r}
# --- Vignette trials per participant ---
trial_dist <- ques_study_within_vignettes %>%
  group_by(ID) %>%
  summarise(n_trials = n(), .groups = 'drop') %>%
  pull(n_trials)

# --- Trap trials per participant ---
trap_dist <- visual_trap_within %>%
  group_by(ID) %>%
  summarise(n_trials = n(), .groups = 'drop') %>%
  pull(n_trials)
```


```{r}
# ============================================================================
# SAVE ALL DATASETS
# ============================================================================

print("SAVING FINAL DATASETS")
print(strrep("=", 80))

# --- Output directories ---
output_dir <- file.path(OUTPUT_QUESTIONNAIRE_DIR)
within_dir <- file.path(OUTPUT_QUESTIONNAIRE_DIR, "within_subjects")

# Create directories if they don't exist
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(within_dir, showWarnings = FALSE, recursive = TRUE)

# ============================================================================
# DATASET 1: BETWEEN-SUBJECTS (Wide Format)
# ============================================================================
print("\n📋 Saving BETWEEN-SUBJECTS dataset (ques_study)...")

# RDS format (preserves data types)
ques_study_file_rds <- file.path(output_dir, "ques_study_wide.rds")
saveRDS(ques_study, file = ques_study_file_rds)
print(paste("  ✓ Saved: ques_study_wide.rds"))

# CSV format
# ques_study_file_csv <- file.path(output_dir, "ques_study_wide.csv")
# write.csv(ques_study, file = ques_study_file_csv, row.names = FALSE)
# print(paste("  ✓ Saved: ques_study_wide.csv"))

# XLSX format
ques_study_file_xlsx <- file.path(output_dir, "ques_study_wide.xlsx")
writexl::write_xlsx(ques_study, path = ques_study_file_xlsx)
print(paste("  ✓ Saved: ques_study_wide.xlsx"))

# ============================================================================
# DATASET 2: WITHIN-SUBJECTS - VIGNETTES (Long Format)
# ============================================================================
print("\n📊 Saving WITHIN-SUBJECTS VIGNETTES dataset (ques_study_within_vignettes)...")

# RDS format
vignettes_file_rds <- file.path(within_dir, "ques_study_within_vignettes.rds")
saveRDS(ques_study_within_vignettes, file = vignettes_file_rds)
print(paste("  ✓ Saved: within_subjects/ques_study_within_vignettes.rds"))

# CSV format
vignettes_file_csv <- file.path(within_dir, "ques_study_within_vignettes.csv")
write.csv(ques_study_within_vignettes, file = vignettes_file_csv, row.names = FALSE)
print(paste("  ✓ Saved: within_subjects/ques_study_within_vignettes.csv"))

# XLSX format
vignettes_file_xlsx <- file.path(within_dir, "ques_study_within_vignettes.xlsx")
writexl::write_xlsx(ques_study_within_vignettes, path = vignettes_file_xlsx)
print(paste("  ✓ Saved: within_subjects/ques_study_within_vignettes.xlsx"))

# ============================================================================
# DATASET 3: WITHIN-SUBJECTS - VISUAL TRAPS (Long Format)
# ============================================================================
print("\n🎯 Saving WITHIN-SUBJECTS VISUAL TRAPS dataset (visual_trap_within)...")

# RDS format
traps_file_rds <- file.path(within_dir, "visual_trap_within.rds")
saveRDS(visual_trap_within, file = traps_file_rds)
print(paste("  ✓ Saved: within_subjects/visual_trap_within.rds"))

# CSV format
traps_file_csv <- file.path(within_dir, "visual_trap_within.csv")
write.csv(visual_trap_within, file = traps_file_csv, row.names = FALSE)
print(paste("  ✓ Saved: within_subjects/visual_trap_within.csv"))

# XLSX format
traps_file_xlsx <- file.path(within_dir, "visual_trap_within.xlsx")
writexl::write_xlsx(visual_trap_within, path = traps_file_xlsx)
print(paste("  ✓ Saved: within_subjects/visual_trap_within.xlsx"))

# ============================================================================
# METADATA & DOCUMENTATION
# ============================================================================
print("\n📄 Creating dataset metadata...")

dataset_metadata <- data.frame(
  Dataset = c(
    "ques_study (Between-subjects)",
    "ques_study_within_vignettes (Within-vignettes)",
    "visual_trap_within (Within-traps)"
  ),
  Format = c("Wide", "Long", "Long"),
  Rows = c(
    nrow(ques_study),
    nrow(ques_study_within_vignettes),
    nrow(visual_trap_within)
  ),
  Columns = c(
    ncol(ques_study),
    ncol(ques_study_within_vignettes),
    ncol(visual_trap_within)
  ),
  Unique_IDs = c(
    n_ids_wide,
    n_ids_vignettes,
    n_ids_traps
  ),
  Primary_Keys = c(
    "ID, PROLIFIC_PID",
    "ID, PROLIFIC_PID, counterScenarios",
    "ID, PROLIFIC_PID, counterVisualTraps, trap_type"
  ),
  Key_Variables = c(
    "Demographics, Scales (52 items), Transcripts, Sentiment",
    "PE (4 items), Attributes (6 items), Trial metadata, Paradata",
    "Response, Correctness, RT (reaction time), Trap type"
  ),
  Trials_Per_Participant = c(
    "1 row per participant",
    paste(min(trial_dist), "-", max(trial_dist), "trials (median:", median(trial_dist), ")"),
    paste(min(trap_dist), "-", max(trap_dist), "trials (median:", median(trap_dist), ")")
  )
)

metadata_file <- file.path(output_dir, "dataset_metadata.csv")
write.csv(dataset_metadata, file = metadata_file, row.names = FALSE)
print(paste("  ✓ Saved: dataset_metadata.csv"))

# ============================================================================
# CODEBOOK SUMMARY
# ============================================================================
print("\n📚 Creating codebook summary...")

codebook_summary <- data.frame(
  Section = c(
    "Identifiers",
    "Demographics",
    "Psychological Scales",
    "Transcripts",
    "Sentiment Analysis",
    "Vignette Variables (within)",
    "Trap Variables (within)"
  ),
  Variables = c(
    "ID, PROLIFIC_PID",
    paste(socio_demo, collapse = ", "),
    "52 items across 10 scales (BRS, SWLS, ARIS, NFC, SJS, etc.)",
    "text_missingUtopia, text_rankingUtopias",
    "sentiment_label, sentiment_score, num_windows, window_*_sentimentLabel",
    "vignette_order_position, vignette_id, condition_FutureSociety, pe1-pe4, att_ideal, att_beneficial, etc.",
    "trap_type, response, response_label, correct, rt_ms"
  ),
  Dataset_Location = c(
    "All datasets",
    "ques_study (wide only)",
    "ques_study (wide only)",
    "ques_study (wide only)",
    "ques_study (wide only)",
    "ques_study_within_vignettes",
    "visual_trap_within"
  ),
  Notes = c(
    "Integer ID and Prolific participant ID",
    "7 socio-demographic variables",
    "See scale documentation for item wording and scoring",
    "Text transcripts from audio recordings",
    "Sentiment labels (positive/negative/neutral) and scores",
    "One row per vignette trial, aggregated from long raw data",
    "One row per visual trap trial, responses and performance"
  )
)

codebook_file <- file.path(output_dir, "codebook_summary.csv")
write.csv(codebook_summary, file = codebook_file, row.names = FALSE)
print(paste("  ✓ Saved: codebook_summary.csv"))

# ============================================================================
# FINAL SUMMARY
# ============================================================================
print("✅ DATA EXPORT COMPLETE")
print(strrep("=", 80))


print("\n📋 BETWEEN-SUBJECTS (1 row per participant):")
print(paste("  • ques_study_wide.rds    ", round(file.size(ques_study_file_rds)/1024/1024, 2), "MB"))
print(paste("  • ques_study_wide.csv    ", round(file.size(ques_study_file_csv)/1024/1024, 2), "MB"))
print(paste("  • ques_study_wide.xlsx   ", round(file.size(ques_study_file_xlsx)/1024/1024, 2), "MB"))

print("\n📊 WITHIN-SUBJECTS VIGNETTES (one row per vignette trial):")
print(paste("  • ques_study_within_vignettes.rds   ", round(file.size(vignettes_file_rds)/1024/1024, 2), "MB"))
print(paste("  • ques_study_within_vignettes.csv   ", round(file.size(vignettes_file_csv)/1024/1024, 2), "MB"))
print(paste("  • ques_study_within_vignettes.xlsx  ", round(file.size(vignettes_file_xlsx)/1024/1024, 2), "MB"))

print("\n🎯 WITHIN-SUBJECTS VISUAL TRAPS (one row per trap trial):")
print(paste("  • visual_trap_within.rds   ", round(file.size(traps_file_rds)/1024/1024, 2), "MB"))
print(paste("  • visual_trap_within.csv   ", round(file.size(traps_file_csv)/1024/1024, 2), "MB"))
print(paste("  • visual_trap_within.xlsx  ", round(file.size(traps_file_xlsx)/1024/1024, 2), "MB"))

print("\n📄 DOCUMENTATION:")
print("  • dataset_metadata.csv")
print("  • codebook_summary.csv")


# Clean up
rm(dataset_metadata, codebook_summary, missing_wide)
gc()

print("\n✓ All done!")
```


```{r}
# ============================================================================
# SAVE FILTERING STATISTICS
# ============================================================================
print("\n📊 Saving filtering statistics...")

# Save filtering_stats as CSV
filtering_stats_file <- file.path(OUTPUT_QUESTIONNAIRE_DIR, "filtering_statistics.csv")
write.csv(filtering_stats, file = filtering_stats_file, row.names = FALSE)
print(paste("  ✓ filtering_statistics.csv"))

# Print summary
print("\n📋 FILTERING STATISTICS SUMMARY:")
print(filtering_stats)
```

# tests


```{r}
library(tidyverse)
library(corrplot)

# Create correlation matrix
ranking_cols <- c("rank_religious", "rank_ai_centered", "rank_primitivist", 
                  "rank_institutional", "rank_moral_anarchic", "rank_futurist", 
                  "rank_modern_green")
demo_cols <- c("lrscale", "rlgdgr")

# Calculate correlations
rank_demo_corr <- ques_study %>%
  select(all_of(c(ranking_cols, demo_cols))) %>%
  cor(use = "complete.obs")

# Extract just the rankings vs demographics
rank_vs_demo <- rank_demo_corr[ranking_cols, demo_cols]

# Visualize
corrplot(rank_vs_demo, method = "circle", type = "full", 
         diag = FALSE, tl.col = "black", tl.srt = 45,
         col = colorRampPalette(c("blue", "white", "red"))(200),
         main = "Utopia Rankings vs Political & Religious Orientation")

# Print correlation values
round(rank_vs_demo, 3)

# ============================================================================
# 2. MEAN RANKINGS by Demographics Groups
# ============================================================================

# Group by political orientation quartiles
ques_study_ranked <- ques_study %>%
  mutate(
    lrscale_group = case_when(
      lrscale <= 3 ~ "Left",
      lrscale <= 5 ~ "Center-Left",
      lrscale <= 7 ~ "Center-Right",
      TRUE ~ "Right"
    ),
    rlgdgr_group = case_when(
      rlgdgr <= 2 ~ "Non-religious",
      rlgdgr <= 4 ~ "Somewhat religious",
      TRUE ~ "Very religious"
    )
  )

# Mean rankings by political orientation
rank_by_lrscale <- ques_study_ranked %>%
  group_by(lrscale_group) %>%
  summarise(
    across(all_of(ranking_cols), 
           list(mean = ~ mean(., na.rm = TRUE)), 
           .names = "{.col}_mean"),
    n = n()
  )

print(rank_by_lrscale)

# Mean rankings by religiosity
rank_by_rlgdgr <- ques_study_ranked %>%
  group_by(rlgdgr_group) %>%
  summarise(
    across(all_of(ranking_cols), 
           list(mean = ~ mean(., na.rm = TRUE)), 
           .names = "{.col}_mean"),
    n = n()
  )

print(rank_by_rlgdgr)

```

```{r}
ggplot(ques_study, aes(x = lrscale, y = rank_moral_anarchic)) +
  geom_jitter(width = 0.15, height = 0.3, alpha = 0.5, size = 2.5, color = "#0C447C") +
  geom_smooth(method = "lm", se = TRUE, color = "#185FA5", fill = "#B5D4F4", linewidth = 1) +
  scale_x_continuous(breaks = 1:7, labels = c("Far Left", "2", "3", "Center", "5", "6", "Far Right")) +
  scale_y_continuous(breaks = 1:7, labels = c("1\n(least)", "2", "3", "4", "5", "6", "7\n(most)")) +
  labs(
    title = "Political Orientation vs Moral Commonwealth Preference",
    subtitle = "Left-leaning participants strongly prefer self-regulating moral societies",
    x = "Political Orientation (Left ← → Right)",
    y = "Ranking: Moral Commonwealth Utopia"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11, color = "#666"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#eee")
  )


ques_study %>%
  mutate(
    political_group = case_when(
      lrscale <= 2 ~ "Far Left",
      lrscale <= 3 ~ "Left",
      lrscale <= 4 ~ "Center-Left",
      lrscale == 5 ~ "Center",
      lrscale <= 6 ~ "Center-Right",
      lrscale <= 7 ~ "Right",
      TRUE ~ "Far Right"
    ),
    political_group = factor(political_group, 
                            levels = c("Far Left", "Left", "Center-Left", "Center", 
                                     "Center-Right", "Right", "Far Right"))
  ) %>%
  ggplot(aes(x = political_group, y = rank_moral_anarchic, fill = political_group)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.3, size = 1.5) +
  scale_fill_manual(values = c(
    "Far Left" = "#042C53",
    "Left" = "#185FA5",
    "Center-Left" = "#85B7EB",
    "Center" = "#D3D1C7",
    "Center-Right" = "#F0997B",
    "Right" = "#993C1D",
    "Far Right" = "#4A1B0C"
  )) +
  scale_y_continuous(breaks = 1:7) +
  labs(
    title = "Moral Commonwealth Ranking by Political Group",
    x = "Political Orientation",
    y = "Ranking (1=least preferred, 7=most preferred)",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14)
  )

```



```{r}
# ============================================================================
# 1. SCATTER PLOT with jitter, trend, and marginal distributions
# ============================================================================

ggplot(ques_study, aes(x = rlgdgr, y = rank_religious)) +
  geom_jitter(width = 0.15, height = 0.3, alpha = 0.5, size = 2.5, color = "#A32D2D") +
  geom_smooth(method = "lm", se = TRUE, color = "#993C1D", fill = "#F0997B", linewidth = 1) +
  scale_x_continuous(breaks = 1:7, labels = c("Not Religious", "2", "3", "Moderate", "5", "6", "Very Religious")) +
  scale_y_continuous(breaks = 1:7, labels = c("1\n(least)", "2", "3", "4", "5", "6", "7\n(most)")) +
  labs(
    title = "Religious Commitment vs Religious Utopia Preference",
    subtitle = "Very religious participants strongly prefer faith-based future societies",
    x = "Religious Commitment (Not Religious ← → Very Religious)",
    y = "Ranking: Religious Utopia"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11, color = "#666"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#eee")
  )

# ============================================================================
# 2. BOX PLOT (by religiosity groups)
# ============================================================================

ques_study %>%
  mutate(
    religion_group = case_when(
      rlgdgr <= 2 ~ "Not Religious",
      rlgdgr <= 3 ~ "Slightly Religious",
      rlgdgr <= 4 ~ "Somewhat Religious",
      rlgdgr == 5 ~ "Moderate",
      rlgdgr <= 6 ~ "Quite Religious",
      rlgdgr <= 7 ~ "Very Religious",
      TRUE ~ "Unknown"
    ),
    religion_group = factor(religion_group, 
                           levels = c("Not Religious", "Slightly Religious", "Somewhat Religious", 
                                    "Moderate", "Quite Religious", "Very Religious"))
  ) %>%
  ggplot(aes(x = religion_group, y = rank_religious, fill = religion_group)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.3, size = 1.5) +
  scale_fill_manual(values = c(
    "Not Religious" = "#042C53",
    "Slightly Religious" = "#185FA5",
    "Somewhat Religious" = "#85B7EB",
    "Moderate" = "#D3D1C7",
    "Quite Religious" = "#F0997B",
    "Very Religious" = "#993C1D"
  )) +
  scale_y_continuous(breaks = 1:7) +
  labs(
    title = "Religious Utopia Ranking by Religious Commitment",
    x = "Religious Commitment",
    y = "Ranking (1=least preferred, 7=most preferred)",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14)
  )

```




```{r}
# Create faceted plot for all rankings
plot_data <- ques_study %>%
  select(lrscale, all_of(ranking_cols)) %>%
  pivot_longer(cols = -lrscale, names_to = "utopia", values_to = "ranking") %>%
  mutate(
    utopia = case_when(
      utopia == "rank_religious" ~ "Religious",
      utopia == "rank_ai_centered" ~ "AI-Centered",
      utopia == "rank_primitivist" ~ "Primitivist",
      utopia == "rank_institutional" ~ "Institutional",
      utopia == "rank_moral_anarchic" ~ "Moral Commonwealth",
      utopia == "rank_futurist" ~ "Futurist",
      utopia == "rank_modern_green" ~ "Modern Green",
      TRUE ~ utopia
    )
  )

ggplot(plot_data, aes(x = lrscale, y = ranking)) +
  geom_jitter(width = 0.15, height = 0.2, alpha = 0.4, size = 2, color = "#185FA5") +
  geom_smooth(method = "lm", se = TRUE, color = "#042C53", fill = "#B5D4F4", linewidth = 0.8) +
  facet_wrap(~ utopia, ncol = 3) +
  scale_x_continuous(breaks = c(1, 4, 7), labels = c("Left", "Center", "Right")) +
  scale_y_continuous(breaks = c(1, 4, 7)) +
  labs(
    title = "Political Orientation vs Each Utopia Ranking",
    x = "Political Orientation",
    y = "Ranking (1=least, 7=most preferred)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    strip.text = element_text(face = "bold", size = 10),
    panel.grid.minor = element_blank()
  )
```


```{r}
# Create faceted plot for all rankings
plot_data <- ques_study %>%
  select(rlgdgr, all_of(ranking_cols)) %>%
  pivot_longer(cols = -rlgdgr, names_to = "utopia", values_to = "ranking") %>%
  mutate(
    utopia = case_when(
      utopia == "rank_religious" ~ "Religious",
      utopia == "rank_ai_centered" ~ "AI-Centered",
      utopia == "rank_primitivist" ~ "Primitivist",
      utopia == "rank_institutional" ~ "Institutional",
      utopia == "rank_moral_anarchic" ~ "Moral Commonwealth",
      utopia == "rank_futurist" ~ "Futurist",
      utopia == "rank_modern_green" ~ "Modern Green",
      TRUE ~ utopia
    )
  )

ggplot(plot_data, aes(x = rlgdgr, y = ranking)) +
  geom_jitter(width = 0.15, height = 0.2, alpha = 0.4, size = 2, color = "#185FA5") +
  geom_smooth(method = "lm", se = TRUE, color = "#042C53", fill = "#B5D4F4", linewidth = 0.8) +
  facet_wrap(~ utopia, ncol = 3) +
  labs(
    title = "Religiousness vs Each Utopia Ranking",
    x = "Religiousness Orientation",
    y = "Ranking (1=least, 7=most preferred)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    strip.text = element_text(face = "bold", size = 10),
    panel.grid.minor = element_blank()
  )
```


## within part

```{r}
library(tidyverse)
library(afex)
library(ggplot2)
 
# Assumption: ques_study_within_vignettes is already loaded
 ques_study_within_vignettes <- ques_study_within_vignettes %>%
  mutate(
    # Convert PE items to numeric
    pe1 = as.numeric(pe1),
    pe2 = as.numeric(pe2),
    pe3 = as.numeric(pe3),
    pe4 = as.numeric(pe4),
    # Convert ATT items to numeric
    att_ideal = as.numeric(att_ideal),
    att_beneficial = as.numeric(att_beneficial),
    att_utopian = as.numeric(att_utopian),
    att_desirable = as.numeric(att_desirable),
    att_innovative = as.numeric(att_innovative),
    att_live_in = as.numeric(att_live_in)
  )
```


```{r}
# ============================================================================
# PART 1: PREPARE DATA FOR ANALYSIS
# ============================================================================
 
analysis_vignettes <- ques_study_within_vignettes %>%
  # Calculate composite PE score (mean of pe1-pe4)
  mutate(
    pe_composite = rowMeans(select(., pe1:pe4), na.rm = TRUE)
  ) %>%
  # Calculate composite ATT score (mean of att_ideal through att_live_in)
  mutate(
    att_composite = rowMeans(select(., att_ideal, att_beneficial, att_utopian, 
                                      att_desirable, att_innovative, att_live_in), 
                             na.rm = TRUE)
  ) %>%
  # Keep only complete cases for main analysis
  filter(!is.na(pe_composite), !is.na(att_composite), !is.na(condition_FutureSociety))
 
print(paste("Analysis sample size:", nrow(analysis_vignettes), "vignette trials"))
print(paste("Unique participants:", n_distinct(analysis_vignettes$ID)))
print(paste("Unique vignette conditions:", paste(unique(analysis_vignettes$condition_FutureSociety), collapse = ", ")))
print(paste("Trials per participant:", paste(table(analysis_vignettes$ID), collapse = ", ")))
 
 
# ============================================================================
# PART 2: DESCRIPTIVE STATISTICS BY CONDITION
# ============================================================================
 
desc_stats <- analysis_vignettes %>%
  pivot_longer(
    cols = c(pe_composite, att_composite),
    names_to = "measure",
    values_to = "value"
  ) %>%
  group_by(condition_FutureSociety, measure) %>%
  summarise(
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = dplyr::n(),
    se = sd / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(
    measure = recode(
      measure,
      pe_composite = "Participative Efficacy",
      att_composite = "Attributes (Positive)"
    )
  )
 
print("\n=== DESCRIPTIVE STATISTICS ===")
print(desc_stats)
 
 
# ============================================================================
# PART 3: WITHIN-SUBJECTS ANOVA ANALYSES
# ============================================================================
 
# ANOVA 1: Participative Efficacy by Condition (within-subjects)
print("\n=== ANOVA 1: PARTICIPATIVE EFFICACY (Within-Subjects) ===")
anova_pe <- afex::aov_ez(
  id = "ID",
  dv = "pe_composite",
  within = "condition_FutureSociety",
  data = analysis_vignettes
)
print(summary(anova_pe))
 
 
# ANOVA 2: Attributes by Condition (within-subjects)
print("\n=== ANOVA 2: ATTRIBUTES COMPOSITE (Within-Subjects) ===")
anova_att <- afex::aov_ez(
  id = "ID",
  dv = "att_composite",
  within = "condition_FutureSociety",
  data = analysis_vignettes
)
print(summary(anova_att))
```



```{r}
# ============================================================================
# ATTRIBUTE ITEMS ANALYSIS (att_ideal, att_beneficial, att_utopian, etc.)
# ============================================================================

# Attribute item label mappings from survey
attr_item_labels <- tibble(
  item_label = c("ideal", "beneficial", "utopian", "desirable", "innovative", "live_in"),
  full_label = c(
    "This society is ideal.",
    "This society is beneficial for the greater good.",
    "This society is utopian.",
    "This society is desirable.",
    "This society is innovative.",
    "I would like to live in a society like the one just described."
  )
)

# Convert attitude variables to long format
attr_long <- ques_study_within_vignettes %>%
  select(ID, condition_FutureSociety, 
         att_ideal, att_beneficial, att_utopian, att_desirable, att_innovative, att_live_in) %>%
  pivot_longer(
    cols = starts_with("att_"),
    names_to = "item_label",
    values_to = "value"
  ) %>%
  mutate(item_label = str_remove(item_label, "^att_"))

# Calculate effect sizes (eta-squared) for each attribute item
attr_effects <- attr_long %>%
  group_by(item_label) %>%
  summarise(
    # Calculate eta-squared for one-way ANOVA
    ss_between = sum(tapply(value, condition_FutureSociety, length) * 
                       (tapply(value, condition_FutureSociety, mean) - mean(value, na.rm=T))^2, na.rm=T),
    ss_total = sum((value - mean(value, na.rm=T))^2, na.rm=T),
    eta_sq = ss_between / ss_total,
    .groups = "drop"
  ) %>%
  left_join(attr_item_labels, by = "item_label") %>%
  mutate(eta_label = paste0("η² = ", round(eta_sq, 3)))

# Order items by effect size (largest first)
item_order <- attr_effects %>%
  mutate(eta_order = ifelse(is.na(eta_sq), -Inf, eta_sq)) %>%
  arrange(desc(eta_order)) %>%
  pull(item_label)

# Create labels with full item text and eta-squared values
label_order <- attr_effects %>%
  mutate(eta_order = ifelse(is.na(eta_sq), -Inf, eta_sq)) %>%
  arrange(desc(eta_order)) %>%
  mutate(facet_label = paste0(full_label, "\n", eta_label)) %>%
  pull(facet_label)

# Summarize by condition and item
attr_summary <- attr_long %>%
  group_by(condition_FutureSociety, item_label) %>%
  summarise(
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = dplyr::n(),
    se = sd / sqrt(n),
    .groups = "drop"
  ) %>%
  left_join(attr_effects %>% select(item_label, full_label, eta_label), by = "item_label") %>%
  mutate(
    item_label = factor(item_label, levels = item_order),
    facet_label = factor(paste0(full_label, "\n", eta_label), levels = label_order)
  )

# Order conditions alphabetically
condition_order <- attr_long %>%
  distinct(condition_FutureSociety) %>%
  arrange(condition_FutureSociety) %>%
  pull(condition_FutureSociety)

attr_summary <- attr_summary %>%
  mutate(condition_FutureSociety = factor(condition_FutureSociety, levels = condition_order))

# Visualization - Attributes (with full item labels)
ggplot(attr_summary, aes(x = condition_FutureSociety, y = mean, fill = condition_FutureSociety)) +
  geom_col(width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  geom_text(aes(label = round(mean, 2)), vjust = -0.3, size = 3) +
  facet_wrap(~ facet_label, ncol = 3) +
  labs(x = "Condition", y = "Mean Rating", fill = "Condition",
       title = "Attributes: How Participants Evaluated Each Society") +
  theme_minimal() +
  theme(
    legend.position = "bottom", 
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 7.5, face = "italic"),
    plot.title = element_text(hjust = 0.5, size = 11),
    panel.spacing = unit(1, "lines")
  )
```




```{r}
# ============================================================================
# PERCEIVED EFFICACY ITEMS ANALYSIS (pe1, pe2, pe3, pe4)
# ============================================================================

# Item label mappings from survey
pe_item_labels <- tibble(
  item_label = c("pe1", "pe2", "pe3", "pe4"),
  full_label = c(
    "Ordinary people like me can play a part in bringing about this kind of society.",
    "This kind of society can be realized through the actions of ordinary individuals like me.",
    "Ordinary people can help realize this version of society.",
    "Ordinary people are needed to realize this version of society."
  )
)

# Convert perceived efficacy variables to long format
pe_long <- ques_study_within_vignettes %>%
  select(ID, condition_FutureSociety, 
         pe1, pe2, pe3, pe4) %>%
  pivot_longer(
    cols = starts_with("pe"),
    names_to = "item_label",
    values_to = "value"
  )

# Calculate effect sizes (eta-squared) for each perceived efficacy item
pe_effects <- pe_long %>%
  group_by(item_label) %>%
  summarise(
    # Calculate eta-squared for one-way ANOVA
    ss_between = sum(tapply(value, condition_FutureSociety, length) * 
                       (tapply(value, condition_FutureSociety, mean) - mean(value, na.rm=T))^2, na.rm=T),
    ss_total = sum((value - mean(value, na.rm=T))^2, na.rm=T),
    eta_sq = ss_between / ss_total,
    .groups = "drop"
  ) %>%
  left_join(pe_item_labels, by = "item_label") %>%
  mutate(eta_label = paste0("η² = ", round(eta_sq, 3)))

# Order items by effect size (largest first)
item_order <- pe_effects %>%
  mutate(eta_order = ifelse(is.na(eta_sq), -Inf, eta_sq)) %>%
  arrange(desc(eta_order)) %>%
  pull(item_label)

# Create labels with full item text and eta-squared values
label_order <- pe_effects %>%
  mutate(eta_order = ifelse(is.na(eta_sq), -Inf, eta_sq)) %>%
  arrange(desc(eta_order)) %>%
  mutate(facet_label = paste0(full_label, "\n", eta_label)) %>%
  pull(facet_label)

# Summarize by condition and item
pe_summary <- pe_long %>%
  group_by(condition_FutureSociety, item_label) %>%
  summarise(
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = dplyr::n(),
    se = sd / sqrt(n),
    .groups = "drop"
  ) %>%
  left_join(pe_effects %>% select(item_label, full_label, eta_label), by = "item_label") %>%
  mutate(
    item_label = factor(item_label, levels = item_order),
    facet_label = factor(paste0(full_label, "\n", eta_label), levels = label_order)
  )

# Order conditions alphabetically
condition_order <- pe_long %>%
  distinct(condition_FutureSociety) %>%
  arrange(condition_FutureSociety) %>%
  pull(condition_FutureSociety)

pe_summary <- pe_summary %>%
  mutate(condition_FutureSociety = factor(condition_FutureSociety, levels = condition_order))

# Visualization - Perceived Efficacy (with full item labels)
ggplot(pe_summary, aes(x = condition_FutureSociety, y = mean, fill = condition_FutureSociety)) +
  geom_col(width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  geom_text(aes(label = round(mean, 2)), vjust = -0.3, size = 3) +
  facet_wrap(~ facet_label, ncol = 2) +
  labs(x = "Condition", y = "Mean Rating", fill = "Condition",
       title = "Perceived Efficacy: Ordinary Citizens Can Contribute to Realizing Society") +
  theme_minimal() +
  theme(
    legend.position = "bottom", 
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 7.5, face = "italic"),
    plot.title = element_text(hjust = 0.5, size = 11),
    panel.spacing = unit(1, "lines")
  )
```


## EFA

```{r}
library(tidyverse)
library(corrplot)
library(psych)
library(factoextra)

# ============================================================================
# PREPARE DATA: Convert from long to wide format
# ============================================================================

# Create wide format with all items for each observation
data_wide <- ques_study_within_vignettes %>%
  select(ID, condition_FutureSociety, vignette_id,
         att_ideal, att_beneficial, att_utopian, att_desirable, att_innovative, att_live_in,
         pe1, pe2, pe3, pe4)

# ============================================================================
# 1. CORRELATION MATRIX & VISUALIZATION
# ============================================================================

# Extract only the items
items_data <- data_wide %>%
  select(att_ideal, att_beneficial, att_utopian, att_desirable, att_innovative, att_live_in,
         pe1, pe2, pe3, pe4)

# Calculate correlation matrix
cor_matrix <- cor(items_data, use = "complete.obs")

# Print correlation matrix
print("Correlation Matrix:")
print(cor_matrix)

# Create correlation plot
pdf("correlation_plot.pdf", width = 10, height = 10)
corrplot(cor_matrix, 
         method = "circle",
         type = "lower",
         addCoef.col = "black",
         number.cex = 0.7,
         tl.cex = 0.8,
         diag = TRUE,
         title = "Correlation Matrix: Attitude and Perceived Efficacy Items",
         mar = c(2, 0, 2, 0))
dev.off()

# Alternative: Circle plot with significance
corrplot(cor_matrix,
         method = "shade",
         type = "lower",
         addCoef.col = "white",
         tl.cex = 0.8,
         title = "Correlation Matrix (Shade)",
         mar = c(2, 0, 2, 0))

# ============================================================================
# 2. PARALLEL ANALYSIS (to determine number of factors)
# ============================================================================

print("\n========== PARALLEL ANALYSIS ==========\n")

# Run parallel analysis
pa <- fa.parallel(items_data, 
                  fm = "minres",           # Factor extraction method
                  fa = "fa",               # Factor analysis (not principal components)
                  n.iter = 100,
                  show.legend = TRUE)

# Print summary
print(pa)

# ============================================================================
# 3. FACTOR ANALYSIS - Test different numbers of factors
# ============================================================================

print("\n========== FACTOR ANALYSIS RESULTS ==========\n")

# 1-Factor Solution
fa1 <- fa(items_data, nfactors = 1, fm = "minres", rotate = "promax")
print("\n--- 1-Factor Solution ---")
print(fa1)

# 2-Factor Solution
fa2 <- fa(items_data, nfactors = 2, fm = "minres", rotate = "promax")
print("\n--- 2-Factor Solution ---")
print(fa2)

# 3-Factor Solution
fa3 <- fa(items_data, nfactors = 3, fm = "minres", rotate = "promax")
print("\n--- 3-Factor Solution ---")
print(fa3)

# ============================================================================
# 4. FACTOR LOADINGS COMPARISON
# ============================================================================

print("\n========== FACTOR LOADINGS COMPARISON ==========\n")

# Extract loadings for comparison
loadings_df <- data.frame(
  Item = rownames(fa2$loadings),
  FA1 = fa2$loadings[, 1],
  FA2 = fa2$loadings[, 2],
  Communality = fa2$communalities
)

print(loadings_df)

# ============================================================================
# 5. VISUALIZE FACTOR STRUCTURE (2-Factor Solution)
# ============================================================================

# Plot factor loadings
pdf("factor_loadings.pdf", width = 10, height = 8)
plot(fa2$loadings[, 1], fa2$loadings[, 2],
     type = "n",
     xlab = "Factor 1 Loadings",
     ylab = "Factor 2 Loadings",
     main = "Factor Loadings Plot (2-Factor Solution)",
     xlim = c(-1, 1),
     ylim = c(-1, 1))

# Add reference lines
abline(h = 0, v = 0, lty = 2, col = "gray")

# Add labels
text(fa2$loadings[, 1], fa2$loadings[, 2], 
     rownames(fa2$loadings), 
     cex = 0.8,
     pos = 3)

# Add circles for item grouping
draw.circle <- function(x, y, r, ...) {
  angles <- seq(0, 2*pi, length.out = 100)
  lines(x + r*cos(angles), y + r*sin(angles), ...)
}

dev.off()

# ============================================================================
# 6. MODEL FIT COMPARISON
# ============================================================================

print("\n========== MODEL FIT INDICES ==========\n")

fit_comparison <- data.frame(
  Factors = c(1, 2, 3),
  RMSR = c(fa1$rms, fa2$rms, fa3$rms),
  BIC = c(fa1$BIC, fa2$BIC, fa3$BIC),
  RMSEA = c(fa1$RMSEA[1], fa2$RMSEA[1], fa3$RMSEA[1]),
  TLI = c(fa1$TLI, fa2$TLI, fa3$TLI)
)

print(fit_comparison)

# ============================================================================
# 7. INTERNAL CONSISTENCY (Cronbach's Alpha)
# ============================================================================

print("\n========== CRONBACH'S ALPHA - INTERNAL CONSISTENCY ==========\n")

# All items
alpha_all <- alpha(items_data)$total$raw_alpha
cat("All items (n=10):", round(alpha_all, 3), "\n")

# Attitude items only
alpha_att <- alpha(items_data %>% select(starts_with("att_")))$total$raw_alpha
cat("Attitude items (n=6):", round(alpha_att, 3), "\n")

# Perceived efficacy items only
alpha_pe <- alpha(items_data %>% select(starts_with("pe")))$total$raw_alpha
cat("Perceived Efficacy items (n=4):", round(alpha_pe, 3), "\n")

# ============================================================================
# 8. SCREE PLOT
# ============================================================================

pdf("scree_plot.pdf", width = 10, height = 6)
scree(items_data, factors = 5, main = "Scree Plot")
dev.off()

print("\nPlots saved as: correlation_plot.pdf, factor_loadings.pdf, scree_plot.pdf")
```



## prompts




```{r}
library(glue)

# ============================================================================
# 3. CREATE LLM VALIDATION PROMPT DATASET
# ============================================================================

# Build validation dataset with context for each participant
llm_validation_data <- ques_study %>%
  select(
    ID, PROLIFIC_PID, age, gender, education, lrscale, rlgdgr,
    all_of(ranking_cols),
    text_missingUtopia, text_rankingUtopias,
    sentiment_label, sentiment_score
  ) %>%
  filter(!is.na(text_rankingUtopias) | !is.na(text_missingUtopia)) %>%
  mutate(
    # Create ranking summary
    ranking_summary = paste(
      "1st (least preferred):", 
      case_when(rank_religious == 1 ~ "Religious",
                rank_ai_centered == 1 ~ "AI-Centered",
                rank_primitivist == 1 ~ "Primitivist",
                rank_institutional == 1 ~ "Institutional",
                rank_moral_anarchic == 1 ~ "Moral Commonwealth",
                rank_futurist == 1 ~ "Futurist",
                rank_modern_green == 1 ~ "Modern Green",
                TRUE ~ "Unknown"),
      "\n7th (most preferred):",
      case_when(rank_religious == 7 ~ "Religious",
                rank_ai_centered == 7 ~ "AI-Centered",
                rank_primitivist == 7 ~ "Primitivist",
                rank_institutional == 7 ~ "Institutional",
                rank_moral_anarchic == 7 ~ "Moral Commonwealth",
                rank_futurist == 7 ~ "Futurist",
                rank_modern_green == 7 ~ "Modern Green",
                TRUE ~ "Unknown"),
      sep = " "
    ),
    # Political orientation label
    political_orientation = case_when(
      lrscale <= 3 ~ "Left-leaning",
      lrscale <= 5 ~ "Center",
      lrscale >= 7 ~ "Right-leaning",
      TRUE ~ "Moderate"
    ),
    # Religiosity label
    religiosity = case_when(
      rlgdgr <= 2 ~ "Non-religious",
      rlgdgr <= 4 ~ "Somewhat religious",
      rlgdgr >= 6 ~ "Very religious",
      TRUE ~ "Moderate religiosity"
    )
  )

head(llm_validation_data)

# ============================================================================
# FIXED: CREATE LLM VALIDATION PROMPT FUNCTION
# ============================================================================

create_llm_validation_prompt <- function(id, data) {
  row <- data %>% filter(ID == id) %>% slice(1)
  
  if (nrow(row) == 0) return("")
  
  prompt <- glue(
    "
# Consistency Check: Participant {row$ID[[1]]}

## Demographics
- Age: {row$age[[1]]}
- Gender: {row$gender[[1]]}
- Political Orientation: {row$political_orientation[[1]]} (lrscale={row$lrscale[[1]]})
- Religiosity: {row$religiosity[[1]]} (rlgdgr={row$rlgdgr[[1]]})

## Rankings (1=least preferred, 7=most preferred)
- Religious Utopia: {row$rank_religious[[1]]}
- AI-Centered Utopia: {row$rank_ai_centered[[1]]}
- Primitivist Utopia: {row$rank_primitivist[[1]]}
- Institutional Utopia: {row$rank_institutional[[1]]}
- Moral Commonwealth: {row$rank_moral_anarchic[[1]]}
- Futurist Utopia: {row$rank_futurist[[1]]}
- Modern Green Utopia: {row$rank_modern_green[[1]]}

## Sentiment Analysis
- Overall sentiment: {row$sentiment_label[[1]]} (score={round(row$sentiment_score[[1]], 2)})

## Text Responses

### Ranking Explanation
\"{row$text_rankingUtopias[[1]]}\"

### Missing Utopia Description
\"{row$text_missingUtopia[[1]]}\"

## Task

Evaluate whether the participant's **text responses** are consistent with their **demographic profile** and **utopia rankings**. 

For each dimension, rate consistency (1-5, where 1=completely inconsistent, 5=highly consistent):

1. **Political-Ideological Consistency**: Do the text responses reflect their political orientation (left/center/right)?
2. **Religious-Philosophical Consistency**: Do the text responses align with their religiosity level?
3. **Ranking-Text Alignment**: Do the rankings they chose match what they describe in text?
4. **Overall Coherence**: Does the person's overall vision (across all responses) hang together logically?

Provide:
- Rating (1-5) for each dimension
- Brief explanation (1-2 sentences per dimension)
- Flag any red flags (contradictions, suspicious patterns, potential comprehension issues)

---
"
  )
  return(prompt)
}

# ============================================================================
# GENERATE PROMPTS FOR ALL PARTICIPANTS
# ============================================================================

all_prompts <- map_chr(
  llm_validation_data$ID,
  ~ create_llm_validation_prompt(., llm_validation_data)
)



# ============================================================================
# ALTERNATIVE: CREATE A DATAFRAME FOR BATCH LLM PROCESSING
# ============================================================================

# More efficient: keep data structured, send to LLM as CSV/JSON
llm_batch_data <- llm_validation_data %>%
  select(
    ID, PROLIFIC_PID, age, gender, 
    political_orientation, religiosity, lrscale, rlgdgr,
    rank_religious, rank_ai_centered, rank_primitivist, 
    rank_institutional, rank_moral_anarchic, rank_futurist, rank_modern_green,
    sentiment_label, sentiment_score,
    text_rankingUtopias, text_missingUtopia
  ) %>%
  rename(
    ranking_explanation = text_rankingUtopias,
    missing_utopia = text_missingUtopia
  )



# Print first 2 samples
cat(all_prompts[[1]])
cat("\n\n===================\n\n")
cat(all_prompts[[3]])

```

