---
title: "Data Preparation: Survey + Transcripts + Sentiment Analysis"
author: "Julius Fenn"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    html-math-method: katex
    number-sections: true
    embed-resources: true
---

# Overview

This document prepares the complete dataset by combining:
1. **Survey Data** from `data/results_survey/` (questionnaire responses)
2. **Audio Transcripts** from `outputs/audio_transcripts/` (text_missingUtopia, text_rankingUtopias)
3. **Sentiment Analysis** from `outputs/sentiment_analysis/` (overall + window-level)

Final output: One row per participant with all metadata integrated.

---

# Setup & Configuration

```{r}
#| label: setup
#| message: false
#| warning: false

# Clear environment
rm(list = ls())

# Set control variables
createRawFiles <- TRUE  # Set to FALSE to skip raw file creation
VERBOSE <- TRUE

# Load packages
require(pacman)
p_load('tidyverse', 'jsonlite', 'writexl', 'lubridate', 'DT', 'psych')

# Set base directory
BASE_DIR <- getwd()
print(paste("Working directory:", BASE_DIR))

# Define paths
DATA_SURVEY_DIR <- file.path(BASE_DIR, "data", "results_survey")
AUDIO_TRANSCRIPT_DIR <- file.path(BASE_DIR, "outputs", "audio_transcripts")
SENTIMENT_DIR <- file.path(BASE_DIR, "outputs", "sentiment_analysis")

OUTPUT_RAW_DIR <- file.path(BASE_DIR, "outputs", "raw")
OUTPUT_QUESTIONNAIRE_DIR <- file.path(BASE_DIR, "outputs", "questionnaire")

# Create output directories if they don't exist
dir.create(OUTPUT_RAW_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(OUTPUT_QUESTIONNAIRE_DIR, showWarnings = FALSE, recursive = TRUE)

print("✓ Setup complete")
```

---

# Helper Functions

## Questionnaire Type Function

```{r}
#| label: questionnaire-function

questionnairetype <- function(dataset,
                              listvars = ques_mixed,
                              notNumeric = vec_notNumeric,
                              verbose = FALSE) {

  # Convert JATOS data to classical 2D data frame
  # 
  # Args:
  #   dataset: Raw data frame from JATOS
  #   listvars: Vector of column names to extract
  #   notNumeric: Vector of column names to keep as character
  #   verbose: Print progress
  # 
  # Returns:
  #   Data frame with 1 row per unique ID

  
  datasetques <- data.frame(ID = unique(dataset$ID))
  
  for (c in 1:length(listvars)) {
    if (verbose) {
      print(c)
    }
    
    if (any(colnames(dataset) == listvars[c])) {
      if (verbose) {
        print(listvars[c])
      }
      
      ## tmp IDs
      tmpid <- dataset$ID[!is.na(dataset[, listvars[c]])]
      ## tmp value variable
      tmpvalue <- dataset[, listvars[c]][!is.na(dataset[, listvars[c]])]
      datasetques[listvars[c]] <- NA
      
      if (listvars[c] %in% notNumeric) {
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- tmpvalue
      } else if (is.list(tmpvalue)) {
        tmpvalue_tmp <- unique(tmpvalue)
        tmpvalue <- c()
        counter <- 1
        for (i in 1:length(tmpvalue_tmp)) {
          if (!is.null(tmpvalue_tmp[[i]])) {
            tmpvalue[counter] <- paste0(tmpvalue_tmp[[i]], collapse = " - ")
            counter <- counter + 1
          }
        }
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- tmpvalue
      } else {
        datasetques[datasetques$ID %in% tmpid, listvars[c]] <- as.numeric(tmpvalue)
      }
    }
  }
  return(datasetques)
}

print("✓ Helper functions defined")
```

---

# Load & Process Raw Survey Data

## Load all data.txt files

```{r}
#| label: load-raw-survey-data
#| message: false

if (createRawFiles) {
  print("Loading raw survey data from data.txt files...")
  
  # Find all data.txt files
  data_files <- list.files(
    path = DATA_SURVEY_DIR,
    pattern = "^data\\.txt$",
    full.names = TRUE,
    recursive = TRUE
  )
  
  print(paste("Found", length(data_files), "survey data files"))
  
  if (VERBOSE) {
    print("Files found:")
    print(data_files)
  }
  
  # Load all files
  results <- list()
  
  for (file in data_files) {
    tryCatch({
      tmp <- fromJSON(file)
      tmp[, colnames(tmp) == ""] <- NULL # remove empty variable
      results[[length(results) + 1]] <- tmp
      
      if (VERBOSE) {
        print(paste("✓ Loaded:", basename(dirname(dirname(file)))))
      }
    }, error = function(e) {
      print(paste("✗ Error loading", file, ":", e$message))
    })
  }
  
  print(paste("Successfully loaded", length(results), "files"))
  
  # Save as RDS
  saveRDS(results, file = file.path(OUTPUT_RAW_DIR, "study.rds"))
  
  # Optionally save as JSONL
  json_lines <- map_chr(results, ~ toJSON(.x, auto_unbox = TRUE))
  writeLines(json_lines, con = file.path(OUTPUT_RAW_DIR, "study.jsonl"))
  
  print("✓ Raw files saved (study.rds and study.jsonl)")
  
} else {
  print("Loading pre-saved raw data...")
  results <- readRDS(file = file.path(OUTPUT_RAW_DIR, "study.rds"))
  print(paste("Loaded", length(results), "records from study.rds"))
}
```

---

# Create Questionnaire DataFrame with Filtering


## Combine and prepare raw data

```{r}
#| label: combine-raw-data

for(i in 1:length(results)) {
  tmp <- results[[i]]$meta
  results[[i]]$meta <- NULL
  
    if(any(colnames(results[[i]]) %in% "paradata_general")) {
      results[[i]]$paradata_general <- NULL
    }
  
  results[[i]] <- cbind(results[[i]], tmp)
}


# Combine all records into single dataframe
suppressMessages({
  dat <- bind_rows(results)
})


# Combine all records into single dataframe
suppressMessages({
  dat <- bind_rows(results)
})

print(paste("Combined data shape:", nrow(dat), "rows x", ncol(dat), "columns"))
rm(results)

# Add ID counter
dat$ID <- cumsum(dat$sender == "Greetings" & !is.na(dat$sender))

print("\n=== Initial ID Distribution ===")
print(table(dat$ID))

# Check record counts per ID
tmp <- dat %>%
  group_by(ID) %>%
  summarise(N = n(), .groups = 'drop')

print("\n=== Record Count Distribution ===")
print(table(tmp$N))

max_N <- max(tmp$N)
print(paste("Maximum records per ID:", max_N))
```


## Identify incomplete records and test data

```{r}
#| label: filter-incomplete-records

# Initialize filtering tracking
filtering_stats <- data.frame(
  metric = character(),
  value = numeric(),
  stringsAsFactors = FALSE
)

starting_N <- nrow(dat)
filtering_stats <- rbind(filtering_stats, 
                         data.frame(metric = "starting_total_records", value = starting_N))

# Find IDs with incomplete records
tmp_full <- dat %>%
  group_by(ID) %>%
  summarise(N = n(), .groups = 'drop')

incomplete_IDs <- tmp_full$ID[tmp_full$N != max_N]
n_incomplete_before <- length(incomplete_IDs)

print(paste("\n=== Filtering Step 1: Identify Incomplete Records ==="))
print(paste("IDs with fewer than", max_N, "records:", n_incomplete_before))
if (VERBOSE & n_incomplete_before > 0) {
  print("Incomplete IDs:")
  print(incomplete_IDs)
  
  filtering_stats <- rbind(filtering_stats, 
                         data.frame(metric = "incomplete_ids", value = paste0(incomplete_IDs, collapse = " // ")))
}

dat <- dat[!dat$ID %in% incomplete_IDs, ]

if (VERBOSE & n_incomplete_before > 0) {
  filtering_stats <- rbind(filtering_stats, 
                         data.frame(metric = "after_filtering_incomplete_records", value = nrow(dat)))
}



# Handle test ID "111" - Special logic
print("\n=== Filtering Step 2: Handle Test ID '111' ===")

test_ids <- dat$ID[dat$PROLIFIC_PID == "111"] # !!!
test_ids <- test_ids[!is.na(test_ids)]
n_test_runs <- length(unique(test_ids))

print(paste("Found", n_test_runs, "test run(s) with ID '111'"))

if (VERBOSE & n_test_runs > 0) {
  print("Test IDs:")
  print(test_ids)
  
  filtering_stats <- rbind(filtering_stats, 
                         data.frame(metric = "test_ids", value = paste0(test_ids, collapse = " // ")))
}

dat <- dat[!dat$ID %in% test_ids, ]


if (VERBOSE & n_test_runs > 0) {
  filtering_stats <- rbind(filtering_stats, 
                         data.frame(metric = "after_removing_testIDs_records", value = nrow(dat)))
}


# Update participant count
n_participants_final <- length(unique(dat$ID))
filtering_stats <- rbind(filtering_stats,
                         data.frame(metric = "final_unique_participants", value = n_participants_final))

cat("\n✓ Final unique participants:", n_participants_final)
```

## Extract questionnaire







```{r}
colnames(dat)[sapply(dat, is.list)]
```


```{r}
colnames(dat)
```

```{r}
tmp <- dat %>%
  select(-where(is.list)) %>%
  mutate(across(-ID, as.character)) %>%
  pivot_longer(-ID) %>%
  group_by(ID, name) %>%
  summarise(n_unique = n_distinct(value, na.rm = TRUE), .groups = 'drop') %>%
  filter(n_unique > 1)

tmp$name
```



```{r}
print("\n=== Creating Questionnaire DataFrame ===")


# ============================================================================
# PSYCHOLOGICAL SCALES - Organized by Type
# ============================================================================

# --- ACTIVISM & RADICALISM INTENTIONS (ARIS-adapted) ---
aris_activism <- c(
  "aris_scale-ais1",   # Join organization
  "aris_scale-ais2",   # Donate money
  "aris_scale-ais3",   # Volunteer time
  "aris_scale-ais4"    # Travel to rally
)

aris_radicalism <- c(
  "aris_scale-ris1",   # Support despite breaking law
  "aris_scale-ris2",   # Support despite violence
  "aris_scale-ris3",   # Protest despite violence
  "aris_scale-ris4"    # Attack security forces
)

# --- UTOPIAN THINKING (Fernando et al. 2018) ---
utopianism <- c(
  "utopianism_scale-utop1",   # Think about ideal society
  "utopianism_scale-utop2",   # Spend time thinking
  "utopianism_scale-utop3",   # What world I'd like
  "utopianism_scale-utop4"    # Important to think
)

anti_utopianism <- c(
  "anti_utopianism_scale-antiutop1",   # Could be dangerous
  "anti_utopianism_scale-antiutop2",   # Negative consequences
  "anti_utopianism_scale-antiutop3",   # Shouldn't envision
  "anti_utopianism_scale-antiutop4"    # Useless to dream
)

# --- NEED FOR CHAOS (NFC) - Arceneaux et al. 2021 ---
need_for_chaos <- c(
  "nfc_scale-nfc1",   # Kick from disasters
  "nfc_scale-nfc2",   # Fantasize wipeout
  "nfc_scale-nfc3",   # Burn society
  "nfc_scale-nfc4",   # Let them burn
  "nfc_scale-nfc5",   # Tear down & restart
  "nfc_scale-nfc6",   # Need chaos
  "nfc_scale-nfc7"    # Destroy beautiful things
)

# --- POLITICAL/INSTITUTIONAL ATTITUDES ---
system_justification <- c(
  "sjs_scale-sjs1",    # Society is fair
  "sjs_scale-sjs2",    # Political system works
  "sjs_scale-sjs3r",   # [REVERSE] Needs restructuring
  "sjs_scale-sjs4",    # Best country
  "sjs_scale-sjs5",    # Policies serve greater good
  "sjs_scale-sjs6",    # Fair shot at wealth
  "sjs_scale-sjs7r",   # [REVERSE] Getting worse
  "sjs_scale-sjs8"     # Get what they deserve
)

# --- WELL-BEING & RESILIENCE ---
satisfaction_with_life <- c(
  "swls_scale-swls1",  # Close to ideal
  "swls_scale-swls2",  # Excellent conditions
  "swls_scale-swls3",  # Satisfied with life
  "swls_scale-swls4",  # Got important things
  "swls_scale-swls5"   # Change almost nothing
)

brief_resilience <- c(
  "brs_scale-brs1",    # Bounce back quickly
  "brs_scale-brs2r",   # [REVERSE] Hard time through stress
  "brs_scale-brs3",    # Don't take long to recover
  "brs_scale-brs4r",   # [REVERSE] Hard to snap back
  "brs_scale-brs5",    # Come through with trouble
  "brs_scale-brs6r"    # [REVERSE] Long time to get over
)

# --- DEPRIVATION & CONSPIRACY ---
personal_relative_deprivation <- c(
  "prds_scale-prds1",  # Feel deprived
  "prds_scale-prds2r", # [REVERSE] Feel privileged
  "prds_scale-prds3",  # Feel resentful
  "prds_scale-prds4r", # [REVERSE] Quite well off
  "prds_scale-prds5"   # Dissatisfied
)

conspiracy_mentality <- c(
  "cmq_scale-cmq1",    # Important things hidden
  "cmq_scale-cmq2",    # Politicians don't tell truth
  "cmq_scale-cmq3",    # Government monitors
  "cmq_scale-cmq4",    # Secret activities
  "cmq_scale-cmq5"     # Secret organizations influence
)



# ============================================================================
# SOCIO-DEMOGRAPHIC VARIABLES, FEEDBACK
# ============================================================================
socio_demo <- c(
  "age",
  "gender",
  "education",
  "employment_status",
  "employment_is_employed",
  "lrscale",           # Left-right political scale
  "rlgdgr"             # Religious degree
)

# --- COMBINED SCALE VECTOR ---
vec_ques <- c(
  "PROLIFIC_PID",
  "dummy_informedconsent",
  "commCheck",
  "mic_test_passed",
  "feedback_critic",
  socio_demo,
  # scales:
  aris_activism,
  aris_radicalism,
  utopianism,
  anti_utopianism,
  need_for_chaos,
  system_justification,
  satisfaction_with_life,
  brief_resilience,
  personal_relative_deprivation,
  conspiracy_mentality
)


vec_notNumeric <- c(
  "PROLIFIC_PID",
  # Socio-demographic non-numeric
  "gender",
  "education",
  "employment_status",
  # Other non-numeric
  "mic_test_passed",
  "feedback_critic"
)


# ============================================================================
# Create questionnaire
# ============================================================================
ques_study <- questionnairetype(
  dataset = dat,
  listvars = vec_ques,
  notNumeric = vec_notNumeric,
  verbose = FALSE
)


ques_study$gender <- factor(x = ques_study$gender)
ques_study$education <- factor(x = ques_study$education)
ques_study$employment_status <- factor(x = ques_study$employment_status)


# Save summary
filtering_stats <- rbind(filtering_stats,
                         data.frame(metric = "final_questionnaire_rows", value = nrow(ques_study)))
```



# Add Transcripts

## Load transcript files

```{r}
#| label: load-transcripts

print("\n=== Loading Transcripts ===")

# Find all transcript files
transcript_files <- list.files(
  path = AUDIO_TRANSCRIPT_DIR,
  pattern = "_transcript\\.json$",
  full.names = TRUE
)

print(paste("Found", length(transcript_files), "transcript files"))

# Parse all transcripts
transcripts <- list()

for (file in transcript_files) {
  tryCatch({
    transcript <- fromJSON(file)
    filename <- basename(file)
    
    # Extract participant ID from filename
    # Format: {id_person}_{type}_{timestamp}_transcript.json
    participant_id <- str_extract(filename, "^[^_]+")
    audio_type <- str_extract(filename, "(?<=_)(audio_[^_]+)(?=_)")
    
    transcripts[[length(transcripts) + 1]] <- list(
      filename = filename,
      participant_id = participant_id,
      audio_type = audio_type,
      transcript_text = transcript$transcript,
      timestamp = transcript$timestamp
    )
    
  }, error = function(e) {
    print(paste("✗ Error loading", file, ":", e$message))
  })
}

print(paste("Successfully parsed", length(transcripts), "transcripts"))

# Convert to dataframe for easier processing
transcripts_df <- bind_rows(lapply(transcripts, as.data.frame, stringsAsFactors = FALSE))

if (VERBOSE) {
  print("Sample transcripts:")
  print(head(transcripts_df))
  print("\nTranscript types found:")
  print(table(transcripts_df$audio_type))
}
```

## Match transcripts to questionnaire and add

```{r}
#| label: add-transcripts-to-questionnaire

print("\n=== Adding Transcripts to Questionnaire ===")

# Initialize columns
ques_study$text_missingUtopia <- NA_character_
ques_study$text_rankingUtopias <- NA_character_

# For each participant in questionnaire
for (i in 1:nrow(ques_study)) {
  participant_id <- ques_study$PROLIFIC_PID[i]
  
  # Find matching transcripts
  matching_transcripts <- transcripts_df %>%
    filter(participant_id == !! participant_id)
  
  if (nrow(matching_transcripts) > 0) {
    # Process missing utopia transcripts
    missing_utopia <- matching_transcripts %>%
      filter(audio_type == "audio_missing") %>%
      arrange(desc(timestamp))
    
    if (nrow(missing_utopia) > 0) {
      # Keep only most recent
      ques_study$text_missingUtopia[i] <- missing_utopia$transcript_text[1]
      
      if (VERBOSE & nrow(missing_utopia) > 1) {
        print(paste("  Participant", participant_id, "- Kept 1 of", nrow(missing_utopia), "missingUtopia transcripts"))
      }
    }
    
    # Process ranking explanation transcripts
    ranking <- matching_transcripts %>%
      filter(audio_type == "audio_ranking") %>%
      arrange(desc(timestamp))
    
    if (nrow(ranking) > 0) {
      # Keep only most recent
      ques_study$text_rankingUtopias[i] <- ranking$transcript_text[1]
      
      if (VERBOSE & nrow(ranking) > 1) {
        print(paste("  Participant", participant_id, "- Kept 1 of", nrow(ranking), "rankingUtopias transcripts"))
      }
    }
  }
}

# Check coverage
n_with_missing_utopia <- sum(!is.na(ques_study$text_missingUtopia))
n_with_ranking <- sum(!is.na(ques_study$text_rankingUtopias))

print(paste("Transcripts added:"))
print(paste("  - text_missingUtopia:", n_with_missing_utopia, "/", nrow(ques_study)))
print(paste("  - text_rankingUtopias:", n_with_ranking, "/", nrow(ques_study)))

print("✓ Transcripts merged successfully")
```

---

# Add Sentiment Analysis

## Load overall sentiment data

```{r}
#| label: load-overall-sentiment

print("\n=== Loading Overall Sentiment Analysis ===")

# Find sentiment file
sentiment_files <- list.files(
  path = SENTIMENT_DIR,
  pattern = "overall_sentiment_and_embedding.*\\.csv$",
  full.names = TRUE
)

if (length(sentiment_files) > 0) {
  # Use most recent if multiple files exist
  sentiment_file <- sentiment_files[which.max(file.info(sentiment_files)$mtime)]
  
  print(paste("Using:", basename(sentiment_file)))
  
  sentiment_overall <- read.csv(sentiment_file, stringsAsFactors = FALSE)
  
  print(paste("Loaded", nrow(sentiment_overall), "sentiment records"))
  print("\nSample data:")
  print(head(sentiment_overall))
  
} else {
  print("✗ No overall sentiment file found!")
  sentiment_overall <- NULL
}
```

## Add overall sentiment to questionnaire

```{r}
#| label: add-overall-sentiment

if (!is.null(sentiment_overall)) {
  print("\n=== Adding Overall Sentiment to Questionnaire ===")
  
  # Initialize columns
  ques_study$sentiment_label <- NA_character_
  ques_study$sentiment_score <- NA_real_
  ques_study$num_windows <- NA_integer_
  
  # For each participant
  for (i in 1:nrow(ques_study)) {
    participant_id <- ques_study$PROLIFIC_PID[i]
    
    # Find matching sentiment records
    # Match using json_file which contains participant_id
    matching_sentiment <- sentiment_overall %>%
      filter(grepl(participant_id, json_file, fixed = TRUE))
    
    if (nrow(matching_sentiment) > 0) {
      # If multiple matches, use most recent (first one after sort)
      # For now, combine or average if multiple
      
      if (nrow(matching_sentiment) > 1) {
        # Average sentiment scores, keep first label
        ques_study$sentiment_label[i] <- matching_sentiment$sentiment_label[1]
        ques_study$sentiment_score[i] <- mean(matching_sentiment$sentiment_score, na.rm = TRUE)
        ques_study$num_windows[i] <- sum(matching_sentiment$num_windows, na.rm = TRUE)
        
        if (VERBOSE) {
          print(paste("  Participant", participant_id, "- Averaged", nrow(matching_sentiment), "sentiment records"))
        }
      } else {
        ques_study$sentiment_label[i] <- matching_sentiment$sentiment_label[1]
        ques_study$sentiment_score[i] <- matching_sentiment$sentiment_score[1]
        ques_study$num_windows[i] <- matching_sentiment$num_windows[1]
      }
    }
  }
  
  n_with_sentiment <- sum(!is.na(ques_study$sentiment_label))
  print(paste("Overall sentiment added to", n_with_sentiment, "participants"))
  
} else {
  print("Skipping overall sentiment (file not found)")
}
```

## Load window-level sentiment data

```{r}
#| label: load-window-sentiment

print("\n=== Loading Window-Level Sentiment ===")

# Find window sentiment file
window_files <- list.files(
  path = SENTIMENT_DIR,
  pattern = "window_sentiments.*\\.csv$",
  full.names = TRUE
)

if (length(window_files) > 0) {
  # Use most recent if multiple files exist
  window_file <- window_files[which.max(file.info(window_files)$mtime)]
  
  print(paste("Using:", basename(window_file)))
  
  sentiment_windows <- read.csv(window_file, stringsAsFactors = FALSE)
  
  # Calculate max window index
  max_window_idx <- max(sentiment_windows$window_index, na.rm = TRUE)
  
  print(paste("Loaded", nrow(sentiment_windows), "window sentiment records"))
  print(paste("Maximum window index:", max_window_idx))
  
  # Show sample
  print("\nSample data:")
  print(head(sentiment_windows))
  
} else {
  print("✗ No window sentiment file found!")
  sentiment_windows <- NULL
  max_window_idx <- 0
}
```

## Add window sentiment columns

```{r}
#| label: add-window-sentiment

if (!is.null(sentiment_windows)) {
  print("\n=== Adding Window-Level Sentiment Columns ===")
  
  # Create column names dynamically
  window_cols <- paste0("window_", 0:max_window_idx, "_sentimentLabel")
  
  print(paste("Creating", length(window_cols), "window sentiment columns"))
  
  # Initialize columns with NA
  for (col in window_cols) {
    ques_study[[col]] <- NA_character_
  }
  
  # Fill window sentiment for each participant
  for (i in 1:nrow(ques_study)) {
    participant_id <- ques_study$PROLIFIC_PID[i]
    
    # Find windows for this participant
    participant_windows <- sentiment_windows %>%
      filter(grepl(participant_id, json_file, fixed = TRUE))
    
    if (nrow(participant_windows) > 0) {
      # Fill window columns
      for (j in 1:nrow(participant_windows)) {
        window_idx <- participant_windows$window_index[j]
        col_name <- paste0("window_", window_idx, "_sentimentLabel")
        
        # Only fill if column exists
        if (col_name %in% colnames(ques_study)) {
          ques_study[i, col_name] <- participant_windows$sentiment_label[j]
        }
      }
      
      if (VERBOSE & i %in% c(1, 5, 10)) {
        print(paste("  Participant", participant_id, "- Added", nrow(participant_windows), "window sentiments"))
      }
    }
  }
  
  print(paste("✓ Window sentiment columns added"))
  
} else {
  print("Skipping window sentiment (file not found)")
}
```
```{r}
# Before
cat("Original columns:", ncol(ques_study), "\n")
cat("Columns that are all NA:", sum(sapply(ques_study, function(x) all(is.na(x)))), "\n")

# Apply removal (Option 1)
ques_study <- ques_study[, colSums(is.na(ques_study)) < nrow(ques_study)]

# After
cat("Final columns:", ncol(ques_study), "\n")
```


---

# Get within part of study

## visual trap part

```{r}
# --- Get one PROLIFIC_PID per ID ---
pid_lookup <- dat %>%
  group_by(ID) %>%
  summarise(PROLIFIC_PID = last_non_empty(PROLIFIC_PID), .groups = 'drop')


trap_types <- c("Planets", "Robot", "Circles", "Lines", "CafeWall", "Collision")

visual_trap_within <- map_df(trap_types, function(trap) {
  dat %>%
    select(
      ID,
      counterVisualTraps,
      all_of(paste0("visualTrap_", trap, "_response")),
      all_of(paste0("visualTrap_", trap, "_responseLabel")),
      all_of(paste0("visualTrap_", trap, "_correct")),
      all_of(paste0("visualTrap_", trap, "_rtMs"))
    ) %>%
    # Remove rows where this trap is all NA
    filter(!if_all(starts_with("visualTrap_"), is.na)) %>%
    distinct(ID, counterVisualTraps, .keep_all = TRUE) %>%
    mutate(
      trap_type = trap,
      response = .data[[paste0("visualTrap_", trap, "_response")]],
      response_label = .data[[paste0("visualTrap_", trap, "_responseLabel")]],
      correct = .data[[paste0("visualTrap_", trap, "_correct")]],
      rt_ms = .data[[paste0("visualTrap_", trap, "_rtMs")]]
    ) %>%
    select(ID, counterVisualTraps, trap_type, response, response_label, correct, rt_ms)
}) %>%
  # Add PROLIFIC_PID from lookup
  left_join(pid_lookup, by = "ID") %>%
  select(ID, PROLIFIC_PID, counterVisualTraps, trap_type, everything()) %>%
  arrange(ID, counterVisualTraps, trap_type)

# --- Inspect results ---
head(visual_trap_within)
cat("Total vignette trials:", nrow(visual_trap_within), "\n")
cat("Trials per participant:\n")
print(table(visual_trap_within$ID))
```


## vignette part

```{r}
# Helper function
last_non_empty <- function(x) {
  if (length(x) == 0) return(NA)
  for (i in seq_along(x)[length(x):1]) {
    val <- x[[i]]
    if (is.null(val)) next
    if (!is.list(val) && length(val) == 1 && is.na(val)) next
    if (is.character(val) && length(val) == 1 && isTRUE(val == "")) next
    return(val)
  }
  return(NA)
}

# --- Get PROLIFIC_PID mapping ---
pid_lookup <- dat %>%
  group_by(ID) %>%
  summarise(PROLIFIC_PID = last_non_empty(PROLIFIC_PID), .groups = 'drop')

# --- Initialize results list ---
vignette_list <- list()

# --- Get unique IDs ---
unique_ids <- unique(dat$ID)

# --- LOOP 1: For each participant ---
for (id in unique_ids) {
  
  # Filter data for this participant
  dat_id <- dat %>% filter(ID == id)
  
  # Get their PROLIFIC_PID
  prolific_pid <- pid_lookup %>% 
    filter(ID == id) %>% 
    pull(PROLIFIC_PID)
  
  # --- LOOP 2: For each vignette trial (counterScenarios) ---
  unique_trials <- unique(na.omit(dat_id$counterScenarios))
  
  for (trial in unique_trials) {
    
    # Filter data for this trial
    trial_data <- dat_id %>% filter(counterScenarios == trial)
    
    # Extract values (take last non-empty per variable)
    row <- tibble(
      ID = id,
      PROLIFIC_PID = prolific_pid,
      counterScenarios = trial,
      vignette_order_position = last_non_empty(trial_data$vignette_order_position),
      vignette_id = last_non_empty(trial_data$vignette_id),
      condition_FutureSociety = last_non_empty(trial_data$condition_FutureSociety),
      # Participative efficacy
      pe1 = last_non_empty(trial_data$pe1),
      pe2 = last_non_empty(trial_data$pe2),
      pe3 = last_non_empty(trial_data$pe3),
      pe4 = last_non_empty(trial_data$pe4),
      # Attributes
      att_ideal = last_non_empty(trial_data$att_ideal),
      att_beneficial = last_non_empty(trial_data$att_beneficial),
      att_utopian = last_non_empty(trial_data$att_utopian),
      att_desirable = last_non_empty(trial_data$att_desirable),
      att_innovative = last_non_empty(trial_data$att_innovative),
      att_live_in = last_non_empty(trial_data$att_live_in),
      # Paradata
      para_countclicks = last_non_empty(trial_data$para_countclicks),
      time_run = last_non_empty(trial_data$time_run),
      time_render = last_non_empty(trial_data$time_render),
      time_show = last_non_empty(trial_data$time_show)
    )
    
    # Add to results list
    vignette_list[[length(vignette_list) + 1]] <- row
  }
  
  # Progress indicator
  # cat("Processed ID:", id, "\n")
}

# --- Combine all results into one dataframe ---
ques_study_within_vignettes <- bind_rows(vignette_list)

# --- Remove trials with all NA in PE/ATT ---
ques_study_within_vignettes <- ques_study_within_vignettes %>%
  filter(!if_all(pe1:att_live_in, is.na)) %>%
  arrange(ID, counterScenarios)

# --- Inspect results ---
head(ques_study_within_vignettes)
cat("Total vignette trials:", nrow(ques_study_within_vignettes), "\n")
cat("Trials per participant:\n")
print(table(ques_study_within_vignettes$ID))
```



# Summary stats and Quality checks


```{r}
print("FINAL DATA SUMMARY & VALIDATION")
print(strrep("=", 80))
 
# --- Count unique IDs across datasets ---
n_ids_wide <- n_distinct(ques_study$ID)
n_ids_vignettes <- n_distinct(ques_study_within_vignettes$ID)
n_ids_traps <- n_distinct(visual_trap_within$ID)
 
print("\n📊 PARTICIPANT COUNTS:")
print(paste("  - Between-subjects (ques_study):", n_ids_wide, "unique IDs"))
print(paste("  - Within-vignettes dataset:", n_ids_vignettes, "unique IDs"))
print(paste("  - Within-traps dataset:", n_ids_traps, "unique IDs"))
```

